### 2025-01-14 main.tex
first commit

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc} % For UTF-8 encoding
\usepackage{amsmath, amssymb} % For mathematical symbols
\usepackage{graphicx} % For including images
\usepackage{hyperref} % For clickable links
\usepackage{geometry} % For page margins
\usepackage{booktabs} % For nice tables
\usepackage{lipsum} % For dummy text
\usepackage{longtable}
\setlength{\parindent}{0em}
% Page layout
\geometry{margin=1in}

% Hyperlink setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
    pdftitle={LaTeX Template},
    pdfpagemode=FullScreen,
}

% ######################## Title  Page ####################################
\title{Recurrent Deep Learning Models and its applications}
\author{Ngai Ho Wang}

\date{\today}

\begin{document}

\maketitle
% ######################## Title  Page ####################################

% ######################## table of content ###############################
\tableofcontents % Generate a table of contents                        
% ######################## table of content ###############################

\newpage

% ######################## Project goals ######################################
\section{Project goals}
The goal of this paper is to analyze, implement, and compare the performance of RNN, LSTM, GRU and own model in selected NLP tasks. This paper aims to propose an enhanced RNN-based model for NLP tasks and make the recommendation for future development.
% ######################## Project goals ######################################

% ######################## Objectives ######################################
\section{Objectives}
\begin{enumerate}
    % 1. Literature Review
    \item Literature Review
    \begin{itemize}
        \item Conduct a comprehensive review of existing paper on RNN, LSTM, and GRU, focusing on their architecture.
        \item Review the NLP task.
    \end{itemize}

    % 2. Model implementation
    \item Model implementation
    \begin{itemize}
        \item Implement RNN, LSTM, GRU and own model by PyTorch for chosen NLP tasks.
    \end{itemize}

    % 3. Performance Comparison
    \item Performance Comparison
    \begin{itemize}
        \item Evaluate the performance by using appropriate metrics (e.g., accuracy, precision, recall, F1 score, BLEU, ROUGE).
    \end{itemize}

    % 4. Propose an advanced model
    \item Propose an advanced model
    \begin{itemize}
        \item Develop an advanced model, aiming to enhance the performance on selected NLP tasks. 
    \end{itemize}
\end{enumerate}
% ######################## Objectives ######################################

% ######################## Expected outcomes ######################################
\section{Expected outcomes}
\begin{enumerate}
    % 1. Model performance metrics
    \item Model performance metrics
    \begin{itemize}
        \item A detailed comparison of performance metrics across RNN, LSTM, GRU and own model. 
    \end{itemize}

    % 2. Practical insights
    \item Practical insights
    \begin{itemize}
        \item Practical recommendations for future work in the area of RNN applications in NLP, based on the findings of this paper. 
    \end{itemize}
\end{enumerate}
% ######################## Expected outcomes ######################################

% ######################## Research plan ######################################
\section{Research plan}
\begin{enumerate}
    % 1. Literature review
    \item Literature review
    \begin{itemize}
        \item Review existing work on RNN, LSTM, GRU and their more sophisticated variants. 
    \end{itemize}

    % 2. Model development
    \item Model development
    \begin{itemize}
        \item Select NLP tasks. (e.g., text classification, summarization).
        \item Implement baseline models (RNN, LSTM, GRU) using PyTroch.
    \end{itemize}

    % 3. Performance Evaluation
    \item Performance Evaluation
    \begin{itemize}
        \item Evaluate and compare the performance of the implemented model using appropriate metrics. 
    \end{itemize}

    % 4. Model Enhancement
    \item Model Enhancement
    \begin{itemize}
        \item Design and implement an advanced model based on findings from the previous phases.
    \end{itemize}

    % 5. Final analysis and reporting
    \item Final analysis and reporting
    \begin{itemize}
        \item Analyze the results of the advanced model against baseline models. 
    \end{itemize}
\end{enumerate}
% ######################## Research plan ######################################

% ######################## Experimental design ######################################
\section{Experimental design}
\begin{enumerate}
    % 1. Dataset selection
    \item Dataset selection
    \begin{itemize}
        \item Select suitable NLP dataset.
    \end{itemize}

    % 2. Model configuration
    \item Model configuration
    \begin{itemize}
        \item Define architecture specifications for each model, including number of layers, number of hidden units, activation function, dropout rates.
    \end{itemize}
\end{enumerate}
% ######################## Experimental design ######################################

\newpage
% ######################## Introduction ######################################
\section{Introduction}

With the rise of the Generative Artificial Intelligence, the development of AI has already made remarkable strides in processing sequential data. In understanding and producing sequential data. It has applications ranging from Natural Language Processing (NLP) to music composition to video generation. Especially NLP, has emerged as a pivotal field in artificial intelligence, enable machines to understand, interpret and generate in human readable format. Siri, Alexa and bixby have shown the possibility. Everyone can communicate with those machines and they with make the reasonable response to user.\\[2ex]
Recurrent Neural Networks (RNNs) have been a foundational architecture in this domain, the architecture of RNNs is design for sequential data. It able to retain the information through hidden states. Unfortunately, early RNNs had limitation in training of networks over long sequence. vanishing and exploding gradient problems significantly affect the training process of RNN (Bengio, Simard, \& Frasconi, 1994). Eliminating many practical applications of RNNs. After that, Hochreiter and Schmidhuber (1997) introduced Long Short-Term Memory (LSTM) networks and are responsible for the breakthrough in how to solve these challenges. Specificized gating mechanisms were introduced in LSTMs to regulate the flow of the information, minimize the vanishing gradient problem and learn the long-term dependencies. This advanced made RNNs much more performant on tasks like a language modeling, machine translation and speech recognition tasks.\\[2ex]
Further improvements were achieved with Gated Recurrent Units (GRUs) by Cho et al. (2014) which diminished the LSTM architecture's complexity, but still provided the same performance. GRUs performed comparably but used fewer parameters, making it computationally and more tractably trainable.
% ######################## Introduction ######################################

\newpage
% ######################## Literature Review ######################################
\section{Literature Review}
\textbf{Backpropagation Through Time}

BPTT is one of the most important algorithms used for training RNNs. Dating back to the original effort to expand the typical backpropagation algorithm, BPTT has been formulated to handle the difficulties of temporal sequences that are inherent in sequential data (Werbos, 1990). This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable (Rumelhart, Hinton, \& Williams, 1986).
\newline
\textbf{Conceptual Framework of BPTT}

BPTT works based on the technique of treating an RNN as a deep feedforward network for across multiple time steps. In the forward pass, the RNN, like other artificial neuronal network, applies operation over the data input in sequence, bringing changes in its own state variables at every time step, depending on the input and the previous state of its general working state or hidden state. This sequential processing produces outputs and stores the internal states of the network in any period (Werbos, 1990).

This unfolds the RNN to construct a traditional Feedforward Neural Network where we can apply backpropagation through time. Below is the conceptual idea of BPTT in RNN.
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{./Pic/pic1.png} % Replace "example-image" with your image file
    \caption{Unfolded RNN}
    % \label{fig:sdfsdf}
\end{figure}

\begin{longtable}{|c|c|c|}
    \hline
    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
    \hline
    $U$          & Weight matrix for input to hidden state       & $input\ size\times hidden\ unites$\\
    $W$          & Weight matrix for hidden to hidden state      & $hidden\ units\times hidden\ unites$\\
    $V$          & Weight matrix for hidden state to output state& $hidden\ units\times number\ of\ class$\\
    $x_t$        & Input vector at time t                        & $input\ size\times 1$\\
    $h_t$        & Hidden state output at time t                 & $hidden\ units\times 1$\\
    $b_h$        & Bias term for hidden state                    & $hidden\ units\times 1$\\
    $b_y$        & Bias term for output state                    & $number\ of\ class\times 1$\\
    $\hat{o}_y$  & Output at time t                              & $number\ of\ class\times 1$\\
    $\hat{y}_t$  & Output at time t                              & $hidden\ units\times 1$\\
    $\mathcal{L}$& Loss at time t                                & $scalar$\\
    \hline
    \caption{Unfolded RNN}
\end{longtable}

% Forward pass
\noindent \textbf{Forward Pass}\\
During the forward pass, the RNN processes the input sequence sequentially, computing hidden states and output at each timestep:
\begin{equation}
    h_t = f(U^Tx_{t}+W^Th_{t-1}+b_h)
\end{equation}
\begin{equation}
    \hat{y}_t = f(V^Th_t+b_y)
\end{equation}
% \begin{equation}
%     \hat{y}_t = f(\hat{o}_t)
% \end{equation}
\newline  % Computing the loss function
\noindent \textbf{Computing the loss function}\\
Assuming the loss is computed only at the final timestep t:
\begin{equation}
    \mathcal{L}_t = L(y_t, \hat{y}_t)
\end{equation}
In order to do backpropagation through time to tune the parameters in RNN, we need to calculate the partial derivative of loss function $\mathcal{L}$ with respect to the differently parameters.\\
\newline  % Backward pass using the chain rule
\noindent \textbf{Backward pass using the chain rule}\\
Using the chain rule for computing the gradient.\\
Partial derivative of loss function $\mathcal{L}$ with respect to $W$ (hidden to hidden state) at time 2. 
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial{W}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial W}
\end{equation}
\begin{equation}
    \dfrac{\partial L_i}{\partial W} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial W}
\end{equation}
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial W} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial W}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial W}
\end{equation}

Partial derivative of loss function $\mathcal{L}$ with respect to $U$ (input to hidden state) at time 2.\\
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial{U}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial U}
\end{equation}
\begin{equation}
    \dfrac{\partial L_i}{\partial U} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial U}
\end{equation}
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial U} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial U}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial U}
\end{equation}

Partial derivative of loss function $\mathcal{L}$ with respect to $V$ (hidden to output state) at time 2.\\
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial{V}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial V}
\end{equation}
\begin{equation}
    \dfrac{\partial L_i}{\partial V} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial V}
\end{equation}
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial V} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial V}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial V}
\end{equation}

Partial derivative of loss function $\mathcal{L}$ with respect to $b_h$ (bias term in hidden state) at time 2.\\
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial{b_h}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial b_h}
\end{equation}
\begin{equation}
    \dfrac{\partial L_i}{\partial b_h} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_h}
\end{equation}
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial b_h} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial b_h}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial b_h}
\end{equation}

Partial derivative of loss function $\mathcal{L}$ with respect to $b_y$ (bias term in output state) at time 2.
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial{b_y}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial b_y}
\end{equation}
\begin{equation}
    \dfrac{\partial L_i}{\partial b_y} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_y}
\end{equation}
\begin{equation}
    \dfrac{\partial\mathcal{L}_2}{\partial b_y} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial b_y}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial b_y}
\end{equation}

\textbf{parameters updates}
\begin{equation}
    W \leftarrow W - \alpha\dfrac{\partial\mathcal{L}}{\partial W}
\end{equation}
\begin{equation}
    U \leftarrow U - \alpha\dfrac{\partial\mathcal{L}}{\partial U}
\end{equation}
\begin{equation}
    V \leftarrow V - \alpha\dfrac{\partial\mathcal{L}}{\partial V}
\end{equation}
\begin{equation}
    b_h \leftarrow b_h - \alpha\dfrac{\partial\mathcal{L}}{\partial b_h}
\end{equation}
\begin{equation}
    b_y \leftarrow b_y - \alpha\dfrac{\partial\mathcal{L}}{\partial b_y}
\end{equation}

\textbf{Pseudocode of BPTT} (Wikipedia, 2023)
\begin{center}
    \begin{tabular}{c}
        $Back_Propagation_Through_Time(a, y)   % a[t] is the input at time t. y[t] is the output
        Unfold the network to contain k instances of f
        do until stopping criterion is met:
        x := the zero-magnitude vector % x is the current context
        % for t from 0 to n − k do      % t is time. n is the length of the training sequence
        % Set the network inputs to x, a[t], a[t+1], ..., a[t+k−1]
        % p := forward-propagate the inputs over the whole unfolded network
        % e := y[t+k] − p;           % error = target − prediction
        Back-propagate the error, e, back across the whole unfolded network
        Sum the weight changes in the k instances of f together.
        Update all the weights in f and g.
        x := f(x, a[t]);$           % compute the context for the next time-step

    \end{tabular}
\end{center}
% ######################## Literature Review ######################################

\section{Inserting Images}
To insert an image, use the `graphicx` package. For example:

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{./Pic/pic1.png} % Replace "example-image" with your image file
    \caption{An example image.}
    \label{fig:example}
\end{figure}

\section{Tables}
You can create tables using the `tabular` environment or `booktabs` for professional-quality tables. For example:

\begin{table}[h!]
\centering
\caption{Example Table}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Item} & \textbf{Description} & \textbf{Quantity} \\ \midrule
Apples        & Fresh red apples     & 10                \\
Oranges       & Juicy oranges        & 5                 \\
Bananas       & Ripe bananas         & 7                 \\ \bottomrule
\end{tabular}
\label{tab:example}
\end{table}

\section{Hyperlinks}
To add a hyperlink, use the `hyperref` package. For example:
\href{https://www.latex-project.org/}{Visit the LaTeX project website}.

\section{Conclusions}
This is the conclusion section. Summarize your findings or leave final remarks.

\newpage
\appendix
\section{Appendix}
This is the appendix section, where you can include supplementary materials.

\end{document}

### 2025-01-15 test.py
add model and other stuff

@@ -1,0 +1,163 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+from torch.utils.data import DataLoader, Dataset
+from transformers import AutoTokenizer
+from datasets import load_dataset
+import numpy as np
+
+# Step 1: Load the SQuAD dataset
+squad = load_dataset("squad")
+
+# Extract data
+train_data = squad['train']
+contexts = [item['context'] for item in train_data]
+questions = [item['question'] for item in train_data]
+answers = [item['answers']['text'][0] for item in train_data]
+answer_starts = [item['answers']['answer_start'][0] for item in train_data]
+
+# Step 2: Preprocess the data
+tokenizer = AutoTokenizer.from_pretrained("bert-base-uncased")
+
+# Tokenize context and question
+def preprocess_data(context, question, answer, answer_start):
+    tokenized_context = tokenizer(context, truncation=True, padding="max_length", max_length=300, return_tensors="pt")
+    tokenized_question = tokenizer(question, truncation=True, padding="max_length", max_length=50, return_tensors="pt")
+    
+    # Compute answer tokens
+    answer_end = answer_start + len(answer)
+    token_start = tokenized_context.char_to_token(answer_start)
+    token_end = tokenized_context.char_to_token(answer_end - 1)
+    
+    if token_start is None or token_end is None:
+        # If the answer cannot be tokenized, return None (to filter out later)
+        return None
+    
+    return {
+        "context": tokenized_context["input_ids"].squeeze(0),
+        "question": tokenized_question["input_ids"].squeeze(0),
+        "start_pos": token_start,
+        "end_pos": token_end
+    }
+
+# Preprocess the entire dataset
+processed_data = []
+for c, q, a, s in zip(contexts, questions, answers, answer_starts):
+    item = preprocess_data(c, q, a, s)
+    if item:
+        processed_data.append(item)
+
+# Convert to PyTorch Dataset
+class SquadDataset(Dataset):
+    def __init__(self, data):
+        self.data = data
+
+    def __len__(self):
+        return len(self.data)
+
+    def __getitem__(self, idx):
+        item = self.data[idx]
+        return item["context"], item["question"], item["start_pos"], item["end_pos"]
+
+dataset = SquadDataset(processed_data)
+dataloader = DataLoader(dataset, batch_size=16, shuffle=True)
+
+# Step 3: Define the RNN-based model
+class RNNQuestionAnsweringModel(nn.Module):
[…]

### 2025-01-21 main.tex
window commit

@@
 \usepackage{longtable}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \setlength{\parindent}{0em}
@@
 
+\newpage
 \begin{longtable}{|c|c|c|}
@@
 
+\newpage
 \textbf{Pseudocode of BPTT} (Wikipedia, 2023)
-\begin{center}
-    \begin{tabular}{c}
-        $Back_Propagation_Through_Time(a, y)   % a[t] is the input at time t. y[t] is the output
-        Unfold the network to contain k instances of f
-        do until stopping criterion is met:
-        x := the zero-magnitude vector % x is the current context
-        % for t from 0 to n − k do      % t is time. n is the length of the training sequence
-        % Set the network inputs to x, a[t], a[t+1], ..., a[t+k−1]
-        % p := forward-propagate the inputs over the whole unfolded network
-        % e := y[t+k] − p;           % error = target − prediction
-        Back-propagate the error, e, back across the whole unfolded network
-        Sum the weight changes in the k instances of f together.
-        Update all the weights in f and g.
-        x := f(x, a[t]);$           % compute the context for the next time-step
+\begin{algorithm}[H]
+    \caption{Backpropagation Through Time (BPTT)}
+    \begin{algorithmic}[1]
+    \STATE \textbf{Input:} 
+    \STATE \hspace{1em} Sequence of input data $\{x_1, x_2, \dots, x_T\}$
+    \STATE \hspace{1em} Sequence of target outputs $\{y_1, y_2, \dots, y_T\}$
+    \STATE \hspace{1em} Learning rate $\eta$
+    \STATE \hspace{1em} Number of time steps to unroll $N$
+    \STATE \textbf{Initialize:} Model parameters $\theta$, hidden state $h_0 = 0$
 
-    \end{tabular}
-\end{center}
+    \STATE \textbf{Forward Pass:}
+    \FOR{$t = 1$ to $T$}
+        \STATE Compute hidden state: $h_t = f(h_{t-1}, x_t; \theta)$
+        \STATE Compute output: $\hat{y}_t = g(h_t; \theta)$
+        \STATE Compute loss for time step $t$: $L_t = \mathcal{L}(\hat{y}_t, y_t)$
+    \ENDFOR
+
+    \STATE \textbf{Backward Pass (BPTT):}
+    \STATE Set total loss: $L = \sum_{t=1}^{T} L_t$
+    \FOR{$t = T$ down to $1$}
+        \STATE Compute gradient of loss with respect to output: $\frac{\partial L_t}{\partial \hat{y}_t}$
+        \STATE Backpropagate through output layer to obtain: $\frac{\partial L_t}{\partial h_t}$
+        \STATE Accumulate gradients for parameters: $\frac{\partial L}{\partial \theta}$
+        \FOR{$k = 1$ to $N$}
+            \STATE Backpropagate through time for $N$ steps:
+            \STATE Compute gradient contribution from step $t-k$: $\frac{\partial L_t}{\partial h_{t-k}}$
+        \ENDFOR
+    \ENDFOR
+
+    \STATE \textbf{Update Parameters:}
+    \STATE $\theta = \theta - \eta \cdot \frac{\partial L}{\partial \theta}$
+
+    \STATE \textbf{Output:} Updated parameters $\theta$
+    \end{algorithmic}
+\end{algorithm}
+
+\newpage
+% Actication function
+\textbf{Activation function}\\
+Activation functions, particularly the sigmoid function, are fundamental components of recurrent neural networks (RNNs).
+They transform input data into output data.
+A key property of these functions is their differentiability.
+Differentiability is crucial for the backpropagation through time (BPTT) algorithm, enabling the application of the chain rule during training.
+\begin{equation}
+    Sigmoid(x) = \frac{ 1 }{ 1+e^{-x} }
+\end{equation}
+\begin{equation}
+    Sigmoid'(x) = Sigmoid(x)(1-Sigmoid(x))
+\end{equation}
+Below is the sigmoid function and its derivative.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{./Pic/sigmoid.png} % Replace "example-image" with your image file
+    % \caption{An example image.}
+    \label{fig:sigmoid}
+\end{figure}
+\newline
+$Domain(Sigmoid(x))=\mathbb{R},\hspace{1em} Codomain(Sigmoid(x))=(0,1)$\\
+$Domain(Sigmoid'(x))=\mathbb{R},\hspace{1em} Codomain(Sigmoid'(x))=[0,0.5]$
+
+% Hyperbolic tangent activation function
+\newpage
+\textbf{Hyperbolic tangent activation function}\\
+The main role of the hyperbolic tangent (tanh) activation function is to normalize candidate values and convert the cell state to a hidden state when performing cell state updates.
+It limits the output between [-1,1] because it has a stable gradient, which is important for discovering long-range dependencies.
+\newline
+\begin{equation}
+    tanh(x) = \frac{ e^{x} - e^{-x} }{e^{x} + e^{-x}}
+\end{equation}
+\begin{equation}
+    tanh'(x) = 1 - tanh^{2}(x)
+\end{equation}
+Below is the Hyperbolic tangent activation function and its derivative.\\
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{./Pic/tanh.png} % Replace "example-image" with your image file
+    % \caption{An example image.}
+    \label{fig:tanh}
+\end{figure}
+\newline
+$Domain(tanh(x))=\mathbb{R},\hspace{1em} Codomain(tanh(x))=[-1,1]$\\
+$Domain(tanh'(x))=\mathbb{R},\hspace{1em} Codomain(tanh'(x))=[0,1]$
+
+\newpage
+\textbf{Gradient vanishing and gradient exploring}\\
+When training the RNN, BPTT was used to update the weight matrix.
+As the number of time steps increase, the problem of gradient instability of often encountered, and this problem is gradient vanishing and gradient explored (Bengio et al.
+1994).
 % ######################## Literature Review ######################################
@@
 \section{Inserting Images}
+asdfpiojaseopritjf
 To insert an image, use the `graphicx` package. For example:

### 2025-01-21 paper/main.tex
change directory

--- 2025-01-14 main.tex
@@
     \centering
-    \includegraphics[width=1\textwidth]{./Pic/pic1.png} % Replace "example-image" with your image file
+    \includegraphics[width=1\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
     \caption{Unfolded RNN}
@@
     \centering
-    \includegraphics[width=1\textwidth]{./Pic/pic1.png} % Replace "example-image" with your image file
+    \includegraphics[width=1\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
     \caption{An example image.}

### 2025-01-21 paper/main.tex
merge

@@
     \centering
-    \includegraphics[width=1\textwidth]{./Pic/sigmoid.png} % Replace "example-image" with your image file
+    \includegraphics[width=1\textwidth]{../Pic/sigmoid.png} % Replace "example-image" with your image file
     % \caption{An example image.}
@@
     \centering
-    \includegraphics[width=1\textwidth]{./Pic/tanh.png} % Replace "example-image" with your image file
+    \includegraphics[width=1\textwidth]{../Pic/tanh.png} % Replace "example-image" with your image file
     % \caption{An example image.}

### 2025-01-21 paper/main.tex
add something

@@
 \textbf{Backpropagation Through Time}
-
+\\[1ex]
 BPTT is one of the most important algorithms used for training RNNs.
@@
 This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable (Rumelhart, Hinton, \& Williams, 1986).
-\newline
+\\[2ex]
 \textbf{Conceptual Framework of BPTT}
-
+\\[1ex]
 BPTT works based on the technique of treating an RNN as a deep feedforward network for across multiple time steps.
@@
 % Forward pass
-\noindent \textbf{Forward Pass}\\
+\noindent \textbf{Forward Pass}
+\\[1ex]
 During the forward pass, the RNN processes the input sequence sequentially, computing hidden states and output at each timestep:
@@
 \newline  % Computing the loss function
-\noindent \textbf{Computing the loss function}\\
+\noindent \textbf{Computing the loss function}
+\\[1ex]
 Assuming the loss is computed only at the final timestep t:
@@
 \newline  % Backward pass using the chain rule
-\noindent \textbf{Backward pass using the chain rule}\\
+\noindent \textbf{Backward pass using the chain rule}
+\\[1ex]
 Using the chain rule for computing the gradient.\\
@@
 % Actication function
-\textbf{Activation function}\\
+\textbf{Activation function}
+\\[1ex]
 Activation functions, particularly the sigmoid function, are fundamental components of recurrent neural networks (RNNs).
@@
 \newpage
-\textbf{Hyperbolic tangent activation function}\\
+\textbf{Hyperbolic tangent activation function}
+\\[1ex]
 The main role of the hyperbolic tangent (tanh) activation function is to normalize candidate values and convert the cell state to a hidden state when performing cell state updates.
@@
 
+% Gradient vanishing and gradient exploring
 \newpage
-\textbf{Gradient vanishing and gradient exploring}\\
+\textbf{Gradient vanishing and gradient exploring}
+\\[1ex]
 When training the RNN, BPTT was used to update the weight matrix.
@@
 1994).
+\\[2ex]
+% Vanishing Gradients
+\textbf{Vanishing Gradients}
+\\[1ex]
+Generally, sigmoid activation function is used commonly in RNNs, has a maximum derivative of 0.25.
+When doing BPTT in long time steps, this multiplication results in exponentially diminishing gradients as the sequence length increases.
+Consequently, the shallow neural receive very small gradient updates, making it difficult to adjust the parameters effectively.
+This leads to the model struggling to learn long time dependencies.
+\\[2ex]
+% Exploding Gradients
+\textbf{Exploding Gradients}
+\\[1ex]
+When we are doing the feedforward and get super large value computed by loss function.
+Then when updating the parameters.
+The updates to the weights will also be large.
+Resulting in higher loss and larger gradients in the next iterations.
+This will lead to exploding gradients.
+
+\newpage
+% Long short-term memory (LSTM)
+\textbf{Long short-term memory (LSTM)}
+\\[2ex]
+Long short-term memory proposed by Hochreiter \& Schmidhuber (1997).
+LSTM is designed for handling long time step problems.
+The architecture of LSTM can prevent vanishing gradient and exploring gradient.
+The main difference between LSTM and RNN is the number of gates.
+LSTM introduced input, forget and output gates.
+This allows LSTM to manage the flow of information more effectively, retaining important information over longer sequences.
+\\[2ex]
+Architecture
+\\[1ex]
+LSTMs introduce a memory cell that can maintain information over long time steps the cell is controlled by three gates, input gate, output gate, and forget gate.
+Each cell of LSTMs inside has 3 sigmoid and 1 tanh layer.
+Below graph unfolds the LSTM hence we can analyze different gates.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm1.png} % Replace "example-image" with your image file
+\end{figure}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm2.png} % Replace "example-image" with your image file
+\end{figure}
+\\
+% Forget gate
+Forget gate:
+\\[1ex]
+The forget gate is a component of the LSTM, designed to manage the flow of information within the cell state.
+The function of forget gate is to determine which information should be retained in memory cell (Hochreiter, S., \& Schmidhuber, J., 1997).
+\begin{equation}
+    f_t = \sigma( W_f \cdot [ h_{t-1} , x_t ] + b_f ) 
+\end{equation}
+% Input gate
+Input gate: 
+\\[1ex]
+The input gate controls how much new information from the current time step is allowed to enter the cell (Hochreiter, S., \& Schmidhuber, J., 1997).
+For the $\widetilde{C}_t$, the purpose is to suggest updates for the cell state.
+\begin{equation}
+    i_t = \sigma( W_i \cdot [ h_{t-1} , x_t ] + b_c ) 
+\end{equation}
+\begin{equation}
+    \widetilde{C}_t = tanh( W_c \cdot [ h_{t-1} , x_t ] + b_c )
+\end{equation}
+% Cell state update
+Cell State Update:
+\\[1ex]
+The forget gate will drop the meaningless information and add some potential information.
+\begin{equation}
+    C_t = f_t * C_{t-1} + i_t * \widetilde{C_t}
+\end{equation}
+% Output gate
+Output gate:
+\\[1ex]
+The output gate is able to control how much or what information from the cell state should be passed to the next layer or used in predictions. 
+\begin{equation}
+    o_t = \sigma( W_o \cdot [ h_{t-1} , x_t ] + b_o )
+\end{equation}
+% Hidden state update
+Hidden State Update:
+\\[1ex]
+The hidden state is influenced by output value and current cell state. 
+\begin{equation}
+    h_t = o_t * tanh(C_t)
+\end{equation}
+\\[1ex]
+n = number of features in the input vector $x_t$.\\
+m = number of units in LSTM.\\
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $x_t$               & Input vector at time t                        & $n \times 1$\\
+    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
+    $C_t$               & Cell state at time t                          & $m \times 1$\\
+    $f_t$               & Forget gate output at time t                  & $m \times 1$\\
+    $i_t$               & Input gate output at time t                   & $m \times 1$\\
+    $o_t$               & Output gate output at time t                  & $m \times 1$\\
+    $\widetilde{C_t}$   & Candidate memory cell at time t               & $m \times 1$\\
+    $W_f$               & Weight matrix for the forget gate             & $m \times (m+n)$\\
+    $W_i$               & Weight matrix for the input gate              & $m \times (m+n)$\\
+    $W_C$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
+    $W_o$               & Weight matrix for the output gate             & $m \times (m+n)$\\
+    $b_f$               & Bias vector for the forget gate               & $m \times 1$\\
+    $b_i$               & Bias vector for the input gate                & $m \times 1$\\
+    $b_C$               & Bias vector for the candidate memory cell     & $m \times 1$\\
+    $b_o$               & Bias vector for the output gate               & $m \times 1$\\
+    \hline
+    \caption{Unfolded RNN}
+\end{longtable}
+
+Number of parameters:\\
+\begin{enumerate}
+    % 1. Weights matrix for the input
+    \item Weights matrix for the input
+    \begin{itemize}
+        \item Forget gate: $n \times m$
+        \item Input gate: $n \times m$
+        \item Cell gate: $n \times m$
+        \item Output gate: $n \times m$
+    \end{itemize}
+
+    % 2. Weight matrix for the hidden state
+    \item Weight matrix for the hidden state
+    \begin{itemize}
+        \item Hidden state for forget gate: $m \times m$
+        \item Hidden state for input gate: $m \times m$
+        \item Hidden state for cell gate: $m \times m$
+        \item Hidden state for output gate: $m \times m$
+    \end{itemize}
+
+    % 3. Bias term
+    \item Bias term
+    \begin{itemize}
+        \item Bias for forget gate: $1 \times m$
+        \item Bias for input gate: $1 \times m$
+        \item Bias for cell gate: $1 \times m$
+        \item Bias for output gate: $1 \times m$
+    \end{itemize}
+\end{enumerate}
+Total parameters: $4 \times ( n + m + 1 ) \times m$
+
+\newpage
+% Gated Recurrent Unit (GRU)
+\textbf{Gated Recurrent Unit (GRU)}
+\\[1ex]
+The gated recurrent unit (GRU) was proposed by Cho et al.(2014) to make each recurrent unit to adaptively capture dependencies of different time scales.
+The GRU has 2 gate, update gate and reset gate.
+\\[2ex]
+Update gate:
+\\[1ex]
+The update gate determines how much of the past information should be retained in the current hidden state. 
+\begin{equation}
+    z_t = \sigma(W_z \cdot [ h_{t-1} , x_t ] + b_z )
+\end{equation}
+% Reset gate
+Reset gate:
+\\[1ex]
+The reset gate is similar with the update gate, but the candidate hidden state is influenced by the reset gate. 
+\begin{equation}
+    r_t = \sigma( W_r \cdot [ h_{t-1} , x_t ] + b_r )
+\end{equation}
+Candidate hidden state:
+\\[1ex]
+The candidate hidden state combined with previous hidden state and current input to form the potential new information that can be added to the current hidden state.
+\begin{equation}
+    \widetilde{h_t} = tanh ( W_h \cdot [ r_t * h_{t-1} , x_t ] + b_h )
+\end{equation}
+Final hidden state
+\\[1ex]
+The final hidden state of the GRU at time t is a linear interpolation between the previous final hidden state and the candidate hidden state. 
+\begin{equation}
+    h_t = ( 1 - z_t ) * h_{t-1} + z_t * \widetilde{ h_t }
+\end{equation}
 % ######################## Literature Review ######################################

### 2025-01-22 paper/main.tex
mac

@@
 \end{equation}
+
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $x_t$               & Input vector at time t                        & $n \times 1$\\
+    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
+    $r_t$               & Reset gate output at time t                   & $m \times 1$\\
+    $z_t$               & Update gate output at time t                  & $m \times 1$\\
+    $W_z$               & Weight matrix for the update gate             & $m \times (m+n)$\\
+    $W_r$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
+    $W_h$               & Weight matrix for the output gate             & $m \times (m+n)$\\
+    $b_z$               & Bias vector for the input gate                & $m \times 1$\\
+    $b_r$               & Bias vector for the candidate memory cell     & $m \times 1$\\
+    $b_h$               & Bias vector for the output gate               & $m \times 1$\\
+    \hline
+\end{longtable}
+
+\newpage
+% Hidden markov model
+\textbf{Hidden Markov Model}
+\\[1ex]
+For the NLP tasks, there may have some unobservable data, for example the implicit sentiment, user intent, emotional state.
+Rabiner (1989, 1993) proposes the hidden Markov model (HMM), the assumption is the existence of the latent process follows a Markov chain from which observations X are generated.
+In other word, there would exists an unobserved state sequence $Z = \{z_1, z_2 \cdots , z_T\}$ in observed sequence $X = \{x_1, x_2, \cdots, x_T\}$.
+Where the hidden states, $z_t$ belonging to state-space $Q = \{q_1, q_2, \cdots, q_M \}$ follow a Markov chain goverened by:
+\begin{itemize}
+    \item A state-transition probability matrix $A = [a_{ij}] \in \mathbb{R}^{M \times M}$ where $a_{ij} = p(z_{t+1} = q_j | z_t = q_i)$ 
+\end{itemize}
+\begin{itemize}
+    \item Initial state matrix $\pi = [\pi_i] \in \mathbb{R}_{1 \times M}$ with $\pi_i = p(z_1 = q_i)$
+\end{itemize}
+% \\[1ex]
+
 % ######################## Literature Review ######################################

### 2025-01-26 paper/main.tex
finish document

@@
 \usepackage{algorithmic}
+\usepackage[style=apa, sorting=nyt, backend=biber]{biblatex} % Use biber as the backend
+\addbibresource{references.bib}
+
 \setlength{\parindent}{0em}
@@
 Unfortunately, early RNNs had limitation in training of networks over long sequence.
-vanishing and exploding gradient problems significantly affect the training process of RNN (Bengio, Simard, \& Frasconi, 1994).
+vanishing and exploding gradient problems significantly affect the training process of RNN \parencite{bengio1994learning}.
 Eliminating many practical applications of RNNs.
-After that, Hochreiter and Schmidhuber (1997) introduced Long Short-Term Memory (LSTM) networks and are responsible for the breakthrough in how to solve these challenges.
+After that, \parencite{hochreiter1997lstm} introduced Long Short-Term Memory (LSTM) networks and are responsible for the breakthrough in how to solve these challenges.
 Specificized gating mechanisms were introduced in LSTMs to regulate the flow of the information, minimize the vanishing gradient problem and learn the long-term dependencies.
 This advanced made RNNs much more performant on tasks like a language modeling, machine translation and speech recognition tasks.\\[2ex]
-Further improvements were achieved with Gated Recurrent Units (GRUs) by Cho et al.
-(2014) which diminished the LSTM architecture's complexity, but still provided the same performance.
+Further improvements were achieved with Gated Recurrent Units (GRUs) by \parencite{cho2014properties} which diminished the LSTM architecture's complexity, but still provided the same performance.
 GRUs performed comparably but used fewer parameters, making it computationally and more tractably trainable.
@@
 BPTT is one of the most important algorithms used for training RNNs.
-Dating back to the original effort to expand the typical backpropagation algorithm, BPTT has been formulated to handle the difficulties of temporal sequences that are inherent in sequential data (Werbos, 1990).
-This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable (Rumelhart, Hinton, \& Williams, 1986).
+Dating back to the original effort to expand the typical backpropagation algorithm, BPTT has been formulated to handle the difficulties of temporal sequences that are inherent in sequential data \parencite{werbos1990bptt}.
+This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable \parencite{rumelhart1986backpropagation}.
 \\[2ex]
@@
 In the forward pass, the RNN, like other artificial neuronal network, applies operation over the data input in sequence, bringing changes in its own state variables at every time step, depending on the input and the previous state of its general working state or hidden state.
-This sequential processing produces outputs and stores the internal states of the network in any period (Werbos, 1990).
+This sequential processing produces outputs and stores the internal states of the network in any period \parencite{werbos1990bptt}.
 
@@
 \newpage
-\textbf{Pseudocode of BPTT} (Wikipedia, 2023)
+\textbf{Pseudocode of BPTT} \parencite{wikipedia2023bptt}
 \begin{algorithm}[H]
@@
 When training the RNN, BPTT was used to update the weight matrix.
-As the number of time steps increase, the problem of gradient instability of often encountered, and this problem is gradient vanishing and gradient explored (Bengio et al.
-1994).
+As the number of time steps increase, the problem of gradient instability of often encountered, and this problem is gradient vanishing and gradient explored \parencite{bengio1994learning}.
 \\[2ex]
@@
 \\[2ex]
-Long short-term memory proposed by Hochreiter \& Schmidhuber (1997).
+Long short-term memory proposed by \parencite{hochreiter1997lstm}.
 LSTM is designed for handling long time step problems.
@@
 The forget gate is a component of the LSTM, designed to manage the flow of information within the cell state.
-The function of forget gate is to determine which information should be retained in memory cell (Hochreiter, S., \& Schmidhuber, J., 1997).
+The function of forget gate is to determine which information should be retained in memory cell \parencite{hochreiter1997lstm}.
 \begin{equation}
@@
 \\[1ex]
-The input gate controls how much new information from the current time step is allowed to enter the cell (Hochreiter, S., \& Schmidhuber, J., 1997).
+The input gate controls how much new information from the current time step is allowed to enter the cell \parencite{hochreiter1997lstm}.
 For the $\widetilde{C}_t$, the purpose is to suggest updates for the cell state.
@@
 \\[1ex]
-The gated recurrent unit (GRU) was proposed by Cho et al.(2014) to make each recurrent unit to adaptively capture dependencies of different time scales.
+The gated recurrent unit (GRU) was proposed by \parencite{cho2014properties} to make each recurrent unit to adaptively capture dependencies of different time scales.
 The GRU has 2 gate, update gate and reset gate.
@@
 For the NLP tasks, there may have some unobservable data, for example the implicit sentiment, user intent, emotional state.
-Rabiner (1989, 1993) proposes the hidden Markov model (HMM), the assumption is the existence of the latent process follows a Markov chain from which observations X are generated.
+\parencite{rabiner1993fundamentals} proposes the hidden Markov model (HMM), the assumption is the existence of the latent process follows a Markov chain from which observations X are generated.
 In other word, there would exists an unobserved state sequence $Z = \{z_1, z_2 \cdots , z_T\}$ in observed sequence $X = \{x_1, x_2, \cdots, x_T\}$.
@@
 \end{itemize}
-% \\[1ex]
+Furthermore, for the hidden state $z_t$, correspoinding to the observe data $x_t$ is release by emission process $B = [b_j(x)]$ where $b_j(x) = p(x|z=q_i)$.
+We can assume $b_j(x)$ is follows the Gaussian mixture model (GMM).
+\begin{equation}
+    p(x_t|z=q_j) = \sum_{l=1}^{k}c_{jl}\mathcal{N}(x_t|\mu_{jl}, \Sigma_{jl})
+\end{equation}
+where $\sum_{l=1}^{k}c_{jl} = 1$, $\forall j = \{1,\cdots, M\}$, k is the number of Gaussian mixture components and $\mathcal{N}(x_t|\mu_{jl}, \Sigma_{jl})$ denotes a Gaussian probability density with mean $\mu_{jl}$ and covariance $\Sigma_{jl}$ for state $j$ and mixture component $l$.
+The number of hidden states ($M$) and mixture component ($k$) are the two hyperparameters of the model which have to be provided apriori.
+\\[1ex]
+Therefore, the joint probability probability density function of the observation $X$ can be expressed as:
+\begin{equation}
+    p(X) = p(z_1) \prod^{t=1}_{T-1} p(z_{t+1}|z_{t}) \prod_{t=1}^{T} p(x_t|z_t)
+\end{equation}
+The optimal parameters $[A, B, \pi]$, which maximize the likelihood of the observation sequence $X$ (Equation 39), are determined using the Baum-Welch algorithm, an expectation-maximization method \parencite{rabiner1993fundamentals}.
+Additionally, the probability of the system in a specific hidden state $z_t$ corresponding to the observation $x_t$ is calculated using the Viterbi algorithm.
+\newpage
 
+\textbf{Word representation}
+\\[1ex]
+In RNN, the due to the architecture, the model can only handle vector, it is important that convert the word into machine readable format.
+By translating words into vectors, these representations capture semantic meanings, relationships and contexts.
+\parencite{mikolov2013efficient} proposed 2 model architectures for leaning distributed representations of words.
+\\[1ex]
+\textbf{Continuous Bag of Words model}
+CBOW operates by taking context words around a target word, it is flexible to adjust the window size which mean that we can control how many words around our target word.
+Hence aggregating their embeddings and passing to hidden layer to produce a probability distribution.
+Capture the semantic meanings and the relationship more effectively.
+\begin{equation}
+    Q = N \times D + D \times log_2(V)
+\end{equation}
+\\[1ex]
+\textbf{Continuous Skip gram model}
+\\[1ex]
+The continuous skip gram model operates in opposite direction compare with CBOW.
+The model takes the surrounding word to predict the target word.
+In other word, treat the current word as an input to a log-linear classifier with continuous projection layer, and predict the probability distribution of surrounding words.
+\begin{equation}
+    Q = C \times (D + D \times log_2(V))
+\end{equation}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/word_representation.png} % Replace "example-image" with your image file
+\end{figure}
+\newpage
 % ######################## Literature Review ######################################
+
+% ######################## Reference ######################################
+\printbibliography
+% ######################## Reference ######################################
 

### 2025-02-01 paper/main.tex
testing
paper

@@
 \end{longtable}
+\newpage
+
+% Deep recurrent neural networks (DRNNs)
+\textbf{Deep Recurrent Neural Networks (DRNNs)}
+\\[1ex]
+Deep architecture networks with multiple layers that can hierarchically learn complex representations \parencite{bengio2009learning}.
+By extending of this concept, stacking recurrent layers to form Deep Recurrent Neural Networks aligns with this principle.
+A number of research papers have been prove that the performance of DRNNs is out-performance then conventional RNNs.
+\parencite{le2010deep, delalleau2011shallow, pascanu2013on}
+\\[1ex]
+The architecture of DRNNs are similar with conventional RNNs.
+We simply stack the recurrent layers vertically.
+For the first layer, this layer receives the input and combines it with its previous hidden state $h_{t-1}$ (Equation 38).
+The second layer receive the hidden state of the first layer and treat as the input of the layer 2 (Equation 39).
+We can extend this concept to $L$ layers (Equation 40).
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../Pic/drnn.png} % Replace "example-image" with your image file
+\end{figure}
+\\[1ex]
+\begin{equation}
+    h^{(1)}_t = f(W^{(1)}_{xh} x_{t} + W^{(1)}_{hh} h^{(1)}_{t-1} + b^{(1)}_h)
+\end{equation}
+\begin{equation}
+    h^{(2)}_t = f(W^{(2)}_{xh} h^{(1)}_{t} + W^{(2)}_{hh} h^{(2)}_{t-1} + b^{(2)}_h)
+\end{equation}
+\begin{equation}
+    h^{(L)}_t = f(W^{(L)}_{xh} h^{(L-1)}_{t} + W^{(L)}_{hh} h^{(L)}_{t-1} + b^{(L)}_h)
+\end{equation}
+\begin{equation}
+    o_t = W_{hy}h^{(L)}_{t} + b_y
+\end{equation}
+\begin{equation}
+    y_t = g(o_t)
+\end{equation}
+Where $x_t$ is the input at time $t$.
+$h^{(l)}_{t}$ is the hidden state for the $l$ layer at time t.
+$W^{(l)}_{xh}, W^{(l)}_{hh}$ are the weight matrices for the input to hidden and hidden to hidden connections in layer $l$, respectively.
+$W_{hy}$ is weight matrix for the output layer.
+$b^{(l)}_h$ is the bias vector for the $l$ layer, $b_y$ is the bias vector for output layer.
+$g(\cdot)$ and $f(\cdot)$ are an activation function.
 
@@
 
+% ######################## My work ######################################
+In this paper, 
+% ######################## My work ######################################
+
 % ######################## Reference ######################################

### 2025-02-08 paper2/main.tex
finalize paper

--- 2025-01-22 paper/main.tex
@@
 \newpage
-
-% ######################## Project goals ######################################
-\section{Project goals}
-The goal of this paper is to analyze, implement, and compare the performance of RNN, LSTM, GRU and own model in selected NLP tasks.
-This paper aims to propose an enhanced RNN-based model for NLP tasks and make the recommendation for future development.
-% ######################## Project goals ######################################
-
-% ######################## Objectives ######################################
-\section{Objectives}
-\begin{enumerate}
-    % 1. Literature Review
-    \item Literature Review
-    \begin{itemize}
-        \item Conduct a comprehensive review of existing paper on RNN, LSTM, and GRU, focusing on their architecture.
-        \item Review the NLP task.
-    \end{itemize}
-
-    % 2. Model implementation
-    \item Model implementation
-    \begin{itemize}
-        \item Implement RNN, LSTM, GRU and own model by PyTorch for chosen NLP tasks.
-    \end{itemize}
-
-    % 3. Performance Comparison
-    \item Performance Comparison
-    \begin{itemize}
-        \item Evaluate the performance by using appropriate metrics (e.g., accuracy, precision, recall, F1 score, BLEU, ROUGE).
-    \end{itemize}
-
-    % 4. Propose an advanced model
-    \item Propose an advanced model
-    \begin{itemize}
-        \item Develop an advanced model, aiming to enhance the performance on selected NLP tasks. 
-    \end{itemize}
-\end{enumerate}
-% ######################## Objectives ######################################
-
-% ######################## Expected outcomes ######################################
-\section{Expected outcomes}
-\begin{enumerate}
-    % 1. Model performance metrics
-    \item Model performance metrics
-    \begin{itemize}
-        \item A detailed comparison of performance metrics across RNN, LSTM, GRU and own model. 
-    \end{itemize}
-
-    % 2. Practical insights
-    \item Practical insights
-    \begin{itemize}
-        \item Practical recommendations for future work in the area of RNN applications in NLP, based on the findings of this paper. 
-    \end{itemize}
-\end{enumerate}
-% ######################## Expected outcomes ######################################
-
-% ######################## Research plan ######################################
-\section{Research plan}
-\begin{enumerate}
-    % 1. Literature review
-    \item Literature review
-    \begin{itemize}
-        \item Review existing work on RNN, LSTM, GRU and their more sophisticated variants. 
-    \end{itemize}
-
-    % 2. Model development
-    \item Model development
-    \begin{itemize}
-        \item Select NLP tasks. (e.g., text classification, summarization).
-        \item Implement baseline models (RNN, LSTM, GRU) using PyTroch.
-    \end{itemize}
-
-    % 3. Performance Evaluation
-    \item Performance Evaluation
-    \begin{itemize}
-        \item Evaluate and compare the performance of the implemented model using appropriate metrics. 
-    \end{itemize}
-
-    % 4. Model Enhancement
-    \item Model Enhancement
-    \begin{itemize}
-        \item Design and implement an advanced model based on findings from the previous phases.
-    \end{itemize}
-
-    % 5. Final analysis and reporting
-    \item Final analysis and reporting
-    \begin{itemize}
-        \item Analyze the results of the advanced model against baseline models. 
-    \end{itemize}
-\end{enumerate}
-% ######################## Research plan ######################################
-
-% ######################## Experimental design ######################################
-\section{Experimental design}
-\begin{enumerate}
-    % 1. Dataset selection
-    \item Dataset selection
-    \begin{itemize}
-        \item Select suitable NLP dataset.
-    \end{itemize}
-
-    % 2. Model configuration
-    \item Model configuration
-    \begin{itemize}
-        \item Define architecture specifications for each model, including number of layers, number of hidden units, activation function, dropout rates.
-    \end{itemize}
-\end{enumerate}
-% ######################## Experimental design ######################################
-
-\newpage
-% ######################## Introduction ######################################
+% ######################## Introduction ####################################
 \section{Introduction}
@@
 Especially NLP, has emerged as a pivotal field in artificial intelligence, enable machines to understand, interpret and generate in human readable format.
-Siri, Alexa and bixby have shown the possibility.
-Everyone can communicate with those machines and they with make the reasonable response to user.\\[2ex]
-Recurrent Neural Networks (RNNs) have been a foundational architecture in this domain, the architecture of RNNs is design for sequential data.
-It able to retain the information through hidden states.
+Some famous Artificial Intellience assistence for example, Siri, Alexa and Bixby have shown the possibility.
+Everyone can communicate with those machines, which make the reasonable response back to user.\\[1ex]
+Recurrent Neural Networks (RNNs) have been a foundational architecture in this domain, Unlike the traditional Artificial Neural Network, RNNs do not treat each input independently, RNNs handle each input by considering the information from previous inputs.
+Conceptually this architecture able to retain the information.
+Thus, this architecture is suitable for handling sequential data.
 Unfortunately, early RNNs had limitation in training of networks over long sequence.
-vanishing and exploding gradient problems significantly affect the training process of RNN (Bengio, Simard, \& Frasconi, 1994).
+vanishing and exploding gradient problems significantly affect the training process of RNN \parencite{bengio1994learning}.
 Eliminating many practical applications of RNNs.
-After that, Hochreiter and Schmidhuber (1997) introduced Long Short-Term Memory (LSTM) networks and are responsible for the breakthrough in how to solve these challenges.
+After that, \parencite{hochreiter1997lstm} introduced Long Short-Term Memory (LSTM) networks and are responsible for the breakthrough in how to solve these challenges.
 Specificized gating mechanisms were introduced in LSTMs to regulate the flow of the information, minimize the vanishing gradient problem and learn the long-term dependencies.
-This advanced made RNNs much more performant on tasks like a language modeling, machine translation and speech recognition tasks.\\[2ex]
-Further improvements were achieved with Gated Recurrent Units (GRUs) by Cho et al.
-(2014) which diminished the LSTM architecture's complexity, but still provided the same performance.
-GRUs performed comparably but used fewer parameters, making it computationally and more tractably trainable.
-% ######################## Introduction ######################################
+This advanced made RNNs much more performant on tasks like a language modeling, machine translation and speech recognition tasks.\\[1ex]
+Further improvements were achieved with Gated Recurrent Units (GRUs) by \parencite{cho2014properties} which diminished the LSTM architecture's complexity, but still provided the same performance.
+GRUs performed comparably but used fewer parameters, making it computationally and more tractably trainable.\\[1ex]
+Since the craze of AI has been revived by generative AI, natural language processing to time series prediction and speech recognition have once again aroused people's interest in RNN.
+This report aims to:
+\begin{itemize}
+    \item Explore the theoretical foundations of recurrent deep learning models.
+    \item Investigate their diverse applications in solving sequential data tasks.
+    \item Analyze their performance, strengths, and inherent limitations.
+\end{itemize}
+% ######################## Introduction ####################################
+\newpage
+% ######################## Background and Literature Review #########################
+% Background and Literature Review 
+\section{Background and Literature Review}
+% Evolution of Deep Learning and Recurrent Neural Networks 
+\subsection{Evolution of Deep Learning and Recurrent Neural Networks}
+In the past few decades, thank to the rapidly development of technology, the computing resource has a incredible increase.
+Thus, substantially deep learning architecture have improved, from simple architectures, which only able to capture simple information from data to sophisticated models that are able to learn complex, abstract representations.
+This was before the early neural networks like perceptrons and multilayer perceptrons (MLPs) laid the footwork of neural computation that first came in the picture, but were burdened by the lack of ability to model sequential dependencies.
+This however imposed a limit on the feed forward paradigm, which prompted the development of recurrent neural networks (RNNs) that extend the old stalactite of feed forward paradigm with cyclic connections.
+Through these connection, RNNs are capable to keep a hidden state that represents information over time steps thereby effectively capture temporal dynamics.
+RNNs have been a decisive step in the evolution of deep learning, as they are able to do tasks that require memory of previous events, including problems of natural language processing and time series modeling.
+Despite that, early RNNs models suffered from serious problems for example,  vanishing gradients and exploding gradients, which prevented these RNN models from learning long ranged dependencies.
+This stimulated the building of more refined architectures intended to side step these obstacles.
 
 \newpage
-% ######################## Literature Review ######################################
-\section{Literature Review}
-\textbf{Backpropagation Through Time}
-\\[1ex]
+
+% Literature Review
+\subsection{Literature Review}
+% Backpropagation Through Time
+\subsubsection{Backpropagation Through Time}
 BPTT is one of the most important algorithms used for training RNNs.
-Dating back to the original effort to expand the typical backpropagation algorithm, BPTT has been formulated to handle the difficulties of temporal sequences that are inherent in sequential data (Werbos, 1990).
-This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable (Rumelhart, Hinton, \& Williams, 1986).
+Dating back to the original effort to expand the typical backpropagation algorithm, BPTT has been formulated to handle the difficulties of temporal sequences that are inherent in sequential data \parencite{werbos1990bptt}.
+This algorithm allows RNNs in learning sequence dependent data by unfold the network over time steps and then updating weights matrix through the gradient of loss function with respect to the variable \parencite{rumelhart1986backpropagation}.
 \\[2ex]
@@
     \caption{Unfolded RNN}
-    % \label{fig:sdfsdf}
 \end{figure}
@@
 \end{longtable}
-
 % Forward pass
@@
 \end{equation}
-% \begin{equation}
-%     \hat{y}_t = f(\hat{o}_t)
-% \end{equation}
 \newline  % Computing the loss function
@@
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial{W}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial W}
+    \dfrac{\partial\mathcal{L}_2}{\partial W} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial W}
 \end{equation}
+By mathematic induction 
 \begin{equation}
-    \dfrac{\partial L_i}{\partial W} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial W}
+    \dfrac{\partial\mathcal{L}_t}{\partial W} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial W})
 \end{equation}
+Where 
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial W} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial W}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial W}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
 \end{equation}
@@
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial{U}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial U}
+    \dfrac{\partial\mathcal{L}_2}{\partial U} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial U}
 \end{equation}
+By mathematic induction 
 \begin{equation}
-    \dfrac{\partial L_i}{\partial U} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial U}
+    \dfrac{\partial\mathcal{L}_t}{\partial U} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial U})
 \end{equation}
+Where 
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial U} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial U}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial U}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
 \end{equation}
@@
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial{V}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial V}
+    \dfrac{\partial\mathcal{L}_2}{\partial V} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial V}
 \end{equation}
+By mathematic induction 
 \begin{equation}
-    \dfrac{\partial L_i}{\partial V} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial V}
+    \dfrac{\partial\mathcal{L}_t}{\partial V} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial V})
 \end{equation}
+Where 
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial V} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial V}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial V}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
 \end{equation}
@@
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial{b_h}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial b_h}
+    \dfrac{\partial\mathcal{L}_2}{\partial b_b} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_h}
 \end{equation}
+By mathematic induction 
 \begin{equation}
-    \dfrac{\partial L_i}{\partial b_h} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_h}
+    \dfrac{\partial\mathcal{L}_t}{\partial b_h} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_h})
 \end{equation}
+Where 
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial b_h} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial b_h}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial b_h}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
 \end{equation}
@@
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial{b_y}} = \sum_{i=1}^{2}\dfrac{\partial L_i}{\partial b_y}
+    \dfrac{\partial\mathcal{L}_2}{\partial b_y} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_y}
 \end{equation}
+By mathematic induction 
 \begin{equation}
-    \dfrac{\partial L_i}{\partial b_y} = \dfrac{\partial L_i}{\partial \hat{y}_i} \cdot \dfrac{\partial \hat{y}_i}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_y}
+    \dfrac{\partial\mathcal{L}_t}{\partial b_y} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_y})
 \end{equation}
+Where 
 \begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial b_y} = \dfrac{\partial\mathcal{L}_1}{\partial\hat{y}_1}\cdot\dfrac{\partial\hat{y}_1}{\partial\hat{h}_1}\cdot\dfrac{\partial h_1}{\partial b_y}+\dfrac{\mathcal{L}_2}{\partial\hat{y}_2}\cdot\dfrac{\partial\hat{y}_2}{\partial\hat{h}_2}\cdot\dfrac{\partial h_2}{\partial h_1}\cdot\dfrac{\partial h_1}{\partial b_y}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
 \end{equation}
@@
 
-\newpage
-\textbf{Pseudocode of BPTT} (Wikipedia, 2023)
+\textbf{Pseudocode of BPTT} \parencite{wikipedia2023bptt}
 \begin{algorithm}[H]
@@
 % Actication function
-\textbf{Activation function}
-\\[1ex]
+\subsubsection{Activation Function}
 Activation functions, particularly the sigmoid function, are fundamental components of recurrent neural networks (RNNs).
@@
 Differentiability is crucial for the backpropagation through time (BPTT) algorithm, enabling the application of the chain rule during training.
+\\[1ex]
+% Sigmoid Activation Function 
+\textbf{Sigmoid activation function}
+\\[1ex]
+The main role of the sigmoid activation function is to normalize candidate values and convert the cell state to a hidden state when performing cell state updates.
+It limits the output between [0,1] because it has a smooth gradient, which is important for discovering long-range dependencies.\\[1ex]
 \begin{equation}
@@
 This leads to the model struggling to learn long time dependencies.
-\\[2ex]
+\\[1ex]
 % Exploding Gradients
@@
 This will lead to exploding gradients.
+\\[1ex]
+We have introduce the backpropagation through time.
+This is the method to update the parameters in RNNs.
+When calculating, for example, the partial derivative of loss function with respect to W.
+Assume the time $t$ go to infinity large.
+We will get this term.
+$ \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}} $, and it will lead to exponential problem.
+if $ \dfrac{\partial h_j}{\partial h_{j-1}} > 1 $.
+Then the product of all term will increase exponentially, then exploding gradients occur.
+On the contrary, if $\dfrac{\partial h_j}{\partial h_{j-1}} < 1$.
+Then the result will decrease exponentially, then vanishing gradients occur.
 
-\newpage
-% Long short-term memory (LSTM)
-\textbf{Long short-term memory (LSTM)}
-\\[2ex]
-Long short-term memory proposed by Hochreiter \& Schmidhuber (1997).
-LSTM is designed for handling long time step problems.
-The architecture of LSTM can prevent vanishing gradient and exploring gradient.
-The main difference between LSTM and RNN is the number of gates.
-LSTM introduced input, forget and output gates.
-This allows LSTM to manage the flow of information more effectively, retaining important information over longer sequences.
-\\[2ex]
-Architecture
-\\[1ex]
-LSTMs introduce a memory cell that can maintain information over long time steps the cell is controlled by three gates, input gate, output gate, and forget gate.
-Each cell of LSTMs inside has 3 sigmoid and 1 tanh layer.
-Below graph unfolds the LSTM hence we can analyze different gates.
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=1\textwidth]{../Pic/lstm1.png} % Replace "example-image" with your image file
-\end{figure}
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=1\textwidth]{../Pic/lstm2.png} % Replace "example-image" with your image file
-\end{figure}
-\\
-% Forget gate
-Forget gate:
-\\[1ex]
-The forget gate is a component of the LSTM, designed to manage the flow of information within the cell state.
-The function of forget gate is to determine which information should be retained in memory cell (Hochreiter, S., \& Schmidhuber, J., 1997).
-\begin{equation}
-    f_t = \sigma( W_f \cdot [ h_{t-1} , x_t ] + b_f ) 
-\end{equation}
-% Input gate
-Input gate: 
-\\[1ex]
-The input gate controls how much new information from the current time step is allowed to enter the cell (Hochreiter, S., \& Schmidhuber, J., 1997).
-For the $\widetilde{C}_t$, the purpose is to suggest updates for the cell state.
-\begin{equation}
-    i_t = \sigma( W_i \cdot [ h_{t-1} , x_t ] + b_c ) 
-\end{equation}
-\begin{equation}
-    \widetilde{C}_t = tanh( W_c \cdot [ h_{t-1} , x_t ] + b_c )
-\end{equation}
-% Cell state update
-Cell State Update:
-\\[1ex]
-The forget gate will drop the meaningless information and add some potential information.
-\begin{equation}
-    C_t = f_t * C_{t-1} + i_t * \widetilde{C_t}
-\end{equation}
-% Output gate
-Output gate:
-\\[1ex]
-The output gate is able to control how much or what information from the cell state should be passed to the next layer or used in predictions. 
-\begin{equation}
-    o_t = \sigma( W_o \cdot [ h_{t-1} , x_t ] + b_o )
-\end{equation}
-% Hidden state update
-Hidden State Update:
-\\[1ex]
-The hidden state is influenced by output value and current cell state. 
-\begin{equation}
-    h_t = o_t * tanh(C_t)
-\end{equation}
-\\[1ex]
-n = number of features in the input vector $x_t$.\\
-m = number of units in LSTM.\\
-\begin{longtable}{|c|c|c|}
-    \hline
-    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
-    \hline
-    $x_t$               & Input vector at time t                        & $n \times 1$\\
-    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
-    $C_t$               & Cell state at time t                          & $m \times 1$\\
-    $f_t$               & Forget gate output at time t                  & $m \times 1$\\
-    $i_t$               & Input gate output at time t                   & $m \times 1$\\
-    $o_t$               & Output gate output at time t                  & $m \times 1$\\
-    $\widetilde{C_t}$   & Candidate memory cell at time t               & $m \times 1$\\
-    $W_f$               & Weight matrix for the forget gate             & $m \times (m+n)$\\
-    $W_i$               & Weight matrix for the input gate              & $m \times (m+n)$\\
-    $W_C$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
-    $W_o$               & Weight matrix for the output gate             & $m \times (m+n)$\\
-    $b_f$               & Bias vector for the forget gate               & $m \times 1$\\
-    $b_i$               & Bias vector for the input gate                & $m \times 1$\\
-    $b_C$               & Bias vector for the candidate memory cell     & $m \times 1$\\
-    $b_o$               & Bias vector for the output gate               & $m \times 1$\\
-    \hline
-    \caption{Unfolded RNN}
-\end{longtable}
-
-Number of parameters:\\
-\begin{enumerate}
-    % 1. Weights matrix for the input
-    \item Weights matrix for the input
-    \begin{itemize}
-        \item Forget gate: $n \times m$
-        \item Input gate: $n \times m$
-        \item Cell gate: $n \times m$
-        \item Output gate: $n \times m$
-    \end{itemize}
-
-    % 2. Weight matrix for the hidden state
-    \item Weight matrix for the hidden state
-    \begin{itemize}
-        \item Hidden state for forget gate: $m \times m$
-        \item Hidden state for input gate: $m \times m$
-        \item Hidden state for cell gate: $m \times m$
-        \item Hidden state for output gate: $m \times m$
-    \end{itemize}
-
-    % 3. Bias term
-    \item Bias term
-    \begin{itemize}
-        \item Bias for forget gate: $1 \times m$
-        \item Bias for input gate: $1 \times m$
-        \item Bias for cell gate: $1 \times m$
-        \item Bias for output gate: $1 \times m$
-    \end{itemize}
-\end{enumerate}
-Total parameters: $4 \times ( n + m + 1 ) \times m$
-
-\newpage
-% Gated Recurrent Unit (GRU)
-\textbf{Gated Recurrent Unit (GRU)}
-\\[1ex]
-The gated recurrent unit (GRU) was proposed by Cho et al.(2014) to make each recurrent unit to adaptively capture dependencies of different time scales.
-The GRU has 2 gate, update gate and reset gate.
-\\[2ex]
-Update gate:
-\\[1ex]
-The update gate determines how much of the past information should be retained in the current hidden state. 
-\begin{equation}
-    z_t = \sigma(W_z \cdot [ h_{t-1} , x_t ] + b_z )
-\end{equation}
-% Reset gate
-Reset gate:
-\\[1ex]
-The reset gate is similar with the update gate, but the candidate hidden state is influenced by the reset gate. 
-\begin{equation}
-    r_t = \sigma( W_r \cdot [ h_{t-1} , x_t ] + b_r )
-\end{equation}
-Candidate hidden state:
-\\[1ex]
-The candidate hidden state combined with previous hidden state and current input to form the potential new information that can be added to the current hidden state.
-\begin{equation}
-    \widetilde{h_t} = tanh ( W_h \cdot [ r_t * h_{t-1} , x_t ] + b_h )
-\end{equation}
-Final hidden state
-\\[1ex]
-The final hidden state of the GRU at time t is a linear interpolation between the previous final hidden state and the candidate hidden state. 
-\begin{equation}
-    h_t = ( 1 - z_t ) * h_{t-1} + z_t * \widetilde{ h_t }
-\end{equation}
-
-\begin{longtable}{|c|c|c|}
-    \hline
-    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
-    \hline
-    $x_t$               & Input vector at time t                        & $n \times 1$\\
-    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
-    $r_t$               & Reset gate output at time t                   & $m \times 1$\\
-    $z_t$               & Update gate output at time t                  & $m \times 1$\\
-    $W_z$               & Weight matrix for the update gate             & $m \times (m+n)$\\
-    $W_r$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
-    $W_h$               & Weight matrix for the output gate             & $m \times (m+n)$\\
-    $b_z$               & Bias vector for the input gate                & $m \times 1$\\
-    $b_r$               & Bias vector for the candidate memory cell     & $m \times 1$\\
-    $b_h$               & Bias vector for the output gate               & $m \times 1$\\
-    \hline
-\end{longtable}
-
-\newpage
-% Hidden markov model
-\textbf{Hidden Markov Model}
-\\[1ex]
-For the NLP tasks, there may have some unobservable data, for example the implicit sentiment, user intent, emotional state.
-Rabiner (1989, 1993) proposes the hidden Markov model (HMM), the assumption is the existence of the latent process follows a Markov chain from which observations X are generated.
-In other word, there would exists an unobserved state sequence $Z = \{z_1, z_2 \cdots , z_T\}$ in observed sequence $X = \{x_1, x_2, \cdots, x_T\}$.
-Where the hidden states, $z_t$ belonging to state-space $Q = \{q_1, q_2, \cdots, q_M \}$ follow a Markov chain goverened by:
-\begin{itemize}
-    \item A state-transition probability matrix $A = [a_{ij}] \in \mathbb{R}^{M \times M}$ where $a_{ij} = p(z_{t+1} = q_j | z_t = q_i)$ 
-\end{itemize}
-\begin{itemize}
-    \item Initial state matrix $\pi = [\pi_i] \in \mathbb{R}_{1 \times M}$ with $\pi_i = p(z_1 = q_i)$
-\end{itemize}
-% \\[1ex]
-
-% ######################## Literature Review ######################################
-
-\section{Inserting Images}
-asdfpiojaseopritjf
-To insert an image, use the `graphicx` package. For example:
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=1\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
-    \caption{An example image.}
-    \label{fig:example}
-\end{figure}
-
-\section{Tables}
-You can create tables using the `tabular` environment or `booktabs` for professional-quality tables. For example:
-
-\begin{table}[h!]
-\centering
-\caption{Example Table}
-\begin{tabular}{@{}llr@{}}
-\toprule
-\textbf{Item} & \textbf{Description} & \textbf{Quantity} \\ \midrule
-Apples        & Fresh red apples     & 10                \\
-Oranges       & Juicy oranges        & 5                 \\
-Bananas       & Ripe bananas         & 7                 \\ \bottomrule
-\end{tabular}
-\label{tab:example}
-\end{table}
-
-\section{Hyperlinks}
-To add a hyperlink, use the `hyperref` package. For example:
-\href{https://www.latex-project.org/}{Visit the LaTeX project website}.
-
-\section{Conclusions}
-This is the conclusion section. Summarize your findings or leave final remarks.
-
-\newpage
-\appendix
-\section{Appendix}
-This is the appendix section, where you can include supplementary materials.
+% ######################## Background and Literature Review #########################
 

### 2025-02-08 Pic/main.py

@@ -1,0 +1,23 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Generate x values
+x = np.linspace(0, 8, 9)  # Values from 0 to 8 (inclusive)
+
+# Generate y values (exponential growth)
+y = 2 ** x  # Exponential function
+
+# Plot the graph
+plt.figure(figsize=(8, 6))
+plt.plot(x, y, marker='o', linestyle='-', color='b')
+
+# Add title and labels
+plt.title("Exponential Growth")
+plt.xlabel("x")
+plt.ylabel("f(x)")
+
+# Add grid
+plt.grid()
+
+# Show the plot
+plt.show()

### 2025-02-08 paper2/main.tex
update paper

@@
 \newpage
-\textbf{Gradient vanishing and gradient exploring}
-\\[1ex]
+\subsubsection{Gradient vanishing and gradient exploring}
 When training the RNN, BPTT was used to update the weight matrix.
@@
 Then the result will decrease exponentially, then vanishing gradients occur.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/exponential_growth.png} % Replace "example-image" with your image file
+\end{figure}
+
+\newpage
+% Long short-term memory (LSTM)
+\subsubsection{Long short-term memory (LSTM)}
+Long short-term memory proposed by \parencite{hochreiter1997lstm}.
+LSTM is designed for handling long time step problems.
+The architecture of LSTM can prevent vanishing gradient and exploring gradient.
+The main difference between LSTM and RNN is the number of gates.
+LSTM introduced input, forget and output gates.
+This allows LSTM to manage the flow of information more effectively, retaining important information over longer sequences.
+\\[2ex]
+Architecture
+\\[1ex]
+LSTMs introduce a memory cell that can maintain information over long time steps the cell is controlled by three gates, input gate, output gate, and forget gate.
+Each cell of LSTMs inside has 3 sigmoid and 1 tanh layer.
+Below graph unfolds the LSTM hence we can analyze different gates.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm1.png} % Replace "example-image" with your image file
+\end{figure}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm2.png} % Replace "example-image" with your image file
+\end{figure}
+\\
+% Forget gate
+Forget gate:
+\\[1ex]
+The forget gate is a component of the LSTM, designed to manage the flow of information within the cell state.
+The function of forget gate is to determine which information should be retained in memory cell \parencite{hochreiter1997lstm}.
+\begin{equation}
+    f_t = \sigma( W_f \cdot [ h_{t-1} , x_t ] + b_f ) 
+\end{equation}
+% Input gate
+Input gate: 
+\\[1ex]
+The input gate controls how much new information from the current time step is allowed to enter the cell \parencite{hochreiter1997lstm}.
+For the $\widetilde{C}_t$, the purpose is to suggest updates for the cell state.
+\begin{equation}
+    i_t = \sigma( W_i \cdot [ h_{t-1} , x_t ] + b_c ) 
+\end{equation}
+\begin{equation}
+    \widetilde{C}_t = tanh( W_c \cdot [ h_{t-1} , x_t ] + b_c )
+\end{equation}
+% Cell state update
+Cell State Update:
+\\[1ex]
+The forget gate will drop the meaningless information and add some potential information.
+\begin{equation}
+    C_t = f_t * C_{t-1} + i_t * \widetilde{C_t}
+\end{equation}
+% Output gate
+Output gate:
+\\[1ex]
+The output gate is able to control how much or what information from the cell state should be passed to the next layer or used in predictions. 
+\begin{equation}
+    o_t = \sigma( W_o \cdot [ h_{t-1} , x_t ] + b_o )
+\end{equation}
+% Hidden state update
+Hidden State Update:
+\\[1ex]
+The hidden state is influenced by output value and current cell state. 
+\begin{equation}
+    h_t = o_t * tanh(C_t)
+\end{equation}
+\\[1ex]
+n = number of features in the input vector $x_t$.\\
+m = number of units in LSTM.\\
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $x_t$               & Input vector at time t                        & $n \times 1$\\
+    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
+    $C_t$               & Cell state at time t                          & $m \times 1$\\
+    $f_t$               & Forget gate output at time t                  & $m \times 1$\\
+    $i_t$               & Input gate output at time t                   & $m \times 1$\\
+    $o_t$               & Output gate output at time t                  & $m \times 1$\\
+    $\widetilde{C_t}$   & Candidate memory cell at time t               & $m \times 1$\\
+    $W_f$               & Weight matrix for the forget gate             & $m \times (m+n)$\\
+    $W_i$               & Weight matrix for the input gate              & $m \times (m+n)$\\
+    $W_C$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
+    $W_o$               & Weight matrix for the output gate             & $m \times (m+n)$\\
+    $b_f$               & Bias vector for the forget gate               & $m \times 1$\\
+    $b_i$               & Bias vector for the input gate                & $m \times 1$\\
+    $b_C$               & Bias vector for the candidate memory cell     & $m \times 1$\\
+    $b_o$               & Bias vector for the output gate               & $m \times 1$\\
+    \hline
+    \caption{Unfolded RNN}
+\end{longtable}
+
+Number of parameters:\\
+\begin{enumerate}
+    % 1. Weights matrix for the input
+    \item Weights matrix for the input
+    \begin{itemize}
+        \item Forget gate: $n \times m$
+        \item Input gate: $n \times m$
+        \item Cell gate: $n \times m$
+        \item Output gate: $n \times m$
+    \end{itemize}
+
+    % 2. Weight matrix for the hidden state
+    \item Weight matrix for the hidden state
+    \begin{itemize}
+        \item Hidden state for forget gate: $m \times m$
+        \item Hidden state for input gate: $m \times m$
+        \item Hidden state for cell gate: $m \times m$
+        \item Hidden state for output gate: $m \times m$
+    \end{itemize}
+
+    % 3. Bias term
+    \item Bias term
+    \begin{itemize}
+        \item Bias for forget gate: $1 \times m$
+        \item Bias for input gate: $1 \times m$
+        \item Bias for cell gate: $1 \times m$
+        \item Bias for output gate: $1 \times m$
+    \end{itemize}
+\end{enumerate}
+Total parameters: $4 \times ( n + m + 1 ) \times m$
+
+\newpage
+% Gated Recurrent Unit (GRU)
+\subsubsection{Gated Recurrent Unit (GRU)}
+The gated recurrent unit (GRU) was proposed by \parencite{cho2014properties} to make each recurrent unit to adaptively capture dependencies of different time scales.
+The GRU has 2 gates, update gate and reset gate.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.65\textwidth]{../Pic/gru.png} % Replace "example-image" with your image file
+\end{figure}
+Update gate:
+\\[1ex]
+The update gate determines how much of the past information should be retained in the current hidden state. 
+\begin{equation}
+    z_t = \sigma(W_z \cdot [ h_{t-1} , x_t ] + b_z )
+\end{equation}
+% Reset gate
+Reset gate:
+\\[1ex]
+The reset gate is similar with the update gate, but the candidate hidden state is influenced by the reset gate. 
+\begin{equation}
+    r_t = \sigma( W_r \cdot [ h_{t-1} , x_t ] + b_r )
+\end{equation}
+Candidate hidden state:
+\\[1ex]
+The candidate hidden state combined with previous hidden state and current input to form the potential new information that can be added to the current hidden state.
+\begin{equation}
+    \widetilde{h_t} = tanh ( W_h \cdot [ r_t * h_{t-1} , x_t ] + b_h )
+\end{equation}
+Final hidden state
+\\[1ex]
+The final hidden state of the GRU at time t is a linear interpolation between the previous final hidden state and the candidate hidden state. 
+\begin{equation}
+    h_t = ( 1 - z_t ) * h_{t-1} + z_t * \widetilde{ h_t }
+\end{equation}
+
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $x_t$               & Input vector at time t                        & $n \times 1$\\
+    $h_t$               & Hidden state output at time t                 & $m \times 1$\\
+    $r_t$               & Reset gate output at time t                   & $m \times 1$\\
+    $z_t$               & Update gate output at time t                  & $m \times 1$\\
+    $W_z$               & Weight matrix for the update gate             & $m \times (m+n)$\\
+    $W_r$               & Weight matrix for the candidate memory cell   & $m \times (m+n)$\\
+    $W_h$               & Weight matrix for the output gate             & $m \times (m+n)$\\
+    $b_z$               & Bias vector for the input gate                & $m \times 1$\\
+    $b_r$               & Bias vector for the candidate memory cell     & $m \times 1$\\
+    $b_h$               & Bias vector for the output gate               & $m \times 1$\\
+    \hline
+\end{longtable}
+
+Number of parameters:\\
+\begin{enumerate}
+    % 1. Weights matrix for the input
+    \item Weights matrix for the input
+    \begin{itemize}
+        \item Update gate: $n \times m$
+        \item Reset gate: $n \times m$
+        \item Candidate hidden state: $n \times m$
+    \end{itemize}
+
+    % 2. Weight matrix for the hidden state
+    \item Weight matrix for the hidden state
+    \begin{itemize}
+        \item Hidden state for update gate: $m \times m$
+        \item Hidden state for reset gate: $m \times m$
+        \item Hidden state candidate hidden state: $m \times m$
+    \end{itemize}
+
+    % 3. Bias term
+    \item Bias term
+    \begin{itemize}
+        \item Bias for update: $1 \times m$
+        \item Bias for reset: $1 \times m$
+        \item Bias for candidate hidden state: $1 \times m$
+    \end{itemize}
+\end{enumerate}
+Total parameters: $3 \times ( n + m + 1 ) \times m$
+\newpage
+
+% Deep recurrent neural networks (DRNNs)
+\subsubsection{Deep recurrent neural networks (DRNNs)}
+Deep architecture networks with multiple layers that can hierarchically learn complex representations \parencite{bengio2009learning}.
+By extending of this concept, stacking recurrent layers to form Deep Recurrent Neural Networks aligns with this principle.
+A number of research papers have been prove that the performance of DRNNs is out-performance then conventional RNNs.
+\parencite{le2010deep, delalleau2011shallow, pascanu2013on}
+\\[1ex]
+The architecture of DRNNs are similar with conventional RNNs.
+We simply stack the recurrent layers vertically.
+For the first layer, this layer receives the input and combines it with its previous hidden state $h_{t-1}$ (Equation 38).
+The second layer receive the hidden state of the first layer and treat as the input of the layer 2 (Equation 39).
+We can extend this concept to $L$ layers (Equation 40).
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../Pic/drnn.png} % Replace "example-image" with your image file
+\end{figure}
+\\[1ex]
+\begin{equation}
+    h^{(1)}_t = f(W^{(1)}_{xh} x_{t} + W^{(1)}_{hh} h^{(1)}_{t-1} + b^{(1)}_h)
+\end{equation}
+\begin{equation}
+    h^{(2)}_t = f(W^{(2)}_{xh} h^{(1)}_{t} + W^{(2)}_{hh} h^{(2)}_{t-1} + b^{(2)}_h)
+\end{equation}
+\begin{equation}
+    h^{(L)}_t = f(W^{(L)}_{xh} h^{(L-1)}_{t} + W^{(L)}_{hh} h^{(L)}_{t-1} + b^{(L)}_h)
+\end{equation}
+\begin{equation}
+    o_t = W_{hy}h^{(L)}_{t} + b_y
+\end{equation}
+\begin{equation}
+    y_t = g(o_t)
+\end{equation}
+Where $x_t$ is the input at time $t$.
+$h^{(l)}_{t}$ is the hidden state for the $l$ layer at time t.
+$W^{(l)}_{xh}, W^{(l)}_{hh}$ are the weight matrices for the input to hidden and hidden to hidden connections in layer $l$, respectively.
+$W_{hy}$ is weight matrix for the output layer.
+$b^{(l)}_h$ is the bias vector for the $l$ layer (except output layer), $b_y$ is the bias vector for output layer.
+$g(\cdot)$ and $f(\cdot)$ are an activation function.
+
+\newpage
+\subsubsection{Hidden Markov Model}
+Before reviewing Hidden Markov Model (HMM), it is essential to understand what Markov models is, or Markov chains.
+Markov chains are fundamental models in probability theory and statistic, and it is a stochastic process describing a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
+This property known as the Markov property.
+In rigorous terms, let the state space be defined as:
+Let the state space be defined as:
+\begin{equation}
+    S = \{ s_1, s_2, \dots, s_N \}.
+\end{equation}
+\begin{equation}
+    P(s_{t+1} \mid s_t, s_{t-1}, s_{t-2}, \cdots, s_{1}) = P(s_{t+1} \mid s_t)
+\end{equation}
+The state transition probability is denoted by:
+\begin{equation}
+    P_{ij} = P(s_{t+1} = s_j \mid s_t = s_i).
+\end{equation}
+
+Given an initial state \( s_1 \) with probability \( P(s_1) \), the joint probability of a state sequence \( \{s_1, s_2, \dots, s_T\} \) can be written as:
+\begin{equation}
+    P(s_1, s_2, \dots, s_T) = P(s_1) \cdot P(s_2 \mid s_1) \cdot P(s_3 \mid s_2) \cdots P(s_T \mid s_{T-1})
+\end{equation}
+We can simplfy the joint probability of a state sequence as equestion 47.
+\begin{equation}
+    P(s_1, s_2, \dots, s_T) = P(s_1) \prod_{t=1}^{T-1} P(s_{t+1} \mid s_{t})
+\end{equation}
+
+In the context of Natural Language Processing (NLP), early language models utilized the Markov assumption by approximating the probability of a word sequence.
+For example, for a bigram model, the probability of a sentence \( w_1, w_2, \dots, w_T \) is given by:
+$$
+P(w_1, w_2, \dots, w_T) \approx \prod_{t=1}^{T} P(w_t \mid w_{t-1}),
+$$
+where each word is assumed to depend only on its immediate predecessor.
+Although this simplification makes computation manageable, it does not account for deeper syntactic or semantic structures inherent in language.
+
+\newpage
+
+Literature Review on Markov Chains
+
+Markov chains are fundamental models in probability theory that describe systems evolving in discrete time where the future behaviour of the process depends solely on its current state—a property known as the **Markov property**.
+In rigorous terms, let \((\mathcal{X}, \mathcal{F})\) be a measurable space and let \(\{X_n\}_{n\ge0}\) be a sequence of random variables taking values in \(\mathcal{X}\).
+The process is said to be a *Markov chain* if, for every \(n\ge0\) and for every \(A\in\mathcal{F}\),
+\[
+P\bigl(X_{n+1}\in A \mid X_n, X_{n-1}, \dots, X_0\bigr) = P\bigl(X_{n+1}\in A \mid X_n\bigr).
+\]
+This memoryless property implies that the current state \(X_n\) fully encapsulates all the probabilistic information necessary for predicting the future evolution of the process.
+
+Stationarity and Convergence
+
+A probability distribution \(\pi\) on \(\mathcal{X}\) is called *stationary* (or invariant) for the Markov chain if it satisfies
+\[
+\pi(A) = \int_{\mathcal{X}}P(x,A)\,\pi(dx) \quad \text{for all } A\in\mathcal{F},
+\]
+where \(P(x,A)\) denotes the one-step transition probability.
+Under appropriate conditions—most notably, **\(\varphi\)-irreducibility** and **aperiodicity**—it can be rigorously shown that the distribution of the chain converges to \(\pi\) in total variation distance.
+Specifically, for \(\pi\)-almost every initial state \(x\),
+\[
+\lim_{n\to\infty} \|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} = 0,
+\]
+where the total variation distance between two probability measures \(\mu\) and \(\nu\) is defined as
+\[
+\|\mu-\nu\|_{\text{TV}} = \sup_{A\in\mathcal{F}}|\mu(A)-\nu(A)|.
+\]
+Such convergence results are established in depth in Roberts and Rosenthal (2004) and form the theoretical backbone for many modern computational methods.
+
+Small Sets, Drift Conditions, and Geometric Ergodicity
+
+For Markov chains on general (and possibly uncountable) state spaces, it is crucial to establish conditions that guarantee not only convergence but also quantitative rates of convergence.
+One such condition involves the concept of a **small set**.
+A subset \(C\subseteq \mathcal{X}\) is called small if there exist a positive integer \(n_0\), a constant \(\epsilon>0\), and a probability measure \(\nu\) on \(\mathcal{X}\) such that
+\[
+P^{n_0}(x,A) \ge \epsilon\,\nu(A), \quad \text{for all } x\in C \text{ and all } A\in\mathcal{F}.
+\]
+This minorisation condition is instrumental in establishing uniform convergence bounds.
+
+Complementary to the small set condition is a **drift condition**.
+A drift (or Lyapunov) function \(V:\mathcal{X}\to[1,\infty)\) is used to quantify the tendency of the chain to return to a particular subset \(C\).
+The chain is said to satisfy a drift condition if there exist constants \(\lambda < 1\) and \(b < \infty\) such that
+\[
+P V(x) \le \lambda\, V(x) + b\,1_C(x) \quad \text{for all } x\in\mathcal{X},
+\]
+where \(P V(x) = \int_{\mathcal{X}} V(y)P(x,dy)\) and \(1_C\) is the indicator function of \(C\).
+When a Markov chain satisfies both a minorisation condition and an appropriate drift condition, it is often **geometrically ergodic**; that is, there exist a function \(M(x)\) and a constant \(\rho\in(0,1)\) such that
+\[
+\|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} \le M(x)\,\rho^n, \quad \text{for all } n\ge0.
+\]
+Such quantitative convergence rates are crucial for both theoretical investigations and practical applications (see Roberts and Rosenthal, 2004; Meyn and Tweedie, 1993).
+
+Coupling, Regeneration, and Quantitative Bounds
+
+A powerful method for analyzing the convergence of Markov chains is the use of **coupling techniques**.
+In this approach, one constructs a joint process \((X_n, Y_n)\) on \(\mathcal{X}\times\mathcal{X}\) such that:
+- The marginal dynamics of \(X_n\) and \(Y_n\) each follow the prescribed transition kernel \(P\).
+- The coupling is designed so that the probability \(P(X_n\neq Y_n)\) decreases over time.
+
+A key consequence of coupling is the following inequality:
+\[
+\|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} \le P(X_n \neq Y_n).
+\]
+By constructing couplings that "meet" (or "couple") quickly, one may deduce explicit bounds on the rate of convergence in total variation.
+Similar ideas underpin **regeneration techniques**, which partition the Markov chain into independent, identically distributed tours.
+Regeneration points allow the application of classical renewal theory to derive central limit theorems (CLTs) for additive functionals of the chain.
+For example, if \(h:\mathcal{X}\to\mathbb{R}\) is a function with \(\pi(|h|^2)<\infty\), then under appropriate conditions
+\[
+\frac{1}{\sqrt{n}}\sum_{i=0}^{n-1}\!\Bigl(h(X_i)-\pi(h)\Bigr) \xrightarrow{d} \mathcal{N}(0,\sigma^2),
+\]
+with the asymptotic variance \(\sigma^2\) often expressible in terms of the chain’s autocovariance structure.
 

### 2025-02-08 Pic/main.py

@@ -2,22 +2,29 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 # Generate x values
-x = np.linspace(0, 8, 9)  # Values from 0 to 8 (inclusive)
+x = np.linspace(0, 15, 9)  # Values from 0 to 8 (inclusive)
 
 # Generate y values (exponential growth)
-y = 2 ** x  # Exponential function
+y1 = 0.5 ** x  # Exponential function
+y2 = 2 ** x  # Exponential function
 
 # Plot the graph
-plt.figure(figsize=(8, 6))
-plt.plot(x, y, marker='o', linestyle='-', color='b')
+fig = plt.figure()
+plt.figure(figsize=(8,4))
 
-# Add title and labels
-plt.title("Exponential Growth")
+plt.subplot(211)
+plt.plot(x, y1, marker='o', linestyle='-', color='b')
+plt.title("Exponential Grow")
 plt.xlabel("x")
 plt.ylabel("f(x)")
 
-# Add grid
-plt.grid()
+plt.subplot(212)
+plt.plot(x, y2, marker='o', linestyle='-', color='r')
+plt.xlabel("x")
+plt.ylabel("f(x)")
+
+# Save the figure as a PNG file
+plt.savefig("exponential_growth.png", dpi=300)  # Save with high resolution (300 DPI)
 
 # Show the plot
 plt.show()

### 2025-02-09 paper2/main.tex
paper update

@@
 Markov chains are fundamental models in probability theory and statistic, and it is a stochastic process describing a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
-This property known as the Markov property.
+This property known as the Markov property \parencite{roberts2004general}.
 In rigorous terms, let the state space be defined as:
@@
 \end{equation}
-
-In the context of Natural Language Processing (NLP), early language models utilized the Markov assumption by approximating the probability of a word sequence.
-For example, for a bigram model, the probability of a sentence \( w_1, w_2, \dots, w_T \) is given by:
-$$
-P(w_1, w_2, \dots, w_T) \approx \prod_{t=1}^{T} P(w_t \mid w_{t-1}),
-$$
-where each word is assumed to depend only on its immediate predecessor.
-Although this simplification makes computation manageable, it does not account for deeper syntactic or semantic structures inherent in language.
+Then it is essential to discuss marginal probability in a Markov model because it provides critical information about the likelihood of being in particular state $s_j$ at a specific time $t$.
+The marginal probability of being in state $s_j$ at time $t$ defined as:
+\begin{equation}
+    \pi_t(s_j) = P(s_t=s_j)
+\end{equation}
+It is the probability of the system being in state $s_j$ at time t, irrespective of how the system transitioned to $s_j$.
+We can obtainde by summing over all possible transition probability $P_{ij}$ where $s_i$ belong in state space.
+\begin{equation}
+    \pi_{t+1}(s_j) = \sum_{s_i \in S} \pi_t(s_i) \cdot P_{ij}
+\end{equation}
+If Markov chain is irreducible and aperiodic, the stationary distribution defined as below:
+\begin{equation}
+    \pi_j = \sum_{i=1}^{n} \pi_i P_{ij}
+\end{equation}
 
 \newpage
+\textbf{Hidden markov model}
+\\[1ex]
+Hidden Markov Model (HMMs) is a statistical model that extends the Markov chains by introducting a layer of latent (unobservable) variables.
+And connecting them to observable outputs (emission probability).
+In contrast to standard Markov chains, hidden states are never be observed directly.
+Instead, we observe a sequence of data (observations) that are probabilitically generated by these hidden state.
+This structure has proven useful for applications in speech recognition, biological sequence analysis, and signal processing \parencite{rabiner1989atutorial}.
+Lets defined the
+% ####################################################
+\newpage
 
-Literature Review on Markov Chains
+Literature Review on Hidden Markov Models
 
-Markov chains are fundamental models in probability theory that describe systems evolving in discrete time where the future behaviour of the process depends solely on its current state—a property known as the **Markov property**.
-In rigorous terms, let \((\mathcal{X}, \mathcal{F})\) be a measurable space and let \(\{X_n\}_{n\ge0}\) be a sequence of random variables taking values in \(\mathcal{X}\).
-The process is said to be a *Markov chain* if, for every \(n\ge0\) and for every \(A\in\mathcal{F}\),
+Hidden Markov Models (HMMs) extend classical Markov chains by incorporating an additional layer of latent (unobservable) variables that govern the generation of the observed data.
+In contrast to standard Markov chains—where the state of the system is directly observed—an HMM assumes that the underlying process is hidden and that each observed datum is stochastically generated from the corresponding hidden state.
+This two-layer structure has proven to be both theoretically interesting and practically useful for applications in speech recognition, biological sequence analysis, and signal processing (Rabiner, 1989).
+
+1. From Markov Chains to Hidden Markov Models
+
+Let \(\{X_n\}_{n \ge 1}\) be a Markov chain on a measurable space \((\mathcal{X}, \mathcal{F})\) satisfying
 \[
-P\bigl(X_{n+1}\in A \mid X_n, X_{n-1}, \dots, X_0\bigr) = P\bigl(X_{n+1}\in A \mid X_n\bigr).
+P\bigl(X_{n+1} \in A \mid X_n, X_{n-1}, \dots, X_1\bigr) = P\bigl(X_{n+1} \in A \mid X_n\bigr) \quad \text{for all } A \in \mathcal{F}.
 \]
-This memoryless property implies that the current state \(X_n\) fully encapsulates all the probabilistic information necessary for predicting the future evolution of the process.
+This *Markov property* ensures that the future evolution of the process depends only on its present state, not on the full past history.
 
-Stationarity and Convergence
+In extending this framework to HMMs, we introduce an observable process \(\{Y_n\}_{n \ge 1}\) defined on another measurable space \((\mathcal{Y}, \mathcal{G})\).
+The connection between the hidden states and the observations is governed by the following assumption:
 
-A probability distribution \(\pi\) on \(\mathcal{X}\) is called *stationary* (or invariant) for the Markov chain if it satisfies
+- **Conditional Independence of Observations:** Given the current hidden state,
   \[
-\pi(A) = \int_{\mathcal{X}}P(x,A)\,\pi(dx) \quad \text{for all } A\in\mathcal{F},
+  P\bigl(Y_n \in B \mid X_n, X_{n-1}, \dots, X_1, Y_{n-1}, \dots, Y_1\bigr) = P\bigl(Y_n \in B \mid X_n\bigr)
   \]
-where \(P(x,A)\) denotes the one-step transition probability.
-Under appropriate conditions—most notably, **\(\varphi\)-irreducibility** and **aperiodicity**—it can be rigorously shown that the distribution of the chain converges to \(\pi\) in total variation distance.
-Specifically, for \(\pi\)-almost every initial state \(x\),
+  for every \(B \in \mathcal{G}\). This states that the observation \(Y_n\) is independent of past states and observations, conditioned on \(X_n\).
+
+Thus, an HMM is fully specified by the following three components:
+1. The **initial distribution** \(\pi\) on \(\mathcal{X}\):
    \[
-\lim_{n\to\infty} \|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} = 0,
+   \pi(x) = P(X_1 = x).
    \]
-where the total variation distance between two probability measures \(\mu\) and \(\nu\) is defined as
+2. The **transition kernel** \(P(x, dy)\) of the hidden Markov chain:
    \[
-\|\mu-\nu\|_{\text{TV}} = \sup_{A\in\mathcal{F}}|\mu(A)-\nu(A)|.
+   p_{ij} = P(X_{n+1} = j \mid X_n = i).
    \]
-Such convergence results are established in depth in Roberts and Rosenthal (2004) and form the theoretical backbone for many modern computational methods.
+3. The **emission (observation) distribution** \(B(x, dy)\) or \(b_x(y)\):
+   \[
+   b_x(y) = P(Y_n = y \mid X_n = x).
+   \]
 
-Small Sets, Drift Conditions, and Geometric Ergodicity
+2. Joint and Marginal Distributions
 
-For Markov chains on general (and possibly uncountable) state spaces, it is crucial to establish conditions that guarantee not only convergence but also quantitative rates of convergence.
-One such condition involves the concept of a **small set**.
-A subset \(C\subseteq \mathcal{X}\) is called small if there exist a positive integer \(n_0\), a constant \(\epsilon>0\), and a probability measure \(\nu\) on \(\mathcal{X}\) such that
+Using the above structure, the joint probability of a sequence of hidden states and observations \(\{(X_1, Y_1), \dots, (X_N, Y_N)\}\) is given by
 \[
-P^{n_0}(x,A) \ge \epsilon\,\nu(A), \quad \text{for all } x\in C \text{ and all } A\in\mathcal{F}.
+\begin{aligned}
+P\bigl(X_1, \dots, X_N,\, Y_1, \dots, Y_N\bigr) 
+&= \pi(X_1) \, b_{X_1}(Y_1) \prod_{n=2}^N P\bigl(X_n \mid X_{n-1}\bigr) \, b_{X_n}(Y_n)\\[1mm]
+&= \pi(X_1)\, b_{X_1}(Y_1)\, p_{X_1X_2}\, b_{X_2}(Y_2) \cdots p_{X_{N-1}X_N}\, b_{X_N}(Y_N).
+\end{aligned}
 \]
-This minorisation condition is instrumental in establishing uniform convergence bounds.
+Marginalization over the hidden states yields the likelihood of the observations:
+\[
+P(Y_1, \dots, Y_N) = \sum_{X_1 \in \mathcal{X}} \cdots \sum_{X_N \in \mathcal{X}} \pi(X_1)\, \prod_{n=2}^N p_{X_{n-1}X_n}\, \prod_{n=1}^N b_{X_n}(Y_n),
+\]
+which in many applications is used for parameter estimation and inference.
 
-Complementary to the small set condition is a **drift condition**.
-A drift (or Lyapunov) function \(V:\mathcal{X}\to[1,\infty)\) is used to quantify the tendency of the chain to return to a particular subset \(C\).
-The chain is said to satisfy a drift condition if there exist constants \(\lambda < 1\) and \(b < \infty\) such that
+3. Inference and Algorithmic Developments
+
+Due to the latent structure in HMMs, direct computation of the likelihood or the posterior distribution of the hidden states is usually intractable when the state space is large or continuous.
+This has spurred the development of several efficient algorithms:
+
+- **The Forward-Backward Algorithm:** Utilizes dynamic programming to recursively compute the probabilities of partial observation sequences. For instance, denoting the forward variable by
   \[
-P V(x) \le \lambda\, V(x) + b\,1_C(x) \quad \text{for all } x\in\mathcal{X},
+  \alpha_n(x) = P(Y_1, Y_2, \dots, Y_n, X_n = x),
   \]
-where \(P V(x) = \int_{\mathcal{X}} V(y)P(x,dy)\) and \(1_C\) is the indicator function of \(C\).
-When a Markov chain satisfies both a minorisation condition and an appropriate drift condition, it is often **geometrically ergodic**; that is, there exist a function \(M(x)\) and a constant \(\rho\in(0,1)\) such that
+  one can compute
   \[
-\|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} \le M(x)\,\rho^n, \quad \text{for all } n\ge0.
+  \alpha_{n+1}(y) = \left[\sum_{x \in \mathcal{X}} \alpha_n(x) \, p_{xy}\right]\, b_y(Y_{n+1}).
   \]
-Such quantitative convergence rates are crucial for both theoretical investigations and practical applications (see Roberts and Rosenthal, 2004; Meyn and Tweedie, 1993).
 
-Coupling, Regeneration, and Quantitative Bounds
+- **The Viterbi Algorithm:** Determines the most probable sequence of hidden states (i.e., finds \(\hat{X}_1, \dots, \hat{X}_N\) such that the joint probability is maximized).
 
-A powerful method for analyzing the convergence of Markov chains is the use of **coupling techniques**.
-In this approach, one constructs a joint process \((X_n, Y_n)\) on \(\mathcal{X}\times\mathcal{X}\) such that:
-- The marginal dynamics of \(X_n\) and \(Y_n\) each follow the prescribed transition kernel \(P\).
-- The coupling is designed so that the probability \(P(X_n\neq Y_n)\) decreases over time.
+- **The Baum-Welch Algorithm (Expectation-Maximization):** Iteratively estimates the unknown parameters (transition probabilities and emission probabilities) of the HMM when they are not known a priori.
 
-A key consequence of coupling is the following inequality:
+These methods rely on the underlying assumption that the hidden state process is a Markov chain, and many of the rigorous convergence and mixing properties discussed in the literature on general state space Markov chains (e.g., Roberts and Rosenthal, 2004; Meyn and Tweedie, 1993) carry over—often with additional considerations due to the unobserved structure.
+
+4. Theoretical Underpinnings and Convergence
+
+As with observable Markov chains, the study of HMMs involves rigorous mathematical analysis centered on convergence properties and ergodicity.
+Under appropriate conditions (e.g., the hidden chain is \(\varphi\)-irreducible, aperiodic, and satisfies a suitable drift condition), one can prove that the filtering distributions
 \[
-\|P^n(x,\cdot)-\pi(\cdot)\|_{\text{TV}} \le P(X_n \neq Y_n).
+P(X_n \in \cdot \mid Y_1, \dots, Y_n)
 \]
-By constructing couplings that "meet" (or "couple") quickly, one may deduce explicit bounds on the rate of convergence in total variation.
-Similar ideas underpin **regeneration techniques**, which partition the Markov chain into independent, identically distributed tours.
-Regeneration points allow the application of classical renewal theory to derive central limit theorems (CLTs) for additive functionals of the chain.
-For example, if \(h:\mathcal{X}\to\mathbb{R}\) is a function with \(\pi(|h|^2)<\infty\), then under appropriate conditions
+converge to a unique invariant measure.
+Techniques such as **coupling** and **regeneration**—which have been successfully employed in establishing quantitative convergence rates for Markov chains (Roberts \& Rosenthal, 2004)—are also adapted to the HMM context to study the stability and consistency of inference procedures.
+
+For example, under the standard assumptions of an HMM, if the hidden state chain is geometrically ergodic then for a function \(h : \mathcal{X} \to \mathbb{R}\) with finite second moment under the stationary measure \(\pi\), the central limit theorem holds for the additive functional
 \[
-\frac{1}{\sqrt{n}}\sum_{i=0}^{n-1}\!\Bigl(h(X_i)-\pi(h)\Bigr) \xrightarrow{d} \mathcal{N}(0,\sigma^2),
+\frac{1}{\sqrt{n}} \sum_{i=1}^n \left[h(X_i) - \pi(h)\right],
 \]
-with the asymptotic variance \(\sigma^2\) often expressible in terms of the chain’s autocovariance structure.
+which is essential for quantifying the uncertainty in estimators derived via the Expectation-Maximization or sequential Monte Carlo methods (see also Billingsley, 1995).
 
+5. Conclusion
+
+In summary, by extending the framework of Markov chains—characterized by the Markov property, established convergence criteria, small set theory, and drift conditions—to a hidden state setting, Hidden Markov Models provide a robust foundation for dealing with complex sequence data where the underlying process is not directly observable.
+This extension leads to rich algorithmic and inferential techniques that have been rigorously analyzed and widely adopted in several applied fields.
 % ######################## Background and Literature Review #########################

### 2025-02-09 paper2/main.tex
paper update

@@
 Markov chains are fundamental models in probability theory and statistic, and it is a stochastic process describing a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
-This property known as the Markov property \parencite{roberts2004general}.
-In rigorous terms, let the state space be defined as:
+This property known as the Markov property \parencite{roberts2004general, rabiner1986anintroduction}.
+In rigorous terms.
 Let the state space be defined as:
@@
 \newpage
+
+% Hidden markov model 
 \textbf{Hidden markov model}
@@
 This structure has proven useful for applications in speech recognition, biological sequence analysis, and signal processing \parencite{rabiner1989atutorial}.
-Lets defined the
-% ####################################################
+Lets the hidden states be denoted by $X_t$, where $X_t$ is the hidden state at time $t$.
+The hidden state space is defined as:
+\begin{equation}
+    \mathcal{X} = \{s_1, s_2, \cdots, s_n\}
+\end{equation}
+where $\mathcal{X}$ is finite set of size $n$, and each $s_i \in \mathcal{X}$ represents a specific possible hidden state.
+Then for the observable space represents the possible outcomes (observations) that are generated by the hidden state, we have discuss previously.
+Let the observations at time $t$ be denoted by $Y_t$.
+The observable space is defined as:
+\begin{equation}
+    \mathcal{Y} = \{o_1, o_2, \cdots, o_m\}
+\end{equation}
+where $\mathcal{Y}$ is finite set of size $m$, and each $o_i \in \mathcal{Y}$ represents a specific possible observation. 
+\\[1ex]
+Then for the transition probability matrix ($P$), the probability of transitioning between hidden states denoted as:
+\begin{equation}
+    P_{n \times n} = 
+    \left[ {\begin{array}{cccc}
+        p_{11} & p_{12} & \cdots & p_{1n}\\
+        p_{21} & p_{22} & \cdots & p_{2n}\\
+        \vdots & \vdots & \ddots & \vdots\\
+        p_{n1} & p_{n2} & \cdots & p_{nn}\\
+    \end{array} } \right]
+\end{equation}
+\begin{equation}
+    p_{ij} = P(X_{t+1} = s_j \mid X_t = s_i) 
+\end{equation}
+where $s_j, s_i \in \mathcal{X}$.
+\\[1ex]
+For emission probability matrix ($B$), the probabilities of generating an observation given a hidden state.
+\begin{equation}
+    B_{n \times n} = 
+    \left[ {\begin{array}{cccc}
+        b_{11} & b_{12} & \cdots & b_{1m}\\
+        b_{21} & b_{22} & \cdots & b_{2m}\\
+        \vdots & \vdots & \ddots & \vdots\\
+        b_{m1} & b_{m2} & \cdots & b_{mm}\\
+    \end{array} } \right]
+\end{equation}
+\begin{equation}
+    b_{ik} = P(Y_{t} = o_k \mid X_t = s_i) 
+\end{equation}
+where $s_i \in \mathcal{X}, o_k \in \mathcal{Y}$.
+\\[1ex]
+And the goal of hidden markov model is the maximize the joint probability $P(X, Y)$ given the initial probability $P(X_1)$.
+\begin{equation}
+    P(X, Y) = P(X_1) \cdot \prod_{t=1}^{T-1}P(X_{t+1}=s_j \mid X_{t}=s_i) \cdot \prod_{t=1}^{T} P(Y_t=o_k \mid X_t=s_i)
+\end{equation}
+
+For the NLP tasks, there may have some unobservable data, for example the implicit sentiment, user intent, emotional state.
+\parencite{rabiner1993fundamentals} proposes the hidden Markov model (HMM), the assumption is the existence of the latent process follows a Markov chain from which observations X are generated.
+In other word, there would exists an unobserved state sequence $Z = \{z_1, z_2 \cdots , z_T\}$ in observed sequence $X = \{x_1, x_2, \cdots, x_T\}$ \parencite{sengupta2023hybrid}.
+Where the hidden states, $z_t$ belonging to state-space $Q = \{q_1, q_2, \cdots, q_M \}$ follow a Markov chain goverened by:
+\begin{itemize}
+    \item A state-transition probability matrix $A = [a_{ij}] \in \mathbb{R}^{M \times M}$ where $a_{ij} = p(z_{t+1} = q_j | z_t = q_i)$ 
+\end{itemize}
+\begin{itemize}
+    \item Initial state matrix $\pi = [\pi_i] \in \mathbb{R}_{1 \times M}$ with $\pi_i = p(z_1 = q_i)$
+\end{itemize}
+Furthermore, for the hidden state $z_t$, correspoinding to the observe data $x_t$ is release by emission process $B = [b_j(x)]$ where $b_j(x) = p(x|z=q_i)$.
+We can assume $b_j(x)$ is follows the Gaussian mixture model (GMM).
+\begin{equation}
+    p(x_t|z=q_j) = \sum_{l=1}^{k}c_{jl}\mathcal{N}(x_t|\mu_{jl}, \Sigma_{jl})
+\end{equation}
+where $\sum_{l=1}^{k}c_{jl} = 1$, $\forall j = \{1,\cdots, M\}$, k is the number of Gaussian mixture components and $\mathcal{N}(x_t|\mu_{jl}, \Sigma_{jl})$ denotes a Gaussian probability density with mean $\mu_{jl}$ and covariance $\Sigma_{jl}$ for state $j$ and mixture component $l$.
+The number of hidden states ($M$) and mixture component ($k$) are the two hyperparameters of the model which have to be provided apriori.
+\\[1ex]
+Therefore, the joint probability probability density function of the observation $X$ can be expressed as:
+\begin{equation}
+    p(X) = p(z_1) \prod_{t=1}^{T-1} p(z_{t+1}|z_{t}) \prod_{t=1}^{T} p(x_t|z_t)
+\end{equation}
+The optimal parameters $[A, B, \pi]$, which maximize the likelihood of the observation sequence $X$ (Equation 39), are determined using the Baum-Welch algorithm, an expectation-maximization method \parencite{rabiner1993fundamentals}.
+Additionally, the probability of the system in a specific hidden state $z_t$ corresponding to the observation $x_t$ is calculated using the Viterbi algorithm.
+
 \newpage
 
-Literature Review on Hidden Markov Models
+% Word representation 
+\subsubsection{Word representation}
+In RNN, the due to the architecture, the model can only handle vector, it is important that convert the word into machine readable format.
+By translating words into vectors, these representations capture semantic meanings, relationships and contexts.
+\parencite{mikolov2013efficient} proposed 2 model architectures for leaning distributed representations of words.
+\\[1ex]
+\textbf{Continuous Bag of Words model}
+CBOW operates by taking context words around a target word, it is flexible to adjust the window size which mean that we can control how many words around our target word.
+Hence aggregating their embeddings and passing to hidden layer to produce a probability distribution.
+Capture the semantic meanings and the relationship more effectively.
+\begin{equation}
+    Q = N \times D + D \times log_2(V)
+\end{equation}
+\\[1ex]
+\textbf{Continuous Skip gram model}
+\\[1ex]
+The continuous skip gram model operates in opposite direction compare with CBOW.
+The model takes the surrounding word to predict the target word.
+In other word, treat the current word as an input to a log-linear classifier with continuous projection layer, and predict the probability distribution of surrounding words.
+\begin{equation}
+    Q = C \times (D + D \times log_2(V))
+\end{equation}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/word_representation.png} % Replace "example-image" with your image file
+\end{figure}
 
-Hidden Markov Models (HMMs) extend classical Markov chains by incorporating an additional layer of latent (unobservable) variables that govern the generation of the observed data.
-In contrast to standard Markov chains—where the state of the system is directly observed—an HMM assumes that the underlying process is hidden and that each observed datum is stochastically generated from the corresponding hidden state.
-This two-layer structure has proven to be both theoretically interesting and practically useful for applications in speech recognition, biological sequence analysis, and signal processing (Rabiner, 1989).
+\newpage
 
-1. From Markov Chains to Hidden Markov Models
-
-Let \(\{X_n\}_{n \ge 1}\) be a Markov chain on a measurable space \((\mathcal{X}, \mathcal{F})\) satisfying
-\[
-P\bigl(X_{n+1} \in A \mid X_n, X_{n-1}, \dots, X_1\bigr) = P\bigl(X_{n+1} \in A \mid X_n\bigr) \quad \text{for all } A \in \mathcal{F}.
-\]
-This *Markov property* ensures that the future evolution of the process depends only on its present state, not on the full past history.
-
-In extending this framework to HMMs, we introduce an observable process \(\{Y_n\}_{n \ge 1}\) defined on another measurable space \((\mathcal{Y}, \mathcal{G})\).
-The connection between the hidden states and the observations is governed by the following assumption:
-
-- **Conditional Independence of Observations:** Given the current hidden state,
-  \[
-  P\bigl(Y_n \in B \mid X_n, X_{n-1}, \dots, X_1, Y_{n-1}, \dots, Y_1\bigr) = P\bigl(Y_n \in B \mid X_n\bigr)
-  \]
-  for every \(B \in \mathcal{G}\). This states that the observation \(Y_n\) is independent of past states and observations, conditioned on \(X_n\).
-
-Thus, an HMM is fully specified by the following three components:
-1. The **initial distribution** \(\pi\) on \(\mathcal{X}\):
-   \[
-   \pi(x) = P(X_1 = x).
-   \]
-2. The **transition kernel** \(P(x, dy)\) of the hidden Markov chain:
-   \[
-   p_{ij} = P(X_{n+1} = j \mid X_n = i).
-   \]
-3. The **emission (observation) distribution** \(B(x, dy)\) or \(b_x(y)\):
-   \[
-   b_x(y) = P(Y_n = y \mid X_n = x).
-   \]
-
-2. Joint and Marginal Distributions
-
-Using the above structure, the joint probability of a sequence of hidden states and observations \(\{(X_1, Y_1), \dots, (X_N, Y_N)\}\) is given by
-\[
-\begin{aligned}
-P\bigl(X_1, \dots, X_N,\, Y_1, \dots, Y_N\bigr) 
-&= \pi(X_1) \, b_{X_1}(Y_1) \prod_{n=2}^N P\bigl(X_n \mid X_{n-1}\bigr) \, b_{X_n}(Y_n)\\[1mm]
-&= \pi(X_1)\, b_{X_1}(Y_1)\, p_{X_1X_2}\, b_{X_2}(Y_2) \cdots p_{X_{N-1}X_N}\, b_{X_N}(Y_N).
-\end{aligned}
-\]
-Marginalization over the hidden states yields the likelihood of the observations:
-\[
-P(Y_1, \dots, Y_N) = \sum_{X_1 \in \mathcal{X}} \cdots \sum_{X_N \in \mathcal{X}} \pi(X_1)\, \prod_{n=2}^N p_{X_{n-1}X_n}\, \prod_{n=1}^N b_{X_n}(Y_n),
-\]
-which in many applications is used for parameter estimation and inference.
-
-3. Inference and Algorithmic Developments
-
-Due to the latent structure in HMMs, direct computation of the likelihood or the posterior distribution of the hidden states is usually intractable when the state space is large or continuous.
-This has spurred the development of several efficient algorithms:
-
-- **The Forward-Backward Algorithm:** Utilizes dynamic programming to recursively compute the probabilities of partial observation sequences. For instance, denoting the forward variable by
-  \[
-  \alpha_n(x) = P(Y_1, Y_2, \dots, Y_n, X_n = x),
-  \]
-  one can compute
-  \[
-  \alpha_{n+1}(y) = \left[\sum_{x \in \mathcal{X}} \alpha_n(x) \, p_{xy}\right]\, b_y(Y_{n+1}).
-  \]
-
-- **The Viterbi Algorithm:** Determines the most probable sequence of hidden states (i.e., finds \(\hat{X}_1, \dots, \hat{X}_N\) such that the joint probability is maximized).
-
-- **The Baum-Welch Algorithm (Expectation-Maximization):** Iteratively estimates the unknown parameters (transition probabilities and emission probabilities) of the HMM when they are not known a priori.
-
-These methods rely on the underlying assumption that the hidden state process is a Markov chain, and many of the rigorous convergence and mixing properties discussed in the literature on general state space Markov chains (e.g., Roberts and Rosenthal, 2004; Meyn and Tweedie, 1993) carry over—often with additional considerations due to the unobserved structure.
-
-4. Theoretical Underpinnings and Convergence
-
-As with observable Markov chains, the study of HMMs involves rigorous mathematical analysis centered on convergence properties and ergodicity.
-Under appropriate conditions (e.g., the hidden chain is \(\varphi\)-irreducible, aperiodic, and satisfies a suitable drift condition), one can prove that the filtering distributions
-\[
-P(X_n \in \cdot \mid Y_1, \dots, Y_n)
-\]
-converge to a unique invariant measure.
-Techniques such as **coupling** and **regeneration**—which have been successfully employed in establishing quantitative convergence rates for Markov chains (Roberts \& Rosenthal, 2004)—are also adapted to the HMM context to study the stability and consistency of inference procedures.
-
-For example, under the standard assumptions of an HMM, if the hidden state chain is geometrically ergodic then for a function \(h : \mathcal{X} \to \mathbb{R}\) with finite second moment under the stationary measure \(\pi\), the central limit theorem holds for the additive functional
-\[
-\frac{1}{\sqrt{n}} \sum_{i=1}^n \left[h(X_i) - \pi(h)\right],
-\]
-which is essential for quantifying the uncertainty in estimators derived via the Expectation-Maximization or sequential Monte Carlo methods (see also Billingsley, 1995).
-
-5. Conclusion
-
-In summary, by extending the framework of Markov chains—characterized by the Markov property, established convergence criteria, small set theory, and drift conditions—to a hidden state setting, Hidden Markov Models provide a robust foundation for dealing with complex sequence data where the underlying process is not directly observable.
-This extension leads to rich algorithmic and inferential techniques that have been rigorously analyzed and widely adopted in several applied fields.
 % ######################## Background and Literature Review #########################
 
+% ######################## Reference ######################################
+\printbibliography
+% ######################## Reference ######################################
+
 \end{document}

### 2025-02-09 Pic/main.py

@@ -1,30 +1,58 @@
+import itertools
 import numpy as np
-import matplotlib.pyplot as plt
 
-# Generate x values
-x = np.linspace(0, 15, 9)  # Values from 0 to 8 (inclusive)
+# Define HMM parameters
+hidden_states = ["Sunny", "Rainy"]
+observations = ["Walk", "Shop", "Clean"]
 
-# Generate y values (exponential growth)
-y1 = 0.5 ** x  # Exponential function
-y2 = 2 ** x  # Exponential function
+# Initial probabilities (π)
+initial_probs = {"Sunny": 0.6, "Rainy": 0.4}
 
-# Plot the graph
-fig = plt.figure()
-plt.figure(figsize=(8,4))
+# Transition probabilities (P)
+transition_probs = {
+    "Sunny": {"Sunny": 0.7, "Rainy": 0.3},
+    "Rainy": {"Sunny": 0.4, "Rainy": 0.6},
+}
 
-plt.subplot(211)
-plt.plot(x, y1, marker='o', linestyle='-', color='b')
-plt.title("Exponential Grow")
-plt.xlabel("x")
-plt.ylabel("f(x)")
+# Emission probabilities (B)
+emission_probs = {
+    "Sunny": {"Walk": 0.6, "Shop": 0.3, "Clean": 0.1},
+    "Rainy": {"Walk": 0.1, "Shop": 0.4, "Clean": 0.5},
+}
 
-plt.subplot(212)
-plt.plot(x, y2, marker='o', linestyle='-', color='r')
-plt.xlabel("x")
-plt.ylabel("f(x)")
+# Observation sequence
+observation_sequence = ["Walk", "Walk", "Shop"]
 
-# Save the figure as a PNG file
-plt.savefig("exponential_growth.png", dpi=300)  # Save with high resolution (300 DPI)
+# Generate all possible hidden state sequences
+num_obs = len(observation_sequence)
+all_hidden_sequences = list(itertools.product(hidden_states, repeat=num_obs))
 
-# Show the plot
-plt.show()
+# Function to calculate joint probability P(X, Y)
+def calculate_joint_probability(hidden_sequence, observation_sequence):
+    joint_prob = initial_probs[hidden_sequence[0]]  # Start with P(X1)
+    joint_prob *= emission_probs[hidden_sequence[0]][observation_sequence[0]]  # P(Y1 | X1)
+    
+    # Iterate through the sequence
+    for t in range(1, len(hidden_sequence)):
[…]

### 2025-02-16 paper2/main.tex
paper update

@@
 \newpage
+% ######################## Background and Literature Review #########################
 
-% ######################## Background and Literature Review #########################
+% ######################## Project Goals and Objectives ######################################
+\subsection{Project Goals and Objectives}
+
+% Project Goals
+\subsubsection{Project Goals}
+The primary goal of this report is to deeply review on RNN and how it can apply in natural language processing (NLP) tasks.
+And by comparing the performance of different models.
+this report aims to provide a detailed comparative analysis of the following RNN-based models:
+\begin{itemize}
+    \item Vanilla RNN
+    \item Long Short-Term Memory (LSTM)
+    \item Gated Recurrent Unit (GRU)
+\end{itemize}
+
+% Objectives
+\subsubsection{Objectives}
+\begin{enumerate}
+    \item Implement Existing RNN Architectures\\
+    Develop models using state-of-the-art deep learning framework, PyTorch, to implement each of the above RNNs model.
+This includes setting up the network layers, loss functions, and appropriate optimization methods.
+    \item Apply the Models on a Benchmark NLP Dataset\\
+    Utilize a selected NLP dataset to train and evaluate the performance of each RNN-based model.
+The dataset will be preprocessed (e.g., tokenization, padding, embedding initialization) to suit sequential data modeling.
+    \item Compare Model Performances\\
+    Compare different RNN-based models on metrics relevant to the target NLP taskquantitatively and qualitatively. Accuracy, F1, training time scores be used.
+    \item Analyze the Impact of Architectural Differences
+    Identify and discuss the strengths and limitations of each model architecture through both experimental results and theoretical insights.
+For example, compare how LSTM and GRU architectures mitigate vanishing gradients and exploding gradients using gating mechanisms relative to a vanilla RNN.
+\end{enumerate}
+
+\newpage
+% ######################## Project Goals and Objectives ######################################
+
+% ######################## Research Plan / Methodology ######################################
+\subsection{Research Plan / Methodology}
+
+% ######################## Research Plan / Methodology ######################################
 

### 2025-02-16 Pic/main.py

@@ -1,58 +1,0 @@
-import itertools
-import numpy as np
-
-# Define HMM parameters
-hidden_states = ["Sunny", "Rainy"]
-observations = ["Walk", "Shop", "Clean"]
-
-# Initial probabilities (π)
-initial_probs = {"Sunny": 0.6, "Rainy": 0.4}
-
-# Transition probabilities (P)
-transition_probs = {
-    "Sunny": {"Sunny": 0.7, "Rainy": 0.3},
-    "Rainy": {"Sunny": 0.4, "Rainy": 0.6},
-}
-
-# Emission probabilities (B)
-emission_probs = {
-    "Sunny": {"Walk": 0.6, "Shop": 0.3, "Clean": 0.1},
-    "Rainy": {"Walk": 0.1, "Shop": 0.4, "Clean": 0.5},
-}
-
-# Observation sequence
-observation_sequence = ["Walk", "Walk", "Shop"]
-
-# Generate all possible hidden state sequences
-num_obs = len(observation_sequence)
-all_hidden_sequences = list(itertools.product(hidden_states, repeat=num_obs))
-
-# Function to calculate joint probability P(X, Y)
-def calculate_joint_probability(hidden_sequence, observation_sequence):
-    joint_prob = initial_probs[hidden_sequence[0]]  # Start with P(X1)
-    joint_prob *= emission_probs[hidden_sequence[0]][observation_sequence[0]]  # P(Y1 | X1)
-    
-    # Iterate through the sequence
-    for t in range(1, len(hidden_sequence)):
-        joint_prob *= transition_probs[hidden_sequence[t - 1]][hidden_sequence[t]]  # P(Xt | Xt-1)
-        joint_prob *= emission_probs[hidden_sequence[t]][observation_sequence[t]]  # P(Yt | Xt)
-    
-    return joint_prob
-
-# Calculate joint probabilities for all hidden state sequences
-results = []
-for hidden_sequence in all_hidden_sequences:
-    joint_prob = calculate_joint_probability(hidden_sequence, observation_sequence)
-    results.append((hidden_sequence, joint_prob))
-
-# Find the most likely hidden sequence
-most_likely_sequence = max(results, key=lambda x: x[1])
-
-# Display results
-print(f"Observation Sequence: {observation_sequence}\n")
-print("All Possible Hidden State Sequences and Their Joint Probabilities:")
-for seq, prob in results:
[…]

### 2025-02-16 paper2/main.tex
delete conda file

@@
 
+To achieve the project goals and objectives, the research will follow a structured approach combining theoretical understanding with empirical evaluation:
+
+1. **Literature Review:**
+   - **Background Study:**  
+     Conduct a detailed review of seminal and recent works on RNNs, LSTMs, GRUs, and RNN encoder-decoder architectures in NLP. Key references include:
+     - Hochreiter \& Schmidhuber (1997) on LSTM.
+     - Cho et al. (2014) and Chung et al. (2014) on GRU.
+     - Recent surveys on the application of RNNs in NLP.
+   - **Problem Context:**  
+     Understand current best practices, challenges (e.g., vanishing/exploding gradients, sequence length handling), and evaluation standards in NLP.
+
+2. **Data Collection and Preprocessing:**
+   - **Dataset Selection:**  
+     Choose a benchmark NLP dataset appropriate to the target task (e.g., language modeling, text classification, or machine translation).
+   - **Preprocessing Steps:**  
+     Perform text normalization, tokenization, vocabulary building, and sequence padding. Optional steps may include initializing word embeddings (either pre-trained or learned from scratch).
+
+3. **Model Development and Experimental Implementation:**
+   - **Model Architecture Design:**  
+     Implement the different RNN-based models:
+     - **Vanilla RNN:**  
+       Define the recurrence as  
+       \[
+       h_t = \sigma(W_{hx} x_t + W_{hh} h_{t-1} + b_h), \quad y_t = f(W_{yh} h_t + b_y),
+       \]
+       where \(x_t\) is the input vector at time \(t\), \(h_t\) is the hidden state, and \(y_t\) is the output.
+     - **LSTM:**  
+       Set up the LSTM cell with input, forget, and output gates to control information flow.
+     - **GRU:**  
+       Implement the GRU cell, which uses update and reset gates as a simplified alternative.
+     - **RNN Encoder-Decoder:**  
+       Develop an encoder to summarize the input sequence and a decoder to generate the target sequence.
+   - **Framework and Tools:**  
+     Develop the models using PyTorch or TensorFlow. Leverage available libraries for reproducibility and efficiency.
+
+4. **Training, Hyperparameter Tuning, and Evaluation:**
+   - **Training Protocol:**  
+     Use similar training configurations (e.g., optimizer choice, learning rate schedule, batch size) for all models to ensure a fair comparison.
+   - **Hyperparameter Optimization:**  
+     Conduct grid search or Bayesian optimization for parameters such as hidden unit size, dropout rates, and learning rates.
+   - **Evaluation Metrics:**  
+     Define metrics based on the nature of the task (e.g., classification accuracy, F1 score, perplexity). Monitor convergence behavior and computational resource usage.
+   - **Cross-Validation and Benchmarking:**  
+     Use cross-validation and hold-out test sets to ensure robustness of performance assessments.
+
+5. **Result Analysis and Reporting:**
+   - **Comparative Analysis:**  
+     Analyze and interpret the performance differences among the RNN, LSTM, GRU, and RNN encoder-decoder models. Assess the trade-offs in terms of processing speed, accuracy, and stability.
+   - **Discussion of Findings:**  
+     Provide a discussion relative to theoretical expectations (e.g., how and why LSTMs may outperform vanilla RNNs in handling long-term dependencies).
+   - **Documentation:**  
+     Compile the findings into a detailed technical report, including visualizations of loss curves, performance metrics over training epochs, and error analysis.
+
+6. **Dissemination:**
+   - **Presentation and Peer Review:**  
+     Present the results at seminars and workshops to receive feedback.
+   - **Final Reporting:**  
+     Prepare a final research report and, if applicable, a manuscript for publication in a relevant conference or journal.
+
 % ######################## Reference ######################################

### 2025-02-17 paper2/main.tex
paper update

@@
 
-\newpage
 % ######################## Introduction ####################################
@@
 % ######################## Introduction ####################################
-\newpage
+
 % ######################## Background and Literature Review #########################
@@
 This stimulated the building of more refined architectures intended to side step these obstacles.
-
-\newpage
 
@@
 
-\newpage
 \begin{longtable}{|c|c|c|}
@@
 
-\newpage
 % Actication function
@@
 % Hyperbolic tangent activation function
-\newpage
 \textbf{Hyperbolic tangent activation function}
@@
 % Gradient vanishing and gradient exploring
-\newpage
 \subsubsection{Gradient vanishing and gradient exploring}
 When training the RNN, BPTT was used to update the weight matrix.
-As the number of time steps increase, the problem of gradient instability of often encountered, and this problem is gradient vanishing and gradient explored \parencite{bengio1994learning}.
+As the number of time steps increase, the problem of gradient instability of often encountered, and this problem is gradient vanishing and gradient exploring \parencite{bengio1994learning}.
 \\[2ex]
@@
 
-\newpage
 % Long short-term memory (LSTM)
@@
 
-\newpage
 % Gated Recurrent Unit (GRU)
@@
 Total parameters: $3 \times ( n + m + 1 ) \times m$
-\newpage
 
@@
 
-\newpage
 \subsubsection{Hidden Markov Model}
@@
 \end{equation}
-
-\newpage
 
@@
 
-\newpage
-
 % Word representation 
@@
 \textbf{Continuous Bag of Words model}
+\\[1ex]
 CBOW operates by taking context words around a target word, it is flexible to adjust the window size which mean that we can control how many words around our target word.
@@
 
-\newpage
 % ######################## Background and Literature Review #########################
@@
 % ######################## Project Goals and Objectives ######################################
-\subsection{Project Goals and Objectives}
+\section{Project Goals and Objectives}
 
 % Project Goals
-\subsubsection{Project Goals}
+\subsection{Project Goals}
 The primary goal of this report is to deeply review on RNN and how it can apply in natural language processing (NLP) tasks.
@@
 % Objectives
-\subsubsection{Objectives}
-\begin{enumerate}
-    \item Implement Existing RNN Architectures\\
+\subsection{Objectives}
+\subsubsection{Implement Existing RNN Architectures}
 Develop models using state-of-the-art deep learning framework, PyTorch, to implement each of the above RNNs model.
 This includes setting up the network layers, loss functions, and appropriate optimization methods.
-    \item Apply the Models on a Benchmark NLP Dataset\\
+\subsubsection{Apply the Models on a Benchmark NLP Dataset}
 Utilize a selected NLP dataset to train and evaluate the performance of each RNN-based model.
 The dataset will be preprocessed (e.g., tokenization, padding, embedding initialization) to suit sequential data modeling.
-    \item Compare Model Performances\\
+\subsubsection{Compare Model Performances}
 Compare different RNN-based models on metrics relevant to the target NLP taskquantitatively and qualitatively. Accuracy, F1, training time scores be used.
-    \item Analyze the Impact of Architectural Differences
+\subsubsection{Analyze the Impact of Architectural Differences}
 Identify and discuss the strengths and limitations of each model architecture through both experimental results and theoretical insights.
 For example, compare how LSTM and GRU architectures mitigate vanishing gradients and exploding gradients using gating mechanisms relative to a vanilla RNN.
-\end{enumerate}
 
-\newpage
 % ######################## Project Goals and Objectives ######################################
@@
 % ######################## Research Plan / Methodology ######################################
-\subsection{Research Plan / Methodology}
+\section{Research Plan / Methodology}
+To achieve the project goals and objectives, this report will follow the step as below:
+\subsection{Literature Review}
+Since this report require some knowledge related on RNN and NLP.
+So it is necessary to review the foundemental of those area.
+Different architecture of RNNs-based model for instance, vanilla RNN, LSTM, GRU.
+And understand the limitation of RNN, vanishing/exploding gradients.
+Also, it is essential to understand natural language processing (NLP).
+\subsection{Data Collection and Preprocessing}
+In this report, ACL-IMDB \parencite{maas-EtAl:2011:ACL-HLT2011} was experiment.
+This is a dataset for binary sentiment classification containing substantially more data than previous benchmark datasets.
+This dataset provide a set of 25,000 highly polar movie reviews for training, and 25,000 for testing.
+I preporcess the dataset by text cleaning and tokenization.
+\subsection{Model Architecture}
 

### 2025-02-17 paper2/main.tex
window paper update

@@
 \subsection{Model Architecture}
+In this section, i will describe the model architecture, i will list out all of the model i have used and the detail information of those models.
+For mathematical notation, i will define all of the variables precisely and clearly.
+And because this project mainly focusing on natural language processing task.
+So that the classification model consists of an embedding layer, this is just like a look-up table to convert token to vector.
+First, the sequence of words $(w_1, w_2, \cdots, w_T)$ are passed through an embedding layer, which is the look-up table to convert those words in vector format.
+In other words, the look-up table map the word in a high-demensional space $(v_1, v_2, \cdots, v_T)$, where $v_t \in \mathbb{R}^d$.
+And the dimension of the vector space is depends on how we define the hidden size of the model.
 

### 2025-02-19 paper2/main.tex
paper update

@@
 And the dimension of the vector space is depends on how we define the hidden size of the model.
+Next, the forward of RNN based model will processes these word vectors $v_t$ in the forward directions, updating corresponding hidden states at each time step $t$.
+\begin{equation}
+    h_t = f(W_{ih}^Tv_t+W_{hh}h_{t-1}+b_h)
+\end{equation}
+where $W_{ih}$ is input to hidden matrix, $v_t$ is the word vector at time step $t$, $W_{hh}$ is hidden to hidden matrix.
+$h_t$ is hidden state at time $t$.
+$b_h$ is biased term for hidden state.
+For the output at time $t$, we will pass the hidden state output to an activation function defined as below.
+\begin{equation}
+    y_t = f(W_{ho}^Th_{t}+b_y)
+\end{equation}
+where $y_t$ is the output at time step $t$ and $W_{ho}$ is the hidden to output matrix and $b_y$ is biased term for output. And $f(\cdot)$ is an activation function.
 
 % ######################## Research Plan / Methodology ######################################
-
+\newpage
 To achieve the project goals and objectives, the research will follow a structured approach combining theoretical understanding with empirical evaluation:

### 2025-02-20 paper2/main.tex
paper update

@@
 \usepackage{algorithmic}
+\usepackage{listings} % For python code 
 \usepackage[style=apa, sorting=nyt, backend=biber]{biblatex} % Use biber as the backend
@@
 I preporcess the dataset by text cleaning and tokenization.
+\begin{enumerate}
+    \item Lowercasing:\\
+    All text was converted to lowercase to ensure consistency and avoid treadting words like "Good" and "good" as different tokens.
+    \item Handing Abbreviations:
+    I hard code a dictionary in python, and marjority of common abbreviations were included and replaced with their full forms. For example:
+    \begin{itemize}
+        \item "u" -$>$ "you"
+        \item "pls" -$>$ "please"
+    \end{itemize}
+    \item Removing HTML Tags:
+    HTML tags were removed using regular expression 
+    \begin{lstlisting}[language=Python] 
+        import re
+        re.sub(r'<.*?>', '', sentence) 
+    \end{lstlisting}
+    where variable "sentence" is movie review. 
+    \item Removing Special Characters
+    Non-alphanumeric characters (e.g. punctuation) were removed using the regular expression
+    \begin{lstlisting}[language=Python]
+        import re
+        re.sub(r'[^a-zA-Z0-9\s]', '', sentence)
+    \end{lstlisting}
+    \item Removing Extra Whitespace
+    Leading and trailing whitespaces were stripped. 
+    \item Removing Step Words
+
+\end{enumerate}
 \subsection{Model Architecture}
@@
 Next, the forward of RNN based model will processes these word vectors $v_t$ in the forward directions, updating corresponding hidden states at each time step $t$.
+
+\subsubsection{Model 1 (Vanilla RNN)}
+This is the vanilla RNN model, with 1 layer. We just use the last output to calculate the loss function and do the backpropagation.
 \begin{equation}
@@
 \begin{equation}
-    y_t = f(W_{ho}^Th_{t}+b_y)
+    \hat{y}_t = f(W_{ho}^Th_{t}+b_y)
 \end{equation}
-where $y_t$ is the output at time step $t$ and $W_{ho}$ is the hidden to output matrix and $b_y$ is biased term for output. And $f(\cdot)$ is an activation function.
+where $y_t$ is the output at time step $t$ and $W_{ho}$ is the hidden to output matrix and $b_y$ is biased term for output.
+And $f(\cdot)$ is an activation function.
+Then we calculate the loss function by the last output.
+\begin{equation}
+    Loss = \mathcal{L}(\hat{y}, y)
+\end{equation}
+
+\subsubsection{Model 2 (Long short-term memory)}
+For the LSTM model, I use 1 layer, and the calculation step was already discuss.
+
+\subsubsection{Model 3 (Gated recurrent unit)}
+For the GRU model, I use 1 layer, and the calculation step was already discuss.
+
+\subsection{Training}
+\subsubsection{Training Protocol}
+In order to ensure a fair comparison, i will use the similar training configurations (e.g., optimizer choice, learning rate schedule, batch size).
+
+4. **Training, Hyperparameter Tuning, and Evaluation:**
+   - **Training Protocol:**  
+     Use similar training configurations (e.g., optimizer choice, learning rate schedule, batch size) for all models to ensure a fair comparison.
+   - **Hyperparameter Optimization:**  
+     Conduct grid search or Bayesian optimization for parameters such as hidden unit size, dropout rates, and learning rates.
+   - **Evaluation Metrics:**  
+     Define metrics based on the nature of the task (e.g., classification accuracy, F1 score, perplexity). Monitor convergence behavior and computational resource usage.
+   - **Cross-Validation and Benchmarking:**  
+     Use cross-validation and hold-out test sets to ensure robustness of performance assessments.
 

### 2025-02-20 paper2/main.tex
paper update

@@
 \usepackage{listings} % For python code 
+\usepackage{tikz} 
+\usetikzlibrary{positioning, arrows.meta}
 \usepackage[style=apa, sorting=nyt, backend=biber]{biblatex} % Use biber as the backend
@@
 \subsection{Data Collection and Preprocessing}
+\subsubsection{Data Collection}
 In this report, ACL-IMDB \parencite{maas-EtAl:2011:ACL-HLT2011} was experiment.
@@
 I preporcess the dataset by text cleaning and tokenization.
+\subsubsection{Data Preprocessing}
 \begin{enumerate}
@@
     All text was converted to lowercase to ensure consistency and avoid treadting words like "Good" and "good" as different tokens.
-    \item Handing Abbreviations:
+    \item Handing Abbreviations:\\
     I hard code a dictionary in python, and marjority of common abbreviations were included and replaced with their full forms. For example:
@@
     \end{itemize}
-    \item Removing HTML Tags:
+    \item Removing HTML Tags:\\
     HTML tags were removed using regular expression 
@@
     where variable "sentence" is movie review. 
-    \item Removing Special Characters
+    \item Removing Special Characters\\
     Non-alphanumeric characters (e.g. punctuation) were removed using the regular expression
@@
     \end{lstlisting}
-    \item Removing Extra Whitespace
+    \item Removing Extra Whitespace\\
     Leading and trailing whitespaces were stripped. 
-    \item Removing Step Words
-
+    \item Removing Stop Words\\
+    Stop words (e.g., "is", "the", "and") were removed to reduce noise in the text. The NLTK library's stop word list was used. Words not in the stop word list were retained.
+    \item Tokenization\\
+    The cleaned text was split into individual tokens, I use white space and the criterion to split the word.
+\\[1ex]
+Once the text was preprocessed, a vocabulary was constructed from the dataset.
+The vocabulary mapping each word to a unique index.
+Words not found in the vocabulary are replaced with an "unknown" token.
+I reserved index 0 as unknown token.
+So that if the word can not map to correspoinding index.
+Then the token should be 0.
+This is a good way to handle Out-of-vocabulary words \parencite{mikolov2013efficient}.
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|}
+        \hline
+        \textbf{Word} & \textbf{Index} \\ \hline
+        \texttt{<UNK>} & 0 \\ \hline
+        acting & 1 \\ \hline
+        good & 2 \\ \hline
+        movie & 3 \\ \hline
+        amazing & 4 \\ \hline
+        performance & 5 \\ \hline
+        bad & 6 \\ \hline
+        terrible & 7 \\ \hline
+        excellent & 8 \\ \hline
+    \end{tabular}
+    \caption{Word to Index Mapping Table}
+    \label{tab:word_index}
+\end{table}
 \end{enumerate}
@@
 First, the sequence of words $(w_1, w_2, \cdots, w_T)$ are passed through an embedding layer, which is the look-up table to convert those words in vector format.
-In other words, the look-up table map the word in a high-demensional space $(v_1, v_2, \cdots, v_T)$, where $v_t \in \mathbb{R}^d$.
+In other words, the look-up table map the word in a high-demensional space $(v_1, v_2, \cdots, v_T)$ (Table 5), where $v_t \in \mathbb{R}^d$.
 And the dimension of the vector space is depends on how we define the hidden size of the model.
@@
 
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \textbf{Word} & \textbf{Index} & \textbf{Embedding Vector} \\ \hline
+        \texttt{<UNK>} & 0 & $\begin{bmatrix} 0.01 & -0.02 \cdots 0.03 & 0.04 \end{bmatrix}$ \\ \hline
+        acting & 1 & $\begin{bmatrix} 0.12 & 0.23 \cdots -0.11 & 0.05 \end{bmatrix}$ \\ \hline
+        good & 2 & $\begin{bmatrix} 0.03 & -0.15 \cdots 0.08 & 0.07 \end{bmatrix}$ \\ \hline
+        movie & 3 & $\begin{bmatrix} 0.10 & 0.05 \cdots -0.09 & -0.01 \end{bmatrix}$ \\ \hline
+        amazing & 4 & $\begin{bmatrix} 0.25 & -0.12 \cdots 0.18 & 0.30 \end{bmatrix}$ \\ \hline
+        performance & 5 & $\begin{bmatrix} -0.10 & -0.20 \cdots 0.15 & -0.05 \end{bmatrix}$ \\ \hline
+        bad & 6 & $\begin{bmatrix} 0.40 & 0.35 \cdots -0.10 & 0.50 \end{bmatrix}$ \\ \hline
+        terrible & 7 & $\begin{bmatrix} 0.30 & 0.25 \cdots -0.11 & 0.40 \end{bmatrix}$ \\ \hline
+        excellent & 8 & $\begin{bmatrix} 0.32 & 0.38 \cdots -0.90 & 0.70 \end{bmatrix}$ \\ \hline
+    \end{tabular}
+    \caption{Word to Index Matching with Lookup Table Embeddings}
+    \label{tab:word_lookup}
+\end{table}
+
 \subsubsection{Model 1 (Vanilla RNN)}
 This is the vanilla RNN model, with 1 layer. We just use the last output to calculate the loss function and do the backpropagation.
+\begin{center} % Center the graph on the page
+    \begin{tikzpicture}[
+        node distance=1.5cm and 2cm, % Distance between nodes
+        every node/.style={draw, minimum size=1cm, font=\small, align=center}, % Node style
+        arrow/.style={-Latex, thick} % Arrow style
+    ]
+
+    % Input nodes
+    \node (x1) [circle, fill=blue!10] {$x_1$};
+    \node (x2) [circle, fill=blue!10, right=of x1] {$x_2$};
+    \node (x3) [circle, fill=blue!10, right=of x2] {$x_3$};
+
+    % Hidden nodes
+    \node (h1) [circle, fill=green!10, above=of x1] {$h_1$};
+    \node (h2) [circle, fill=green!10, above=of x2] {$h_2$};
+    \node (h3) [circle, fill=green!10, above=of x3] {$h_3$};
+
+    % Output nodes
+    \node (y1) [circle, fill=red!10, above=of h1] {$y_1$};
+    \node (y2) [circle, fill=red!10, above=of h2] {$y_2$};
+    \node (y3) [circle, fill=red!10, above=of h3] {$y_3$};
+
+    % Horizontal connections (input to hidden)
+    \draw[arrow] (x1) -- (h1);
+    \draw[arrow] (x2) -- (h2);
+    \draw[arrow] (x3) -- (h3);
+
+    % Horizontal connections (hidden to output)
+    \draw[arrow] (h1) -- (y1);
+    \draw[arrow] (h2) -- (y2);
+    \draw[arrow] (h3) -- (y3);
+
+    % Recurrent connections
+    \draw[arrow] (h1.east) to[out=30, in=150] (h2.west);
+    \draw[arrow] (h2.east) to[out=30, in=150] (h3.west);
+
+    % Labels
+    \node[above=0.3cm of y2, font=\bfseries] {Recurrent Neural Network (Unfolded)};
+    \node[below=0.5cm of x2, font=\itshape] {Time Steps};
+
+    \end{tikzpicture}
+\end{center}
+
 \begin{equation}

### 2025-02-24 paper2/main.tex
paper update

@@
     % Input nodes
-    \node (x1) [circle, fill=blue!10] {$x_1$};
-    \node (x2) [circle, fill=blue!10, right=of x1] {$x_2$};
-    \node (x3) [circle, fill=blue!10, right=of x2] {$x_3$};
+    \node (v1) [circle, fill=blue!10] {$v_1$};
+    \node (v2) [circle, fill=blue!10, right=of v1] {$v_2$};
+    \node (v3) [circle, fill=blue!10, right=of v2] {$v_3$};
 
     % Hidden nodes
-    \node (h1) [circle, fill=green!10, above=of x1] {$h_1$};
-    \node (h2) [circle, fill=green!10, above=of x2] {$h_2$};
-    \node (h3) [circle, fill=green!10, above=of x3] {$h_3$};
+    \node (h1) [circle, fill=green!10, above=of v1] {$h_1$};
+    \node (h2) [circle, fill=green!10, above=of v2] {$h_2$};
+    \node (h3) [circle, fill=green!10, above=of v3] {$h_3$};
 
@@
     % Horizontal connections (input to hidden)
-    \draw[arrow] (x1) -- (h1);
-    \draw[arrow] (x2) -- (h2);
-    \draw[arrow] (x3) -- (h3);
+    \draw[arrow] (v1) -- (h1);
+    \draw[arrow] (v2) -- (h2);
+    \draw[arrow] (v3) -- (h3);
 
@@
     \node[above=0.3cm of y2, font=\bfseries] {Recurrent Neural Network (Unfolded)};
-    \node[below=0.5cm of x2, font=\itshape] {Time Steps};
+    \node[below=0.5cm of v2, font=\itshape] {Time Steps};
 
@@
 \subsubsection{Model 2 (Long short-term memory)}
-For the LSTM model, I use 1 layer, and the calculation step was already discuss.
+For the LSTM model, I use 1 layer:
+\begin{figure}[H]
+    \centering
+    \fbox{
+    \resizebox{0.95\textwidth}{!}{
+    \begin{tikzpicture}[
+    font=\large,
+        >=LaTeX,
+        cell/.style={% For the main box
+        rectangle, 
+        rounded corners=5mm, 
+        draw,
+        very thick,
+        },
+        operator/.style={
+        circle,
+        draw,
+        inner sep=-0.5pt,
+        minimum width =.3cm,
+        minimum height =.3cm,
+        },
+        function/.style={
+        ellipse,
+        draw,
+        inner sep=1pt
+        },
+        ct/.style={
+        circle,
+        draw,
+        line width = .75pt,
+        minimum width=1cm,
+        inner sep=1pt,
+        },
+        gt1/.style={
+        rectangle,
+        draw,
+        fill=green, 
+        minimum width=4mm,
+        minimum height=3mm,
+        inner sep=1pt
+        },
+        gt2/.style={
+        rectangle,
+        draw,
+        fill=red, 
+        minimum width=4mm,
+        minimum height=3mm,
+        inner sep=1pt
+        },
+        gt3/.style={
+        rectangle,
+        draw,
+        fill=yellow, 
+        minimum width=4mm,
+        minimum height=3mm,
+        inner sep=1pt
+        },
+        ArrowC1/.style={% Arrows with rounded corners
+        rounded corners=.25cm,
+        thick,
+        },
+        ArrowC2/.style={% Arrows with big rounded corners
+        rounded corners=.5cm,
+        thick,
+        },
+        cella/.pic={
+            \node [cell, minimum height =4cm, minimum width=6cm] at (0,0){};
+            \node [gt1] (ibox1a) at (-2,-0.75) {$\sigma$};
+            \node [gt2] (ibox2a) at (-1.5,-0.75) {$\sigma$};
+            \node [gt2, minimum width=1cm] (ibox3a) at (-0.5,-0.75) {Tanh};
+            \node [gt3] (ibox4a) at (0.5,-0.75) {$\sigma$};
+            \node [operator] (mux1a) at (-2,1.5) {$\times$};
+            \node [operator] (adda) at (-0.5,1.5) {+};
+            \node [operator] (mux2a) at (-0.5,0) {$\times$};
+            \node [gt3] (mux3a) at (1.5,0) {$\times$};
+            \node [gt3] (funca) at (1.5,0.75) {Tanh};
+            \node[] (ca) at (-4,1.5) {$C_{t-2}$};
+            \node[] (ha) at (-4,-1.5) {$y_{t-2}$};
+            \node[] (xa) at (-2.5,-3) {$x_{t-1}$};
+            \node[] (c2a) at (4,1.5) {};
+            \node[] () at (3.5, 2) {$C_{t-1}$};
+            \node[] (h2a) at (4,-1.5) {};
+            \node[] (x2a) at (2.5,3) {$y_{t-1}$};
+            \node[] () at (3.5, -2) {$y_{t-1}$};
+            % \draw [->, ArrowC1] (c) -- (mux1) -- (add1) -- (c2);
+            \draw [->, ArrowC1] (ca) -- (mux1a);
+            \draw [->, ArrowC1] (mux1a) -- (adda);
+            \draw [->, ArrowC1] (adda) -- (c2a);
+            \draw [ArrowC2] (ha) -| (ibox4a);
+            \draw [ArrowC1] (ha -| ibox1a)++(-0.5,0) -| (ibox1a); 
+            \draw [ArrowC1] (ha -| ibox2a)++(-0.5,0) -| (ibox2a);
+            \draw [ArrowC1] (ha -| ibox3a)++(-0.5,0) -| (ibox3a);
+            \draw [ArrowC1] (xa) -- (xa |- ha)-| (ibox3a);
+            \draw [->, ArrowC2] (ibox1a) -- (mux1a);
+            \draw [->, ArrowC2] (ibox2a) |- (mux2a);
+            \draw [->, ArrowC2] (ibox3a) -- (mux2a);
+            \draw [->, ArrowC2] (ibox4a) |- (mux3a);
+            \draw [->, ArrowC2] (mux2a) -- (adda);
+            \draw [->, ArrowC1] (adda -| funca)++(-0.5,0) -| (funca);
+            \draw [->, ArrowC2] (funca) -- (mux3a);
+            \draw [->, ArrowC2] (mux3a) |- (h2a);
+            \draw (c2a -| x2a) ++(0,-0.1) coordinate (i1a);
+            \draw [-, ArrowC2] (h2a -| x2a)++(-0.5,0) -| (i1a);
+            \draw [->, ArrowC2] (i1a)++(0,0.2) -- (x2a);
+        },
+        cellb/.pic={
+            \node [cell, minimum height =4cm, minimum width=6cm] at (0,0){};
+            \node [gt1] (ibox1b) at (-2,-0.75) {$\sigma$};
+            \node [gt2] (ibox2b) at (-1.5,-0.75) {$\sigma$};
+            \node [gt2, minimum width=1cm] (ibox3b) at (-0.5,-0.75) {Tanh};
+            \node [gt3] (ibox4b) at (0.5,-0.75) {$\sigma$};
+            \node [operator] (mux1b) at (-2,1.5) {$\times$};
+            \node [operator] (add1b) at (-0.5,1.5) {+};
+            \node [operator] (mux2b) at (-0.5,0) {$\times$};
+            \node [gt3] (mux3b) at (1.5,0) {$\times$};
+            \node [gt3] (func1b) at (1.5,0.75) {Tanh};
+            \node[] (cb) at (-4,1.5) {};
+            \node[] (hb) at (-4,-1.5) {};
+            \node[] (xb) at (-2.5,-3) {$x_{t}$};
+            \node[] (c2b) at (4,1.5) {};
+            \node[] () at (3.5,2) {$C_{t}$};
+            \node[] (h2b) at (4,-1.5) {};
+            \node[] (x2b) at (2.5,3) {$y_{t}$};
+            \node[] () at (3.5, -2) {$y_{t}$};
+            % \draw [->, ArrowC1] (c) -- (mux1) -- (add1) -- (c2);
+            \draw [->, ArrowC1] (cb) -- (mux1b);
+            \draw [->, ArrowC1] (mux1b) -- (add1b);
+            \draw [->, ArrowC1] (add1b) -- (c2b);
+            \draw [ArrowC2] (hb) -| (ibox4b);
+            \draw [ArrowC1] (hb -| ibox1b)++(-0.5,0) -| (ibox1b); 
+            \draw [ArrowC1] (hb -| ibox2b)++(-0.5,0) -| (ibox2b);
+            % \draw [ArrowC1] (h -| ibox3)++(-0.5,0) -| (ibox3);
+            \draw [ArrowC1] (xb) -- (xb |- hb)-| (ibox3b);
+            \draw [->, ArrowC2] (ibox1b) -- (mux1b);
+            \draw [->, ArrowC2] (ibox2b) |- (mux2b);
+            \draw [->, ArrowC2] (ibox3b) -- (mux2b);
+            \draw [->, ArrowC2] (ibox4b) |- (mux3b);
+            \draw [->, ArrowC2] (mux2b) -- (add1b);
+            \draw [->, ArrowC1] (add1b -| func1b)++(-0.5,0) -| (func1b);
+            \draw [->, ArrowC2] (func1b) -- (mux3b);
+            \draw [->, ArrowC2] (mux3b) |- (h2b);
+            \draw (c2b -| x2b) ++(0,-0.1) coordinate (i1b);
+            \draw [-, ArrowC2] (h2b -| x2b)++(-0.5,0) -| (i1b);
+            \draw [->, ArrowC2] (i1b)++(0,0.2) -- (x2b);
+        },
+        cellc/.pic={
+            \node [cell, minimum height =4cm, minimum width=6cm] at (0,0){};
+            \node [gt1] (ibox1c) at (-2,-0.75) {$\sigma$};
+            \node [gt2] (ibox2c) at (-1.5,-0.75) {$\sigma$};
+            \node [gt2, minimum width=1cm] (ibox3c) at (-0.5,-0.75) {Tanh};
+            \node [gt3] (ibox4c) at (0.5,-0.75) {$\sigma$};
+            \node [operator] (mux1c) at (-2,1.5) {$\times$};
+            \node [operator] (add1c) at (-0.5,1.5) {+};
+            \node [operator] (mux2c) at (-0.5,0) {$\times$};
+            \node [gt3] (mux3c) at (1.5,0) {$\times$};
+            \node [gt3] (func1c) at (1.5,0.75) {Tanh};
+            \node[] (cc) at (-4,1.5) {};
+            \node[] (hc) at (-4,-1.5) {};
+            \node[] (xc) at (-2.5,-3) {$x_{t+1}$};
+            \node[] (c2c) at (4,1.5) {};
+            \node[] () at (3.5,2) {$C_{t+1}$};
+            \node[] (h2c) at (4,-1.5) {};
+            \node[] (x2c) at (2.5,3) {$y_{t+1}$};
+            \node[] () at (3.5, -2) {$y_{t+1}$};
+            % \draw [->, ArrowC1] (c) -- (mux1) -- (add1) -- (c2);
+            \draw [->, ArrowC1] (cc) -- (mux1c);
+            \draw [->, ArrowC1] (mux1c) -- (add1c);
+            \draw [->, ArrowC1] (add1c) -- (c2c);
+            \draw [ArrowC2] (hc) -| (ibox4c);
+            \draw [ArrowC1] (hc -| ibox1c)++(-0.5,0) -| (ibox1c); 
+            \draw [ArrowC1] (hc -| ibox2c)++(-0.5,0) -| (ibox2c);
+            \draw [ArrowC1] (hc -| ibox3c)++(-0.5,0) -| (ibox3c);
+            \draw [ArrowC1] (xc) -- (xc |- hc)-| (ibox3c);
+            \draw [->, ArrowC2] (ibox1c) -- (mux1c);
+            \draw [->, ArrowC2] (ibox2c) |- (mux2c);
+            \draw [->, ArrowC2] (ibox3c) -- (mux2c);
+            \draw [->, ArrowC2] (ibox4c) |- (mux3c);
+            \draw [->, ArrowC2] (mux2c) -- (add1c);
+            \draw [->, ArrowC1] (add1c -| func1c)++(-0.5,0) -| (func1c);
+            \draw [->, ArrowC2] (func1c) -- (mux3c);
+            \draw [->, ArrowC2] (mux3c) |- (h2c);
+            \draw (c2c -| x2c) ++(0,-0.1) coordinate (i1c);
+            \draw [-, ArrowC2] (h2c -| x2c)++(-0.5,0) -| (i1c);
+            \draw [->, ArrowC2] (i1c)++(0,0.2) -- (x2c);
+        }, 
+        very thick,
+    ]
+    \begin{scope}[xshift=-7cm]
+    \pic {cella};
+    \draw[rounded corners=0.5cm, fill=green!40, opacity=0.8] (-3,-2) rectangle +(6,4);
+    \node[scale=3] {$t-1$};
+    \end{scope}
+    \begin{scope}[xshift=0cm]
+    \draw[rounded corners=0.5cm, fill=green!40, opacity=0.8] (-3,-2) rectangle +(6,4);
+    \pic {cellb};
+    \end{scope}
+    \begin{scope}[xshift=7cm]
+    \pic {cellc};
+    \draw[rounded corners=0.5cm, fill=green!40, opacity=0.8] (-3,-2) rectangle +(6,4);
+    \node[scale=3] {$t+1$};
+    \end{scope}
+    \end{tikzpicture}
+    }
+    }
+\end{figure}
+
+For an LSTM, the overall structure is similar to the Vanilla RNN shown in the diagram, but the hidden state computation is replaced by the LSTM's more complex gating mechanism.
+Using its gating mechanisms to control how information flows through time.
+Let us consider an input sequence $\{v_1, v_2, \dots, v_T\}$, where $v_t$ represents the input vector at time step $t$.
+Here’s how the LSTM processes the input:
+
+\paragraph{Step 1: Initialize Hidden and Cell States}
+At the beginning of the sequence, the LSTM initializes the hidden state $h_0$ and cell state $C_0$. For our model, we use 0 matrix as $h_0$ and $C_0$:
+\begin{eqnarray}
+    h_0 = \mathbf{0}, \quad C_0 = \mathbf{0}
+\end{eqnarray}
+
+\paragraph{Step 2: Process Each Input Vector}
+For each time step $t$, LSTM processes the current input $v_t$ along with the previous hidden state $h_{t-1}$ and cell state $C_{t-1}$ to compute:
+\begin{itemize}
+    \item The \textbf{forget gate} $f_t$, determine which information should be retained in memory cell $C_{t-1}$.
+    \item The \textbf{input gate} $i_t$, controls how much new information from the current time step is allowed to enter the cell.
+    \item The \textbf{cell candidate} $\tilde{C}_t$, represents the new information to be added to the cell state.
+    \item The updated \textbf{cell state} $C_t$, which combines the retained information from $C_{t-1}$ and the new candidate $\tilde{C}_t$.
+    \item The \textbf{output gate} $o_t$, control how much or what information from the cell state should be passed to the next layer or used in predictions.
+    \item The updated \textbf{hidden state} $h_t$, summarizes the information at time $t$.
+\end{itemize}
+
+\paragraph{Step 3: Compute Forget Gate}
+The forget gate $f_t$ is computed as:
+\begin{equation}
+    f_t = \sigma(W_f v_t + U_f h_{t-1} + b_f)
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\sigma$ is the sigmoid activation function.
+    \item $W_f$ is the input to hidden weight matrix for the forget gate.
+    \item $U_f$ is the hidden to hidden weight matrix for the forget gate.
+    \item $b_f$ is the bias term for the forget gate.
+\end{itemize}
+The forget gate determines which parts of the previous cell state $C_{t-1}$ to retain.
+
+\paragraph{Step 4: Compute Input Gate}
+The input gate $i_t$ is computed as:
+\begin{eqnarray}
+    i_t = \sigma(W_i v_t + U_i h_{t-1} + b_i)
+\end{eqnarray}
+
+where:
+\begin{itemize}
+    \item $W_i$, $U_i$, and $b_i$ are the weight matrices and bias for the input gate.
+\end{itemize}
+The input gate determines how much new information to add to the cell state.
+
+\paragraph{Step 5: Compute Cell Candidate}
+The cell candidate $\tilde{C}_t$ is computed as:
+\begin{equation}
+    \tilde{C}_t = \tanh(W_c v_t + U_c h_{t-1} + b_c)
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\tanh$ is the hyperbolic tangent activation function.
+    \item $W_c$, $U_c$, and $b_c$ are the weight matrices and bias for the cell candidate.
+\end{itemize}
+The cell candidate represents potential new information to add to the cell state.
+
+\paragraph{Step 6: Update Cell State}
+The cell state $C_t$ is updated as:
+\begin{equation}
+    C_t = f_t \odot C_{t-1} + i_t \odot \tilde{C}_t
+\end{equation}
+
+where $\odot$ represents element-wise multiplication. The updated cell state combines retained information from $C_{t-1}$ (via $f_t$) and new information from $\tilde{C}_t$ (via $i_t$).
+
+\paragraph{Step 7: Compute Output Gate}
+The output gate $o_t$ is computed as:
+\begin{equation}
+    o_t = \sigma(W_o v_t + U_o h_{t-1} + b_o)
+\end{equation}
+
+where $W_o$, $U_o$, and $b_o$ are the weight matrices and bias for the output gate.
+
+\paragraph{Step 8: Compute Hidden State}
+The hidden state $h_t$ is computed as:
+\begin{equation}
+    h_t = o_t \odot \tanh(C_t)
+\end{equation}
+
+The hidden state summarizes the information at time $t$ and is passed to the next time step.
+
+\paragraph{Step 9: Use Final Hidden State for Output}
+After processing all time steps, the final hidden state $h_T$ is used as the input to a fully connected layer for classification:
+\begin{equation}
+    \hat{y} = f(W_{ho}^T h_T + b_y)    
+\end{equation}
+
+where $W_{ho}$ and $b_y$ are the weight matrix and bias for the output layer, and $f$ is the activation function. Then we calculate the loss function by the last output.
+\begin{equation}
+    Loss = \mathcal{L}(\hat{y}, y)
+\end{equation}
 
 \subsubsection{Model 3 (Gated recurrent unit)}
-For the GRU model, I use 1 layer, and the calculation step was already discuss.
+For the GRU model, I use 1 layer.
+This architecture is similar with LSTM but reduce the among of parameters.
+Let us consider an input sequence $\{v_1, v_2, \dots, v_T\}$, where $v_t$ represents the input vector at time step $t$.
+Here's how the GRU processes the input:
+\paragraph{Step 1: Initialize Hidden State}
+At the beginning of the sequence, the GRU initializes the hidden state $h_0$ to zero:
+\begin{equation}
+    h_0 = \mathbf{0}
+\end{equation}
+
+\paragraph{Step 2: Process Each Input Vector}
+For each time step $t$, the GRU processes the current input $v_t$ along with the previous hidden state $h_{t-1}$ to compute:
+\begin{itemize}
+    \item The \textbf{reset gate} $r_t$, which determines how much of the previous hidden state $h_{t-1}$ to forget.
+    \item The \textbf{update gate} $z_t$, which determines how much of the previous hidden state $h_{t-1}$ to retain.
+    \item The \textbf{candidate hidden state} $\tilde{h}_t$, which represents the new information to be combined with the previous hidden state.
+    \item The updated \textbf{hidden state} $h_t$, which summarizes the information at time $t$.
+\end{itemize}
+
+\paragraph{Step 3: Compute Reset Gate}
+The reset gate $r_t$ is computed as:
+\begin{equation}
+    r_t = \sigma(W_r v_t + U_r h_{t-1} + b_r)
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\sigma$ is the sigmoid activation function.
+    \item $W_r$ is the input to hidden weight matrix for the reset gate.
+    \item $U_r$ is the hidden to hidden weight matrix for the reset gate.
+    \item $b_r$ is the bias term for the reset gate.
+\end{itemize}
+The reset gate determines how much of the previous hidden state $h_{t-1}$ to use when computing the candidate hidden state.
+
+\paragraph{Step 4: Compute Update Gate}
+The update gate $z_t$ is computed as:
+\begin{equation}
+    z_t = \sigma(W_z v_t + U_z h_{t-1} + b_z)
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $W_z$ is the input to hidden weight matrix for the update gate.
+    \item $U_z$ is the hidden to hidden weight matrix for the update gate.
+    \item $b_z$ is the bias term for the update gate.
+\end{itemize}
+The update gate determines how much of the previous hidden state $h_{t-1}$ to retain and how much of the new candidate hidden state to use.
+
+\paragraph{Step 5: Compute Candidate Hidden State}
+The candidate hidden state $\tilde{h}_t$ is computed as:
+\begin{equation}
+    \tilde{h}_t = \tanh(W_h v_t + U_h (r_t \odot h_{t-1}) + b_h)    
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\tanh$ is the hyperbolic tangent activation function.
+    \item $r_t \odot h_{t-1}$: The reset gate modulates the contribution of the previous hidden state $h_{t-1}$.
+    \item $W_h$, $U_h$, and $b_h$ are the weight matrices and bias for the candidate hidden state.
+\end{itemize}
+
+\paragraph{Step 6: Update Hidden State}
+The updated hidden state $h_t$ is computed as:
+\begin{equation}
+    h_t = z_t \odot h_{t-1} + (1 - z_t) \odot \tilde{h}_t    
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $z_t$: The update gate decides how much of the previous hidden state $h_{t-1}$ to retain.
+    \item $(1 - z_t)$: The complement of the update gate decides how much of the candidate hidden state $\tilde{h}_t$ to use.
+\end{itemize}
+
+\paragraph{Step 7: Use Final Hidden State for Output}
+After processing all time steps, the final hidden state $h_T$ is used as the input to a fully connected layer for classification:
+\begin{equation}
+    \hat{y} = f(W_{ho}^T h_T + b_y)    
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $W_{ho}$ and $b_y$ are the weight matrix and bias for the output layer.
+    \item $f$ is the activation function (e.g., sigmoid for binary classification).
+\end{itemize}
+
+Then we calculate the loss function by the last output.
+\begin{equation}
+    Loss = \mathcal{L}(\hat{y}, y)
+\end{equation}
 

### 2025-02-24 paper2/main.tex
paper update

@@
 \usetikzlibrary{positioning, arrows.meta}
+\usetikzlibrary{shapes.geometric, arrows, positioning}
+
+\tikzstyle{block} = [rectangle, draw, text centered, rounded corners, minimum height=2em, minimum width=4em]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+
 \usepackage[style=apa, sorting=nyt, backend=biber]{biblatex} % Use biber as the backend
@@
 Here's how the GRU processes the input:
+
+\begin{figure}[tbh!]
+    \centering   
+    \begin{tikzpicture}[
+        font=\sf \tiny,
+        >=LaTeX,
+        % Styles
+        cell/.style={% For the main box
+            rectangle, 
+            rounded corners=5mm, 
+            draw,
+            very thick,
+            },
+        cell2/.style={% For the main box
+        rectangle, 
+        rounded corners=2mm, 
+        draw,
+        thin,
+        },
+        operator/.style={% For operators like +  and  x
+            circle,
+            draw,
+            inner sep=-0.5pt,
+            minimum height =.2cm,
+            },
+        function/.style={% For functions
+            ellipse,
+            draw,
+            inner sep=1pt
+            },
+        ct/.style={% For external inputs and outputs
+            circle,
+            draw,
+            line width = .75pt,
+            minimum width=1cm,
+            inner sep=0pt,
+            },
+        gt/.style={% For internal inputs
+            rectangle,
+            draw,
+            minimum width=3mm,
+            minimum height=3mm,
+            inner sep=1pt
+            },
+        mylabel/.style={% something new that I have learned
+            font=\tiny\sffamily
+            },
+        ArrowC1/.style={% Arrows with rounded corners
+            rounded corners=.25cm,
+            thick,
+            },
+        ArrowC2/.style={% Arrows with big rounded corners
+            rounded corners=.5cm,
+            thick,
+            },
+        ]
+
+    % Nodes    
+        % Draw the cell boundary: 
+        \node [cell, minimum height=3.5cm, minimum width=3.5cm] at (0,0){} ;
+        \node [dashed, red, cell2, minimum height=3cm, minimum width=0.6cm] at (0.05,0.175){} ;
+        \node [dashed, blue, cell2, minimum height=3cm, minimum width=0.85cm] at (-0.75,0.15){} ;
+
+        % Draw inputs named ibox#
+        \node [gt] (sigma1) at (-0.5,-0.75) {$\sigma$};
+        \node [gt] (sigma2) at (0,-0.75) {$\sigma$};
+        \node [gt, minimum width=0.5cm] (tanh) at (1,-0.75) {Tanh};
+        \node [gt] (oneminus) at (0,0.25) {$1-$};
+
+       % Draw operators named mux# , add# and func#
+        \node [operator] (mux1) at (-1,0.5) {$\times$};
+        \node [operator] (add1) at (1,1.5) {+};
+        \node [operator] (mux2) at (1,0) {$\times$};
+        \node [operator] (mux3) at (0,1.5) {$\times$};
+
+        % Draw External inputs named as basis h,x
+        \node[ct, label={[mylabel]Previous hidden state}, align=center, inner sep=0pt, minimum size=4mm] (h) at (-2.5,1.5) {$h^{t-1}$};
+        \node[ct, label={[mylabel]left:Input}, align=center, inner sep=0pt, minimum size=4mm] (x) at (-1,-2.5) {$x^{t}$};
+
+        % Draw External outputs named as basis h2,y2
+        \node[ct, label={[mylabel]Current hidden state}, align=center, inner sep=0pt, minimum size=4mm] (h2) at (2.25,1.5) {$h^{t}$};
+        \node[ct, label={[mylabel]}, align=center, inner sep=0pt, minimum size=4mm] (y2) [below of= h2] {$y^{t}$};
+
+    % Start connecting all.
+        % Intersections and displacements are used. 
+        % Drawing arrows    
+        \draw [ArrowC1] (h) -- (mux3) -- (add1) -- (h2);
+        \draw [ArrowC1] (add1) coordinate[auto] -|(y2.west);
+
+        % Inputs
+        \draw [ArrowC1] (x) -- coordinate[auto] (link1) (mux1); 
+        \draw [ArrowC1] (x -| sigma2)++(-1,1.25) -| (sigma2);
+        \draw [ArrowC1] (x -| tanh)++(-2,1) -| (tanh);
+        \draw [ArrowC1] (x.north)++(0,1.04) -| (sigma1.south); 
+        \draw [->, ArrowC1] (mux1) ++(0,0.1) |- (mux3) ;
+        \draw [->,ArrowC1] (h.east)++(0.6,0) |- (mux1.west);
+        \draw [->, ArrowC1] (sigma1.north) |- coordinate[auto]  node[pos=0.5] {$r_t$} (mux1.east);
+
+        % Internal
+        \draw [->, ArrowC2] (sigma2) |- (mux2);
+        \draw [->, ArrowC2] (tanh) -- node[pos=0.4, right] {$\widetilde{h}_t$}(mux2);
+        \draw [->, ArrowC2] (oneminus) -- coordinate[auto] node[pos=0.75, right] {$u_t$} (mux3);
+        \draw [->, ArrowC2] (sigma2)++(0,0.25) -| node[pos=0.4, right] {$z_t$} (oneminus);
+        \draw [->, ArrowC2] (mux2) -- (add1);
+        % \draw [ArrowC2] (h.east)++(0.6,0) |- ($(x.north)!0.775!(link1)$);  
+        % \draw[->, ArrowC2] (h.east)++(0.25,0) to [out=0,in=100, looseness=0.85] ($(x.north)!0.2!(h.east)$);
+
+    \end{tikzpicture}
+\end{figure}
+
 \paragraph{Step 1: Initialize Hidden State}
@@
 
-\subsection{Training}
-\subsubsection{Training Protocol}
-In order to ensure a fair comparison, i will use the similar training configurations (e.g., optimizer choice, learning rate schedule, batch size).
-
-4. **Training, Hyperparameter Tuning, and Evaluation:**
-   - **Training Protocol:**  
-     Use similar training configurations (e.g., optimizer choice, learning rate schedule, batch size) for all models to ensure a fair comparison.
-   - **Hyperparameter Optimization:**  
-     Conduct grid search or Bayesian optimization for parameters such as hidden unit size, dropout rates, and learning rates.
-   - **Evaluation Metrics:**  
-     Define metrics based on the nature of the task (e.g., classification accuracy, F1 score, perplexity). Monitor convergence behavior and computational resource usage.
-   - **Cross-Validation and Benchmarking:**  
-     Use cross-validation and hold-out test sets to ensure robustness of performance assessments.
-
-% ######################## Research Plan / Methodology ######################################
 \newpage
-To achieve the project goals and objectives, the research will follow a structured approach combining theoretical understanding with empirical evaluation:
-
-1. **Literature Review:**
-   - **Background Study:**  
-     Conduct a detailed review of seminal and recent works on RNNs, LSTMs, GRUs, and RNN encoder-decoder architectures in NLP. Key references include:
-     - Hochreiter \& Schmidhuber (1997) on LSTM.
-     - Cho et al. (2014) and Chung et al. (2014) on GRU.
-     - Recent surveys on the application of RNNs in NLP.
-   - **Problem Context:**  
-     Understand current best practices, challenges (e.g., vanishing/exploding gradients, sequence length handling), and evaluation standards in NLP.
-
-2. **Data Collection and Preprocessing:**
-   - **Dataset Selection:**  
-     Choose a benchmark NLP dataset appropriate to the target task (e.g., language modeling, text classification, or machine translation).
-   - **Preprocessing Steps:**  
-     Perform text normalization, tokenization, vocabulary building, and sequence padding. Optional steps may include initializing word embeddings (either pre-trained or learned from scratch).
-
-3. **Model Development and Experimental Implementation:**
-   - **Model Architecture Design:**  
-     Implement the different RNN-based models:
-     - **Vanilla RNN:**  
-       Define the recurrence as  
-       \[
-       h_t = \sigma(W_{hx} x_t + W_{hh} h_{t-1} + b_h), \quad y_t = f(W_{yh} h_t + b_y),
-       \]
-       where \(x_t\) is the input vector at time \(t\), \(h_t\) is the hidden state, and \(y_t\) is the output.
-     - **LSTM:**  
-       Set up the LSTM cell with input, forget, and output gates to control information flow.
-     - **GRU:**  
-       Implement the GRU cell, which uses update and reset gates as a simplified alternative.
-     - **RNN Encoder-Decoder:**  
-       Develop an encoder to summarize the input sequence and a decoder to generate the target sequence.
-   - **Framework and Tools:**  
-     Develop the models using PyTorch or TensorFlow. Leverage available libraries for reproducibility and efficiency.
-
-4. **Training, Hyperparameter Tuning, and Evaluation:**
-   - **Training Protocol:**  
-     Use similar training configurations (e.g., optimizer choice, learning rate schedule, batch size) for all models to ensure a fair comparison.
-   - **Hyperparameter Optimization:**  
-     Conduct grid search or Bayesian optimization for parameters such as hidden unit size, dropout rates, and learning rates.
-   - **Evaluation Metrics:**  
-     Define metrics based on the nature of the task (e.g., classification accuracy, F1 score, perplexity). Monitor convergence behavior and computational resource usage.
-   - **Cross-Validation and Benchmarking:**  
-     Use cross-validation and hold-out test sets to ensure robustness of performance assessments.
-
-5. **Result Analysis and Reporting:**
-   - **Comparative Analysis:**  
-     Analyze and interpret the performance differences among the RNN, LSTM, GRU, and RNN encoder-decoder models. Assess the trade-offs in terms of processing speed, accuracy, and stability.
-   - **Discussion of Findings:**  
-     Provide a discussion relative to theoretical expectations (e.g., how and why LSTMs may outperform vanilla RNNs in handling long-term dependencies).
-   - **Documentation:**  
-     Compile the findings into a detailed technical report, including visualizations of loss curves, performance metrics over training epochs, and error analysis.
-
-6. **Dissemination:**
-   - **Presentation and Peer Review:**  
-     Present the results at seminars and workshops to receive feedback.
-   - **Final Reporting:**  
-     Prepare a final research report and, if applicable, a manuscript for publication in a relevant conference or journal.
+\subsection{Training, Hyperparameter Tuning, and Evaluation}
+In my experiments, I implemented the training loop using PyTorch.
+The following details describe the key components of my training code, which are designed to ensure robust model optimization and fair evaluation:
+
+\subsubsection{Model Hyperparameters and Setup}
+The following hyperparameters were chosen based on preliminary experiments and standard practices for sentiment classification tasks:
+\begin{itemize}
+    \item \textbf{Architecture Hyperparameters:}
+    \begin{itemize}
+        \item Number of layers: \texttt{no\_layers = 1}
+        \item Vocabulary size: \(\mathtt{vocab\_size = len(vocab) + 1}\) (the extra one is for unknown word)
+        \item Embedding dimension: \texttt{embedding\_dim = 100}
+        \item Hidden dimension: \texttt{hidden\_dim = 50}
+        \item Output dimension: \texttt{output\_dim = 1}
+        \item Dropout probability: \texttt{drop\_prob = 0.5}
+        \item Batch size: \texttt{batch\_size = 50}
+    \end{itemize}
+    \item \textbf{Optimization Hyperparameters:}
+    \begin{itemize}
+        \item Learning rate: \texttt{lr = 0.001}
+        \item Loss Function: Binary Cross Entropy Loss
+        \item Optimizer: Adam with predefined learning rate.
+    \end{itemize}
+\end{itemize}
+In order to keep the fair experiment. I train 3 RNN model with the same hyperparameters
+
+\subsubsection{Training Loop Overview}
+\begin{itemize}
+    \item \textbf{Mini-Batch Training with Hidden State Initialization:}  
+    At the start of each epoch, the model is set to training mode using \texttt{model.train()}. For each mini-batch from the training data:
+    \begin{enumerate}
+        \item The model’s hidden state is initialized with zero matrix.
+        \item Before the forward pass, gradients are zeroed via \texttt{model.zero\_grad()} and \texttt{optimizer.zero\_grad()}.
+        \item The forward pass is executed, producing outputs and an updated hidden state.
+        \item The hidden state is detached from the computation graph using \texttt{h.detach()} to avoid backpropagating through all previous batches.
+        \item The loss is computed using binary cross entropy.
+        \item Backpropagation is performed via \texttt{loss.backward()}.
+        \item The gradients are clipped using \texttt{nn.utils.clip\_grad\_norm\_} with the specified clip value to mitigate exploding gradients.
+        \item Finally, the optimizer updates the model parameters using \texttt{optimizer.step()}.
+    \end{enumerate}
+
+    \item \textbf{Validation Stage:}  
+    After each epoch, the model is switched to evaluation mode using \texttt{model.eval()} and evaluated on the validation set.
+Similar to training, the hidden state is updated parameters and the model will use that parameters to make the prediction.
+And the accuracy is computed and recorded.
+
+    \item \textbf{Performance Monitoring and Checkpointing:}  
+    At the end of each epoch, the mean training and validation losses, as well as accuracy, are calculated.
+The validation loss is compared against the best loss seen so far.
+If the current validation loss is lower than the lowerest one, then the model's state dictionary is saved as the best model in .pth format.
+\end{itemize}
+
+\subsubsection{Pseudocode Description}
+The following Python pseudocode summarizes the training procedure implemented in my experiment:
+
+\begin{lstlisting}[language=Python, caption={Training Loop Pseudocode}]
+clip = 5
+epochs = 5 
+valid_loss_min = np.inf
+epoch_tr_loss, epoch_vl_loss = [], []
+epoch_tr_acc, epoch_vl_acc = [], []
+
+for epoch in range(epochs):
+    train_losses = []
+    train_acc = 0.0
+    model.train()
+    # Initialize the hidden state for the first mini-batch
+    h = model.init_hidden(batch_size)
+    for inputs, labels in train_loader:
+        inputs, labels = inputs.to(device), labels.to(device)
+        model.zero_grad()
+        optimizer.zero_grad()
+
+        output, h = model(inputs, h)
+        h = h.detach()
+
+        loss = criterion(output, labels.float())
+        loss.backward()
+        train_losses.append(loss.item())
+
+        accuracy = acc(output, labels)
+        train_acc += accuracy
+
+        nn.utils.clip_grad_norm_(model.parameters(), clip)
+        optimizer.step()
+
+    val_h = model.init_hidden(batch_size)
+    val_losses = []
+    val_acc = 0.0
+    model.eval()
+    for inputs, labels in valid_loader:
+        inputs, labels = inputs.to(device), labels.to(device)
+        output, val_h = model(inputs, val_h)
+        val_loss = criterion(output.squeeze(), labels.float())
+        val_losses.append(val_loss.item())
+
+        accuracy = acc(output, labels)
+        val_acc += accuracy
+
+    epoch_train_loss = np.mean(train_losses)
+    epoch_val_loss = np.mean(val_losses)
+    epoch_train_acc = train_acc / len(train_loader.dataset)
+    epoch_val_acc = val_acc / len(valid_loader.dataset)
+
+    epoch_tr_loss.append(epoch_train_loss)
+    epoch_vl_loss.append(epoch_val_loss)
+    epoch_tr_acc.append(epoch_train_acc)
+    epoch_vl_acc.append(epoch_val_acc)
+
+    print(f'Epoch {epoch+1}')
+    print(f'train_loss : {epoch_train_loss}  val_loss : {epoch_val_loss}')
+    print(f'train_accuracy : {epoch_train_acc*100}  val_accuracy : {epoch_val_acc*100}')
+
+    if epoch_val_loss <= valid_loss_min:
+        torch.save(model.state_dict(), './working/state_dict.pt')
+        print('Validation loss decreased ({:.6f} --> {:.6f}). Saving model...'
+              .format(valid_loss_min, epoch_val_loss))
+        valid_loss_min = epoch_val_loss
+    print(25 * '==')
+\end{lstlisting}
+\newpage
 

### 2025-03-07 paper2/main.tex
paper update

@@
 
+% ######################## Results and Analysis ######################################
+\section{Results and Analysis}
+\subsection{Results}
+In this section, I report the classification accuracy of three model on the testing set. And I will show the performance of RNN, GRU and LSTM on IMDB dataset respectively. 
+\\[1ex]
+For the Figure 2, RNN performance on IMDB dataset.
+Here is the deeply interpretation.
+For the left hand side is the accuracy.
+We can see that the training accuracy improves steadly over the 5 epochs and starting around 50\% and reaching roughly 70\%.
+And the validation accuracy also increase and follow the similar trend to the training accuracy.
+Starting at around 50\% and reaching around 60\% at the end of the validation.
+\\[1ex]
+For the training loss, it decreases steadily over 5 epochs, starting around 0.7 and dropping to approximately 0.56, the validation loss initially decreases but begins to raise after $4^{th}$ epoch.
+This suggests that the model's performance on unseen data is not improving as much after the $4^{th}$ epoch and may even start to degrade due to overfitting or some gradient problem.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/rnn_performance.png} 
+    \caption{RNN performance on IMDB dataset}
+\end{figure}
+\\[1ex]
+And next is the performance of GRU, we can see that the training accuracy improves significantly, starting around 60\% and reaching over 95\% by the end of training.
+This indicates that the GRU model is fitting the training data very well.
+On the contrast, the validation accuracy starts at around 80\% and improves slightly in the first 2 epochs but then remain on the same value around 85\% for the rest of the training.
+The widening gap between training and validation accuracy after the $2^{nd}$ epoch suggests overfitting.
+\\[1ex]
+Then for the training loss and the validation loss for the GRU model.
+The training loss decreases steadily and starting at over 0.6 and dropping to around 0.1 by the $5^{th}$ epoch.
+The gap between training and validation accuracy after the second epoch suggests overfitting.
+\\[1ex]
+For the loss, the training loss decreases steadily, starting at over 0.6 and dropping to around 0.1 by the epoch.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/gru_performance.png} 
+    \caption{GRU performance on IMDB dataset}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm_performance.png}
+    \caption{LSTM performance on IMDB dataset}
+\end{figure}
+
+\newpage
+% ######################## Results and Analysis ######################################
+
 % ######################## Reference ######################################

### 2025-03-08 paper2/main.tex
paper update

@@
 This unfolds the RNN to construct a traditional Feedforward Neural Network where we can apply backpropagation through time. Below is the conceptual idea of BPTT in RNN.
+\newpage
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
+    \includegraphics[width=0.8\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
     \caption{Unfolded RNN}
@@
 
+\newpage
 \textbf{Pseudocode of BPTT} \parencite{wikipedia2023bptt}
@@
 Below is the sigmoid function and its derivative.
-\begin{figure}[!htb]
+\newpage
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/sigmoid.png} % Replace "example-image" with your image file
-    % \caption{An example image.}
-    \label{fig:sigmoid}
+    \includegraphics[width=0.8\textwidth]{../Pic/sigmoid.png} % Replace "example-image" with your image file
 \end{figure}
-\newline
+
 $Domain(Sigmoid(x))=\mathbb{R},\hspace{1em} Codomain(Sigmoid(x))=(0,1)$\\
 $Domain(Sigmoid'(x))=\mathbb{R},\hspace{1em} Codomain(Sigmoid'(x))=[0,0.5]$
+\newline
 
@@
 \end{equation}
-Below is the Hyperbolic tangent activation function and its derivative.\\
+Below is the Hyperbolic tangent activation function and its derivative.
 \begin{figure}[!htb]
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/tanh.png} % Replace "example-image" with your image file
-    % \caption{An example image.}
-    \label{fig:tanh}
+    \includegraphics[width=0.8\textwidth]{../Pic/tanh.png} % Replace "example-image" with your image file
 \end{figure}
-\newline
+
 $Domain(tanh(x))=\mathbb{R},\hspace{1em} Codomain(tanh(x))=[-1,1]$\\
@@
 Then the result will decrease exponentially, then vanishing gradients occur.
-\begin{figure}[!htb]
+\begin{figure}[!h]
     \centering
@@
 In this section, I report the classification accuracy of three model on the testing set. And I will show the performance of RNN, GRU and LSTM on IMDB dataset respectively. 
-\\[1ex]
+\subsubsection{Vanilla RNN}
 For the Figure 2, RNN performance on IMDB dataset.
@@
 \end{figure}
-\\[1ex]
+\subsubsection{GRU}
 And next is the performance of GRU, we can see that the training accuracy improves significantly, starting around 60\% and reaching over 95\% by the end of training.
@@
 \\[1ex]
-For the loss, the training loss decreases steadily, starting at over 0.6 and dropping to around 0.1 by the epoch.
-
+For the loss, the training loss decreases steadily, starting at over 0.6 and dropping to around 0.1 by the $5^{th}$ epoch.
+For the validation loss decreases initially and reaching the lowest and then the loss raise over 0.5.
+Which indicate that overfitting.
 \begin{figure}[h!]
@@
 \end{figure}
+
+\subsubsection{LSTM}
+The third one is the performance of LSTM, the training accuracy improves steadily over the epochs and starting at around 55\% and reaching about 90\% at the end of the training.
+And for the validation accuracy starting at approximately 65\% and reaching about 82.5\%.
+And the loss of the training dataset decreases steadily from about 0.7 to around 0.2.
+And the validation loss show that the model did learn something from the data.
 \begin{figure}[h!]
@@
 \newpage
+\section{Analysis}
+\subsection{Learning Behavior}
+All models show the effective learning during the initial epochs, roughtly 55\%, as both training and validation accuracy improve, and loss decrease.
+But the common problem is overfitting occure in 3 models.
+The possible reasons is dataset.
+Since every sentence in dataset was truncated.
+I just use the first 100 token to training the model.
+So the existing tokens might not be fully represent the sentence.
+The second reason is the embedding layers, I used the built-in function in Pytorch to create the embedding layer by randomise the matrix.
+\\[1ex]
+Among the 3 models, LSTM and GRU have the better performance compare with vanilla RNN.
+The reason of GRU and LSTM perform better than RNN is the vanilla RNN struggles with long term dependencies because of the vanishing gradient problem when processing long sequences.
+In this experience is 100 time step.
+And this limits its ability to learn the meaningful patterns from data.
+On the contrast, GRU and LSTM use gate mechanisms to retain the relevant information and discard irrelevant information, allowing them to capture the long term dependencies more effectively.
+\\[1ex]
+And for the overfitting problem, we can see that overfitting problem occur in three models.
+
+\newpage
 % ######################## Results and Analysis ######################################

### 2025-03-08 paper2/main.tex
paper update

@@
 
+\subsubsection{Model 4 (vanilla RNN with 2 layer)}
+Same setting with model 1, but I stack 2 RNN block. In the other word, this is deep recurrent neural networks. 
+\subsubsection{Model 5 (GRU with 2 layer)}
+Same setting with model 1, but I stack 2 GRU block. In the other word, this is deep recurrent neural networks. 
+\subsubsection{Model 6 (LSTM RNN with 2 layer)}
+Same setting with model 1, but I stack 2 LSTM block. In the other word, this is deep recurrent neural networks.
+
 \newpage
@@
 And for the overfitting problem, we can see that overfitting problem occur in three models.
-
+For the vanilla RNN, since this is the basic and simple architecture to handle sequential data, it lacks mechanisms to retain or forget information selectively.
+As a result, it relying on the memorizing the training data rather than learning the meaningful patterns.
+This causes it to overfit.
+And for the GRU and LSTM, although gating mechanisms provide a more sophisticated architecture to learning from data, they aggressively optimizes the training loss, which can result in overfitting when the model continues to learn unnecessary information from the dataset.
+\\[1ex] 
+The above result and analyse shown that GRU and LSTM perform better than vanilla RNN, which show that the graident vanishing and graident exploding problem were improved by the gating mechanism.
+And there is so many way will affect the experiment result for example, the embedding layer, tokenization.
 \newpage
+
 % ######################## Results and Analysis ######################################

### 2025-03-08 decode_text.py

@@ -1,48 +1,0 @@
-from transformers import AutoTokenizer
-
-class TextDecoder:
-    def __init__(self, tokenizer_name="bert-base-uncased"):
-        """
-        Initialize the TextDecoder with a tokenizer.
-        
-        Args:
-            tokenizer_name (str): Name of the tokenizer to use (default: 'bert-base-uncased').
-        """
-        self.tokenizer = AutoTokenizer.from_pretrained(tokenizer_name)
-    
-    def decode_text(self, token_ids, skip_special_tokens=True):
-        """
-        Decode a list of token IDs into human-readable text.
-        
-        Args:
-            token_ids (list[int]): List of token IDs to decode.
-            skip_special_tokens (bool): Whether to skip special tokens like [CLS], [SEP], etc.
-        
-        Returns:
-            str: The decoded human-readable text.
-        """
-        return self.tokenizer.decode(token_ids, skip_special_tokens=skip_special_tokens)
-    
-    def decode_answer(self, context_ids, start_pos, end_pos, skip_special_tokens=True):
-        """
-        Decode the answer from a context given the start and end positions.
-        
-        Args:
-            context_ids (list[int]): List of token IDs for the context.
-            start_pos (int): Start position of the answer in the context.
-            end_pos (int): End position of the answer in the context.
-            skip_special_tokens (bool): Whether to skip special tokens like [CLS], [SEP], etc.
-        
-        Returns:
-            str: The decoded human-readable answer.
-        """
-        answer_ids = context_ids[start_pos:end_pos + 1]
-        return self.tokenizer.decode(answer_ids, skip_special_tokens=skip_special_tokens)
-    
-if __name__ == '__main__':
-    # Initialize the decoder
-    decoder = TextDecoder(tokenizer_name="bert-base-uncased")
-    # Example: Decoding a single tokenized text
-    token_ids = [101, 7592, 2003, 1037, 6584, 102, 0]  # Example token IDs
-    decoded_text = decoder.decode_text(token_ids)
-    print("Decoded Text:", decoded_text)

### 2025-03-08 test.py

@@ -1,163 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.optim as optim
-from torch.utils.data import DataLoader, Dataset
-from transformers import AutoTokenizer
-from datasets import load_dataset
-import numpy as np
-
-# Step 1: Load the SQuAD dataset
-squad = load_dataset("squad")
-
-# Extract data
-train_data = squad['train']
-contexts = [item['context'] for item in train_data]
-questions = [item['question'] for item in train_data]
-answers = [item['answers']['text'][0] for item in train_data]
-answer_starts = [item['answers']['answer_start'][0] for item in train_data]
-
-# Step 2: Preprocess the data
-tokenizer = AutoTokenizer.from_pretrained("bert-base-uncased")
-
-# Tokenize context and question
-def preprocess_data(context, question, answer, answer_start):
-    tokenized_context = tokenizer(context, truncation=True, padding="max_length", max_length=300, return_tensors="pt")
-    tokenized_question = tokenizer(question, truncation=True, padding="max_length", max_length=50, return_tensors="pt")
-    
-    # Compute answer tokens
-    answer_end = answer_start + len(answer)
-    token_start = tokenized_context.char_to_token(answer_start)
-    token_end = tokenized_context.char_to_token(answer_end - 1)
-    
-    if token_start is None or token_end is None:
-        # If the answer cannot be tokenized, return None (to filter out later)
-        return None
-    
-    return {
-        "context": tokenized_context["input_ids"].squeeze(0),
-        "question": tokenized_question["input_ids"].squeeze(0),
-        "start_pos": token_start,
-        "end_pos": token_end
-    }
-
-# Preprocess the entire dataset
-processed_data = []
-for c, q, a, s in zip(contexts, questions, answers, answer_starts):
-    item = preprocess_data(c, q, a, s)
-    if item:
-        processed_data.append(item)
-
-# Convert to PyTorch Dataset
-class SquadDataset(Dataset):
-    def __init__(self, data):
-        self.data = data
-
-    def __len__(self):
-        return len(self.data)
-
-    def __getitem__(self, idx):
-        item = self.data[idx]
-        return item["context"], item["question"], item["start_pos"], item["end_pos"]
-
-dataset = SquadDataset(processed_data)
-dataloader = DataLoader(dataset, batch_size=16, shuffle=True)
-
-# Step 3: Define the RNN-based model
-class RNNQuestionAnsweringModel(nn.Module):
[…]

### 2025-03-09 paper2/main.tex
paper update

@@
 In this section, I report the classification accuracy of three model on the testing set. And I will show the performance of RNN, GRU and LSTM on IMDB dataset respectively. 
-\subsubsection{Vanilla RNN}
-For the Figure 2, RNN performance on IMDB dataset.
+\subsubsection{Vanilla RNN (1 layer)}
+For the figure 2, RNN with 1 layer performance on IMDB dataset.
 Here is the deeply interpretation.
 For the left hand side is the accuracy.
-We can see that the training accuracy improves steadly over the 5 epochs and starting around 50\% and reaching roughly 70\%.
-And the validation accuracy also increase and follow the similar trend to the training accuracy.
-Starting at around 50\% and reaching around 60\% at the end of the validation.
-\\[1ex]
-For the training loss, it decreases steadily over 5 epochs, starting around 0.7 and dropping to approximately 0.56, the validation loss initially decreases but begins to raise after $4^{th}$ epoch.
-This suggests that the model's performance on unseen data is not improving as much after the $4^{th}$ epoch and may even start to degrade due to overfitting or some gradient problem.
+We can see that the training accuracy increasing steadily from 50\% to roughly 65\%.
+And the validation accuracy shows a more erratic behavior.
+It starts from 50\% and increases to a peak around epoch 3, drops slightly at epoch 4 and the rises to 65\% by the end of the training.
+\\[1ex]
+For the training loss, it decreases steadily over 5 epochs from 0.7 to 0.6, showing that the model is learning.
+On the contrary, the validation loss decreases initially and stabilizes around epoch 4 and then rises sharply in later epochs.
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/rnn_performance.png} 
-    \caption{RNN performance on IMDB dataset}
+    \includegraphics[width=1\textwidth]{../Pic/rnn1.png} 
+    \caption{RNN with 1 layer performance on IMDB dataset}
 \end{figure}
-\subsubsection{GRU}
-And next is the performance of GRU, we can see that the training accuracy improves significantly, starting around 60\% and reaching over 95\% by the end of training.
+\subsubsection{Vanilla RNN (2 layer)}
+The training accuracy improves over the eopchs, increasing steadily from 50\% to 75\%.
+For the validation accuracy also improves, although at a slower pace than training accuracy.
+Starts at around 50\% and reaching approximately 67\% by epoch 5.
+\\[1ex]
+Then for the tarining loss, it decreases steadily from 0.7 to 0.54, showing that the model is learning effectively and optimizing the training objective.
+Next for the validation loss, decreases initially, but plateaus and slightly increases after epoch 4.
+\newpage
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/rnn2.png} 
+    \caption{RNN with 2 layer performance on IMDB dataset}
+\end{figure}
+\subsubsection{GRU (1 layer)}
+For the figure 4, GRU with 1 layer, we can see that the training accuracy improves steadily and reaching around 95\%, but the validation accuracy remain the same value after epoch 2.
+Which mean that the model did not learning any new information from the training dataset.
+\\[1ex]
+Then for the training loss, the performance is relatively good and drop from 0.7 to 0.15. But for the validation loss, it drop from 0.55 to 0.4 and then rises to 0.45. 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/gru1.png} 
+    \caption{GRU with 1 layer performance on IMDB dataset}
+\end{figure}
+\subsubsection{GRU (2 layer)}
+And next is the performance of GRU with 1 layer, we can see that the training accuracy improves significantly, starting around 60\% and reaching over 95\% by the end of training.
 This indicates that the GRU model is fitting the training data very well.
@@
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/gru_performance.png} 
-    \caption{GRU performance on IMDB dataset}
+    \includegraphics[width=1\textwidth]{../Pic/gru2.png} 
+    \caption{GRU with 2 layer performance on IMDB dataset}
 \end{figure}
 
-\subsubsection{LSTM}
-The third one is the performance of LSTM, the training accuracy improves steadily over the epochs and starting at around 55\% and reaching about 90\% at the end of the training.
+\subsubsection{LSTM (1 layer)}
+And then for the training accuracy of LSTM with 1 layer, the train and validation accuracy increase steadily, for the training accuracy increase from 55\% to 90\%, for the validation loss is increase from 63\% to 83\%.
+For the training loss also decrease steadily, from 0.7 to 0.3.
+And the validation loss decrease from 0.7 to 0.43.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/lstm1performance.png} 
+    \caption{LSTM with 1 layer performance on IMDB dataset}
+\end{figure}
+\subsubsection{LSTM (2 layer)}
+Next is the performance of LSTM with 1 layer, the training accuracy improves steadily over the epochs and starting at around 55\% and reaching about 90\% at the end of the training.
 And for the validation accuracy starting at approximately 65\% and reaching about 82.5\%.
@@
     \centering
-    \includegraphics[width=1\textwidth]{../Pic/lstm_performance.png}
-    \caption{LSTM performance on IMDB dataset}
+    \includegraphics[width=1\textwidth]{../Pic/lstm2performance.png}
+    \caption{LSTM with 2 layer performance on IMDB dataset}
 \end{figure}

### 2025-03-26 paper2/main.tex
paper update

@@
 
-\subsubsection{Model 4 (vanilla RNN with 2 layer)}
+\subsubsection{Model 4 (Vanilla RNN with 2 layer)}
 Same setting with model 1, but I stack 2 RNN block. In the other word, this is deep recurrent neural networks. 
@@
 
+% ######################## Discussion and Conclusion ######################################
+\section{Discussion and Conclusion}
+
+\subsection{Discussion}
+Throughout the project, several unexpected results and limitations were encountered that shed light on both the robustness and challenges of using RNN-based models for sentiment classification.
+
+\begin{itemize}
+    \item \textbf{Unexpected Results:}  
+    \begin{itemize}
+        \item The vanilla RNN model showed signs of overfitting relatively quickly in comparison to the LSTM and GRU models.
+This behavior was attributed to its vulnerability to vanishing and exploding gradient issues, despite the application of gradient clipping.
+        \item Although gradient clipping (with a threshold value of 5) generally stabilized training, occasional oscillations in the validation loss were observed.
+Such fluctuations suggest that even with robust gradient management, the training dynamics of sequential models can be highly sensitive to hyperparameter settings.
+    \end{itemize}
+
+    \item \textbf{Limitations:}  
+    \begin{itemize}
+        \item The pre-processing pipeline, while extensive, was sometimes insufficient to handle complex or ambiguous tokens (e.g., domain-specific abbreviations).
+This highlights the need for more advanced pre-processing—such as incorporating contextualized word embeddings—to handle semantic variations effectively.
+        \item The scope of hyperparameter tuning was limited to grid or random search methods.
+A more comprehensive hyperparameter optimization (e.g., Bayesian optimization) might yield further improvements.
+        \item Resource constraints (such as batch size and training epochs) influenced the convergence behavior and model performance.
+Future experiments could benefit from increased computational resources to explore deeper architectures or longer sequences.
+    \end{itemize}
+\end{itemize}
+
+\subsection{Conclusion}
+In summary, this project conducted a detailed comparative analysis of three RNN-based architectures (vanilla RNN, LSTM, and GRU) for binary sentiment classification on the ACL-IMDB dataset.
+The main findings, contributions, and implications of the project are as follows:
+
+\begin{itemize}
+    \item \textbf{Main Findings:}
+    \begin{itemize}
+        \item Gated architectures (LSTM and GRU) outperformed the vanilla RNN in managing long-term dependencies and reducing vanishing/exploding gradient issues.
+        \item The structured training approach—incorporating hidden state management, gradient clipping, and periodic validation—proved critical to achieving stable convergence.
+    \end{itemize}
+
+    \item \textbf{Contributions:}
+    \begin{itemize}
+        \item The project establishes a robust training framework and experimental methodology, which includes comprehensive data pre-processing, model initialization, and performance evaluation.
+        \item It provides a systematic comparison of RNN variants, offering insights that can guide future research on neural network architectures for natural language processing.
+    \end{itemize}
+
+    \item \textbf{Implications:}
+    \begin{itemize}
+        \item The results suggest that practitioners working on sentiment analysis or related NLP tasks should consider employing gated architectures like LSTM or GRU to achieve better performance.
+        \item Future work may build on these findings by exploring more advanced architectures (e.g., Transformer models or hybrid models incorporating convolutional layers) as well as more sophisticated hyperparameter optimization techniques.
+    \end{itemize}
+\end{itemize}
+
+In conclusion, while the project revealed certain limitations, it nevertheless contributes valuable insights into the training and evaluation of RNN-based models for sentiment classification, thereby setting the stage for further research in the domain of efficient and scalable NLP methodologies.
+% ######################## Discussion and Conclusion ######################################
+
 % ######################## Reference ######################################

### 2025-03-28 paper2/main.tex
paper update

--- 2025-03-09 paper2/main.tex
@@
 \tableofcontents % Generate a table of contents       
+\newpage                 
 % ######################## table of content ###############################
@@
 This unfolds the RNN to construct a traditional Feedforward Neural Network where we can apply backpropagation through time. Below is the conceptual idea of BPTT in RNN.
-\newpage
 \begin{figure}[h!]
@@
 
+\subsubsection{RNN encoder decoder model}
+RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering.
+The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data.
+\\[1ex]
+In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector.
+And the decoder is to decode the vector into the target sequence.
+\\[1ex]
+\textbf{Encoder}
+\\[1ex]
+The encoder is a RNN based block, which processes the input sequence step by step and calculate the hidden state representation for each input token.
+Let the input sequence be $X = [x_1, x_2, \cdots, x_T]$ where $T$ is the length of the input sequence, and $x_t$ is the input at time step $t$.
+We use the below formula to computes the hidden states $h_t$.
+\\[1ex]
+For the Vanilla RNN encoder
+\begin{equation}
+    h_t = f_{enc}(x_t, h_{t-1})
+\end{equation}
+
+where the $f_{enc}$ is the recurrent transition function, (e.g.
+LSTM, GRU or Vanilla RNN cell), and $h_t$ is the hidden state at time step $t$.
+And then the final hidden state of the encoder, $h_T$ will act as the context vector.
+\begin{equation}
+    c = h_T
+\end{equation}
+\textbf{Decoder}
+\\[1ex]
+The decoder is the second component of RNN encoder decoder, the usage of the decoder is to generates the output sequence step by step, conditioned on the context vector $c$ and its previous hidden state.
+\begin{equation}
+    s_t = f_{dec}(y_{t-1}, s_{t-1}, c)
+\end{equation}
+where $f_{dec}$ is the recurrent function of the decoder, and it is just like the encoder part, $y_{t-1}$ is the embedding of the previous output token, $s_{t-1}$ is the decoder's hidden state.
+And $c$ is the context vector from the encdoer, which remains constant throughout decoding.
+\\[1ex]
+After the model computed the $s_t$, the model will compute the output probability distribution.
+\begin{equation}
+    P(y_t|y_1, c) = softmax(W_os_t+b_o)
+\end{equation}
+And the predicted token $y_t = arg maxP(y_t)$, the process will repeat until the decoder predicts the end of sequence token $(<EOS>)$.
+
 \subsubsection{Hidden Markov Model}
 Before reviewing Hidden Markov Model (HMM), it is essential to understand what Markov models is, or Markov chains.
-Markov chains are fundamental models in probability theory and statistic, and it is a stochastic process describing a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
+Markov chains are fundamental models in probability theory and statistic, and it is a stochastic process, which describe a sequence of possible events in which the probability of each event depends only on the state attained in the previous event.
 This property known as the Markov property \parencite{roberts2004general, rabiner1986anintroduction}.
@@
 \subsubsection{Data Collection}
-In this report, ACL-IMDB \parencite{maas-EtAl:2011:ACL-HLT2011} was experiment.
-This is a dataset for binary sentiment classification containing substantially more data than previous benchmark datasets.
-This dataset provide a set of 25,000 highly polar movie reviews for training, and 25,000 for testing.
+In this report, ACL-IMDB \parencite{maas-EtAl:2011:ACL-HLT2011} was use and for this dataset is a binary sentiment classification containing the substantially more data than the last version benchmark datasets.
+This dataset provide a set of 25,000 highly polar movie reviews for training, and 25,000 for validation.
 I preporcess the dataset by text cleaning and tokenization.
@@
 And there is so many way will affect the experiment result for example, the embedding layer, tokenization.
+
 \newpage
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/gradient.png}
+\end{figure}
+\subsubsection{Vanilla RNN (Red line)}
+The gradients in the Vanilla RNN is unstable, as shown by the oscillations and spikes in the early timesteps.
+Over time, the gradients decay rapidly and go to 0, indicating that the presence of the vanishing gradient problem.
+This problem occurs because the Vanilla RNNs do not have the gating mechanisms in which this architecture is not perform well in long time dependencies.
+In other words, as the sequence length increases, repeated matrix multiplications lead to the gradients to shrink exponentially, making it difficult for the model to update the weighting matrix effectively during the BPTT.
+\\[1ex]
+\subsubsection{LSTM (Blue line)}
+The gradients of the LSTM are much more stable compared to the simple vanilla RNN.
+The gradient magnitudes remain consistent across timesteps and do not decay to zero.
+This stability is achieved through the use of gating machanisms that regulate the flow of information.
+These gates allow the LSTM to keep the relevant information form the input data while remove some irrelevant information from the data.
+As a result, LSTM is highly effective at learning long term dependencies in sequences.
+\\[1ex]
+\subsubsection{GRU (Green line)}
+The gradients in the GRU are also stable, similar to the LSTM, but the performance is more smoother.
+GRU achieves this stability using its gate mechanism, which control the degree of information through the network.
+Compared to the LSTM, the GRU has a simpler architecture, which makes it more computationally efficient while still addressing the vanishing gradient problem.
+
+\newpage
+
 % ######################## Results and Analysis ######################################
 
+% ######################## Discussion and Conclusion ######################################
+\section{Discussion and Conclusion}
+
+\subsection{Discussion}
+During the experimental phase, several insights and challenges emerged, which are discussed below:
+
+\begin{itemize}
+    \item \textbf{Unexpected Results:}  
+    It was observed that the loss of a two-layer LSTM model was higher compared to that of a one-layer LSTM.
+This was unexpected because additional layers are generally presumed to capture more complex patterns and improve performance.
+One possible explanation for this anomaly is the inherent randomness during model initialization and training.
+Since the performance difference can vary from run to run, it is possible that in some exceptions the two-layer model does not outperform the one-layer configuration.
+
+    \item \textbf{Limitations:}  
+    Due to limited computing resources, the experiments were constrained to training the models from scratch with relatively simpler architectures.
+This restricted the exploration of more complex or state-of-the-art models, such as Transformer-based architectures, which might offer further performance improvements but require substantially more computational power.
+\end{itemize}
+
+\subsection{Conclusion}
+The primary objective of this report was to employ various RNN-based models and compare their performance on a sentiment classification task. The main findings of the project are:
+
+\begin{itemize}
+    \item \textbf{Comparative Analysis:}  
+    Different RNN-based models, including vanilla RNN, LSTM, and GRU, were implemented and evaluated.
+The experiments show that LSTM and GRU models effectively address the long-term dependency problem inherent in sequential data.
+Their internal gating mechanisms allow them to control and preserve information over longer sequences.
+
+    \item \textbf{Efficacy of Gated Architectures:}  
+    The results indicate that the gated architectures (LSTM and GRU) significantly outperform the vanilla RNN in terms of handling vanishing and exploding gradient issues, thereby offering more stable and reliable performance.
+\end{itemize}
+
+In conclusion, while some unexpected results were observed—such as the higher loss in the two-layer LSTM model—the overall findings strongly advocate for the use of gated RNN architectures in tasks requiring modeling of long-term dependencies.
+Despite limitations due to computational resources, this report provides a valuable comparative study and lays the groundwork for future research on more complex models like Transformers in natural language processing tasks.
+
+In conclusion, while the project revealed certain limitations, it nevertheless contributes valuable insights into the training and evaluation of RNN-based models for sentiment classification, thereby setting the stage for further research in the domain of efficient and scalable NLP methodologies.
+% ######################## Discussion and Conclusion ######################################
+
 % ######################## Reference ######################################
+\newpage
 \printbibliography

### 2025-03-29 paper2/main.tex
fix bug

@@
 $g(\cdot)$ and $f(\cdot)$ are an activation function.
+\newpage
 
 \subsubsection{RNN encoder decoder model}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder.jpg} % Replace "example-image" with your image file
+\end{figure}
 RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering.
 The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data.
-\\[1ex]
 In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector.
@@
 We use the below formula to computes the hidden states $h_t$.
-\\[1ex]
-For the Vanilla RNN encoder
 \begin{equation}
@@
 And the predicted token $y_t = arg maxP(y_t)$, the process will repeat until the decoder predicts the end of sequence token $(<EOS>)$.
+
+\subsubsection{Attention Mechanism}
+Vaswani et al.
+introduces the Transformer architecture, which replaces RNN model with self attention.
+In this section, I want to review the attention machanism.
+\parencite{vaswani2023attentionneed}.
+Attention machanism is good for handle the sequence-to-sequence tasks such as machine translation, text summarization and image captioning.
+Bahdanau et al.
+\parencite{bahdanau2016neuralmachinetranslationjointly} introduces the attention mechanism in the machine translation task.
+At each decoding step $t'$, the model model will compute the attention scores $\alpha_{t',t}$ that represent how important of the current hidden state $h_t$ to the current decoder state $s_{t'}$.
+We have different attention formula.
+\\[1ex]
+\textbf{Bahdanau (Additive) Attention}
+\\[1ex]
+This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
+\\[1ex]
+\begin{equation}
+    e_{t',t} = v_a^Ttanh(W_as_{t'-1}+U_ah_t)
+\end{equation}                    
+\textbf{Luong (Multiplicative) Attention}
+\\[1ex]
+This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
+\begin{equation}
+    e_{t',t} = s_{t'-1}^Th_t
+\end{equation}
+\textbf{Scaled Dot-Product Attention}
+\\[1ex]
 

### 2025-03-29 paper2/main.tex
fixed bug

--- 2025-03-28 paper2/main.tex
@@
 
+\newpage
+
 \subsubsection{RNN encoder decoder model}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder.jpg} % Replace "example-image" with your image file
+\end{figure}
 RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering.
@@
 And the predicted token $y_t = arg maxP(y_t)$, the process will repeat until the decoder predicts the end of sequence token $(<EOS>)$.
+
+\subsubsection{Attention Mechanism}
+% Vaswani et al.
+introduces the Transformer architecture, which replaces RNN model with self attention.
+In this section, I want to review the attention machanism.
+\parencite{vaswani2023attentionneed}.
+Attention machanism is good for handle the sequence-to-sequence tasks such as machine translation, text summarization and image captioning.
+Bahdanau et al.
+\parencite{bahdanau2016neuralmachinetranslationjointly} introduces the attention mechanism in the machine translation task.
+At each decoding step $t'$, the model model will compute the attention scores $\alpha_{t',t}$ that represent how important of the current hidden state $h_t$ to the current decoder state $s_{t'}$.
+We have different attention formula.
+
+\textbf{Bahdanau (Additive) Attention}
+\\[1ex]
+% This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
+\\[1ex]
+\begin{equation}
+    e_{t',t} = v_a^Ttanh(W_as_{t'-1}+U_ah_t)
+\end{equation}                    
+\textbf{Luong (Multiplicative) Attention}
+\\[1ex]
+% This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
+\begin{equation}
+    e_{t',t} = s_{t'-1}^Th_t
+\end{equation}
+\textbf{Scaled Dot-Product Attention}
+\\[1ex]
 

### 2025-03-29 paper2/main.tex
paper update

@@
 This sequential processing produces outputs and stores the internal states of the network in any period \parencite{werbos1990bptt}.
-
+\\[1ex]
 This unfolds the RNN to construct a traditional Feedforward Neural Network where we can apply backpropagation through time. Below is the conceptual idea of BPTT in RNN.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
+    \includegraphics[width=0.65\textwidth]{../Pic/pic1.png} % Replace "example-image" with your image file
     \caption{Unfolded RNN}
 \end{figure}
+
+% Forward pass
+\noindent \textbf{Forward Pass}
+\\[1ex]
+During the forward pass, the RNN processes the input sequence sequentially, computing hidden states and output at each timestep:
+\begin{equation}
+    h_t = f(U^Tx_{t}+W^Th_{t-1}+b_h)
+\end{equation}
+\begin{equation}
+    \hat{y}_t = f(V^Th_t+b_y)
+\end{equation}
+\newline  % Computing the loss function
+\noindent \textbf{Computing the loss function}
+\\[1ex]
+Assuming the loss is computed only at the final timestep t:
+\begin{equation}
+    \mathcal{L}_t = L(y_t, \hat{y}_t)
+\end{equation}
+In order to do backpropagation through time to tune the parameters in RNN, we need to calculate the partial derivative of loss function $\mathcal{L}$ with respect to the differently parameters.\\
+\newline  % Backward pass using the chain rule
+\noindent \textbf{Backward pass using the chain rule}
+\\[1ex]
+Using the chain rule for computing the gradient.\\
+Partial derivative of loss function $\mathcal{L}$ with respect to $W$ (hidden to hidden state) at time 2. 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_2}{\partial W} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial W}
+\end{equation}
+By mathematic induction 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_t}{\partial W} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial W})
+\end{equation}
+Where 
+\begin{equation}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
+\end{equation}
+
+Partial derivative of loss function $\mathcal{L}$ with respect to $U$ (input to hidden state) at time 2.\\
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_2}{\partial U} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial U}
+\end{equation}
+By mathematic induction 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_t}{\partial U} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial U})
+\end{equation}
+Where 
+\begin{equation}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
+\end{equation}
+
+Partial derivative of loss function $\mathcal{L}$ with respect to $V$ (hidden to output state) at time 2.\\
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_2}{\partial V} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial V}
+\end{equation}
+By mathematic induction 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_t}{\partial V} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial V})
+\end{equation}
+Where 
+\begin{equation}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
+\end{equation}
+
+Partial derivative of loss function $\mathcal{L}$ with respect to $b_h$ (bias term in hidden state) at time 2.\\
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_2}{\partial b_b} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_h}
+\end{equation}
+By mathematic induction 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_t}{\partial b_h} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_h})
+\end{equation}
+Where 
+\begin{equation}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
+\end{equation}
+
+Partial derivative of loss function $\mathcal{L}$ with respect to $b_y$ (bias term in output state) at time 2.
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_2}{\partial b_y} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_y}
+\end{equation}
+By mathematic induction 
+\begin{equation}
+    \dfrac{\partial\mathcal{L}_t}{\partial b_y} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_y})
+\end{equation}
+Where 
+\begin{equation}
+    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
+\end{equation}
+
+\textbf{parameters updates}
+\begin{equation}
+    W \leftarrow W - \alpha\dfrac{\partial\mathcal{L}}{\partial W}
+\end{equation}
+\begin{equation}
+    U \leftarrow U - \alpha\dfrac{\partial\mathcal{L}}{\partial U}
+\end{equation}
+\begin{equation}
+    V \leftarrow V - \alpha\dfrac{\partial\mathcal{L}}{\partial V}
+\end{equation}
+\begin{equation}
+    b_h \leftarrow b_h - \alpha\dfrac{\partial\mathcal{L}}{\partial b_h}
+\end{equation}
+\begin{equation}
+    b_y \leftarrow b_y - \alpha\dfrac{\partial\mathcal{L}}{\partial b_y}
+\end{equation}
 
@@
     \hline
-    $U$          & Weight matrix for input to hidden state       & $input\ size\times hidden\ unites$\\
-    $W$          & Weight matrix for hidden to hidden state      & $hidden\ units\times hidden\ unites$\\
+    $U$          & Weight matrix for input to hidden state       & $input\ size\times hidden\ units$\\
+    $W$          & Weight matrix for hidden to hidden state      & $hidden\ units\times hidden\ units$\\
     $V$          & Weight matrix for hidden state to output state& $hidden\ units\times number\ of\ class$\\
@@
 \end{longtable}
-% Forward pass
-\noindent \textbf{Forward Pass}
-\\[1ex]
-During the forward pass, the RNN processes the input sequence sequentially, computing hidden states and output at each timestep:
-\begin{equation}
-    h_t = f(U^Tx_{t}+W^Th_{t-1}+b_h)
-\end{equation}
-\begin{equation}
-    \hat{y}_t = f(V^Th_t+b_y)
-\end{equation}
-\newline  % Computing the loss function
-\noindent \textbf{Computing the loss function}
-\\[1ex]
-Assuming the loss is computed only at the final timestep t:
-\begin{equation}
-    \mathcal{L}_t = L(y_t, \hat{y}_t)
-\end{equation}
-In order to do backpropagation through time to tune the parameters in RNN, we need to calculate the partial derivative of loss function $\mathcal{L}$ with respect to the differently parameters.\\
-\newline  % Backward pass using the chain rule
-\noindent \textbf{Backward pass using the chain rule}
-\\[1ex]
-Using the chain rule for computing the gradient.\\
-Partial derivative of loss function $\mathcal{L}$ with respect to $W$ (hidden to hidden state) at time 2. 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial W} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial W}
-\end{equation}
-By mathematic induction 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_t}{\partial W} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial W})
-\end{equation}
-Where 
-\begin{equation}
-    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
-\end{equation}
-
-Partial derivative of loss function $\mathcal{L}$ with respect to $U$ (input to hidden state) at time 2.\\
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial U} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial U}
-\end{equation}
-By mathematic induction 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_t}{\partial U} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial U})
-\end{equation}
-Where 
-\begin{equation}
-    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
-\end{equation}
-
-Partial derivative of loss function $\mathcal{L}$ with respect to $V$ (hidden to output state) at time 2.\\
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial V} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial V}
-\end{equation}
-By mathematic induction 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_t}{\partial V} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial V})
-\end{equation}
-Where 
-\begin{equation}
-    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
-\end{equation}
-
-Partial derivative of loss function $\mathcal{L}$ with respect to $b_h$ (bias term in hidden state) at time 2.\\
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial b_b} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_h}
-\end{equation}
-By mathematic induction 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_t}{\partial b_h} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_h})
-\end{equation}
-Where 
-\begin{equation}
-    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
-\end{equation}
-
-Partial derivative of loss function $\mathcal{L}$ with respect to $b_y$ (bias term in output state) at time 2.
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_2}{\partial b_y} = \dfrac{\partial\mathcal{L}_2}{\partial y_2} \cdot \dfrac{\partial y_2}{\partial h_2} \cdot \dfrac{\partial h_2}{\partial h_1} \cdot \dfrac{\partial h_1}{\partial b_y}
-\end{equation}
-By mathematic induction 
-\begin{equation}
-    \dfrac{\partial\mathcal{L}_t}{\partial b_y} = \dfrac{\partial\mathcal{L}_t}{\partial y_t} \cdot \dfrac{\partial y_t}{\partial h_t} \cdot (\sum_{i=1}^{t} \dfrac{\partial h_t}{\partial h_i} \cdot \dfrac{\partial h_i}{\partial b_y})
-\end{equation}
-Where 
-\begin{equation}
-    \dfrac{\partial h_t}{\partial h_i} = \prod_{j=i+1}^{t} \dfrac{\partial h_j}{\partial h_{j-1}}
-\end{equation}
-
-\textbf{parameters updates}
-\begin{equation}
-    W \leftarrow W - \alpha\dfrac{\partial\mathcal{L}}{\partial W}
-\end{equation}
-\begin{equation}
-    U \leftarrow U - \alpha\dfrac{\partial\mathcal{L}}{\partial U}
-\end{equation}
-\begin{equation}
-    V \leftarrow V - \alpha\dfrac{\partial\mathcal{L}}{\partial V}
-\end{equation}
-\begin{equation}
-    b_h \leftarrow b_h - \alpha\dfrac{\partial\mathcal{L}}{\partial b_h}
-\end{equation}
-\begin{equation}
-    b_y \leftarrow b_y - \alpha\dfrac{\partial\mathcal{L}}{\partial b_y}
-\end{equation}
 
@@
 \begin{equation}
-    h_t = f_{enc}(x_t, h_{t-1})
-\end{equation}
-
-where the $f_{enc}$ is the recurrent transition function, (e.g.
-LSTM, GRU or Vanilla RNN cell), and $h_t$ is the hidden state at time step $t$.
-And then the final hidden state of the encoder, $h_T$ will act as the context vector.
-\begin{equation}
-    c = h_T
+    h^{enc}_t = tanh(W^{enc}_{ih}x_t + W^{enc}_{hh}h_{t-1}+b^{enc}_h)
+\end{equation}
+
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $W^{enc}_{ih}$          & Weight matrix for input to hidden state in encoder      & $hidden\ units\times input\ size$\\
+    $W^{enc}_{hh}$          & Weight matrix for hidden to hidden state in encoder     & $hidden\ units\times hidden\ units$\\
+    $b^{enc}_{h}$           & Bias vector for the hidden state in encoder             & $hidden\ units\times hidden\ units$\\
+    $h^{enc}_{t}$           & Hidden state at timestep t in encoder                   & $hidden\ units\times hidden\ units$\\
+    \hline
+    \caption{Encoder}
+\end{longtable}
+After the model processed the entire input sequence from $t=1$ to $t=T$, the last hidden state $h^{enc}_T$ treat as the context vector $c$. 
+\begin{equation}
+    c = h^{enc}_T
 \end{equation}
@@
 The decoder is the second component of RNN encoder decoder, the usage of the decoder is to generates the output sequence step by step, conditioned on the context vector $c$ and its previous hidden state.
-\begin{equation}
-    s_t = f_{dec}(y_{t-1}, s_{t-1}, c)
-\end{equation}
-where $f_{dec}$ is the recurrent function of the decoder, and it is just like the encoder part, $y_{t-1}$ is the embedding of the previous output token, $s_{t-1}$ is the decoder's hidden state.
-And $c$ is the context vector from the encdoer, which remains constant throughout decoding.
-\\[1ex]
-After the model computed the $s_t$, the model will compute the output probability distribution.
-\begin{equation}
-    P(y_t|y_1, c) = softmax(W_os_t+b_o)
+At each timestep $t$, the decoder updates its hidden state $s_t'$ based on the previous hidden state $s_{t'-1}$ and the embedding of the previous output $y_{t'-1}$.
+And we will define the $s_0 = c$ first.
+\begin{equation}
+    h^{dec}_t = tanh(W^{dec}_{ih}y_{t-1}+W^{dec}_{hh}+b^{dec}_h)
+\end{equation}
+After the model computed the $h^{dec}_t$, the model will compute the output probability distribution.
+\begin{equation}
+    P(y_t|y_{t-1}, c) = softmax(W_oh^{dec}_t+b^{dec}_o)
 \end{equation}
 And the predicted token $y_t = arg maxP(y_t)$, the process will repeat until the decoder predicts the end of sequence token $(<EOS>)$.
+\begin{longtable}{|c|c|c|}
+    \hline
+    \textbf{Notation} & \textbf{Meaning} & \textbf{Dimension}\\
+    \hline
+    $W^{dec}_{ih}$          & Weight matrix for input to hidden      & $hidden\ units\times input\ size$\\
+    $W^{dec}_{hh}$          & Weight matrix for hidden to hidden     & $hidden\ units\times hidden\ units$\\
+    $b^{dec}_{h}$           & Bias vector for the hidden              & $hidden\ units\times hidden\ units$\\
+    $h^{dec}_{t}$           & Hidden state at timestep t                    & $hidden\ units\times hidden\ units$\\
+    $b^{dec}_{o}$           & Bias vector for the output layer             & $hidden\ units\times hidden\ units$\\
+    $W^{dec}_{o}$           & Weight matrix for hidden to output     & $output\ feature\ space\times hidden\ units$\\
+    \hline
+    \caption{Decoder}
+\end{longtable}
 
@@
 \\[1ex]
-% This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
-\\[1ex]
-\begin{equation}
-    e_{t',t} = v_a^Ttanh(W_as_{t'-1}+U_ah_t)
+This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
+\\[1ex]
+\begin{equation}
+    e_{t',t} = v_a^Ttanh(W_ah^{dec}_{t'-1}+U_ah^{enc}_t+b_a)
 \end{equation}                    
@@
 \\[1ex]
-% This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
-\begin{equation}
-    e_{t',t} = s_{t'-1}^Th_t
+This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
+\begin{equation}
+    e_{t',t} = h^{dec}_{t'-1} \cdot h^{enc}_t
 \end{equation}
@@
 \\[1ex]
+The scaled dot-product attention was introduced by \parencite{vaswani2017attentionneed}, this is the key component in Transformer architecture.
+It improves from the Luong's dot-product attention \parencite{luong2015effectiveapproachesattentionbasedneural} by addressing he numerical stability issues when working with the high dimensional vectors and making attention computation more efficient for parallelization.
+Below is the formula of attention scores.
+\begin{equation}
+    S = \dfrac{QK^T}{\sqrt{d_k}}
+\end{equation}
+Where $S \in \mathbb{R}^{n \times n}$ is pairwise similarity scores between all tokens.
 

### 2025-03-30 paper2/main.tex
paper update

@@
 \usepackage{tikz} 
+\usepackage{amsmath}
 \usetikzlibrary{positioning, arrows.meta}
@@
 And by comparing the performance of different models.
-this report aims to provide a detailed comparative analysis of the following RNN-based models:
+this report aims to provide a detailed comparative analysis of the following RNN-based models by evaluating their performance on the IMDb dataset.
 \begin{itemize}
@@
     \item Gated Recurrent Unit (GRU)
+    \item RNN Encoder-Decoder
+    \item RNN Encoder-Decoder with Attention Mechanism
 \end{itemize}
@@
 \subsubsection{Compare Model Performances}
-Compare different RNN-based models on metrics relevant to the target NLP taskquantitatively and qualitatively. Accuracy, F1, training time scores be used.
+Compare different RNN-based models on metrics relevant to the target NLP taskquantitatively and qualitatively. For example accuracy.
 \subsubsection{Analyze the Impact of Architectural Differences}
@@
 Same setting with model 1, but I stack 2 RNN block. In the other word, this is deep recurrent neural networks. 
-\subsubsection{Model 5 (GRU with 2 layer)}
+\subsubsection{Model 5 (Gated recurrent unit with 2 layer)}
 Same setting with model 1, but I stack 2 GRU block. In the other word, this is deep recurrent neural networks. 
-\subsubsection{Model 6 (LSTM RNN with 2 layer)}
+\subsubsection{Model 6 (Long short-term memory with 2 layer)}
 Same setting with model 1, but I stack 2 LSTM block. In the other word, this is deep recurrent neural networks.
+
+\subsubsection{Model 7 (RNN Encoder-Decoder)}
+For the RNN encoder-decoder architecture, I will explain how to data be processes and I will show the mathmatical transformations for each time step.
+We first assume the input $X = [x_1, x_2, \cdots, x_T]$ be the input sequence and the $T$ is the sequence length.
+$x$ is word index in the vocabulary.
+\\[1ex]
+Then the embedding layer maps each word index $x_t$ to a high dimensional space. 
+\begin{equation}
+    e_t = Embedding(x_t)
+\end{equation}
+For each time step, the model will train the parameters $h_t$ by using the current input and the previous hidden state.
+\begin{equation}
+    h_t = tanh(W_{ih}e_t+W_{hh}h_{t-1}+b_h)
+\end{equation}
+Where $W_{ih} \in \mathbb{R}^{hidden\ dim \times embedding\ dim}$, $W_{hh} \in \mathbb{R}^{hidden\ dim \times hidden\ dim}$. $b_h \in \mathbb{R}^{hidden\ dim}$.
+\\[1ex]
+For the decoder part is do the fully connected layer.
+\begin{equation}
+    z = W_oh_T+b_o
+\end{equation}
+And the apply sigmoid activation function to find the highest probability output.
+\begin{eqnarray}
+    \hat{y} = \sigma(z) = \dfrac{1}{1+e^{-z}}
+\end{eqnarray}
+
+\subsubsection{Model 8 (RNN Encoder-Decoder with attention mechanism)}
 
@@
 The following hyperparameters were chosen based on preliminary experiments and standard practices for sentiment classification tasks:
+\\[1ex]
+For model 1 (Vanilla RNN), model 2 (Long short-term memory), model 3 (Gated recurrent unit), model 7 (RNN encoder-decoder), model 8 (RNN encoder-decoder with attention mechanism). 
 \begin{itemize}
@@
 \end{itemize}
-In order to keep the fair experiment. I train 3 RNN model with the same hyperparameters
+
+For model 4 (Vanilla RNN with 2 layer), model (GRU with 2 layer), model 6 (Long short-term memory with 2 layer)
+\begin{itemize}
+    \item \textbf{Architecture Hyperparameters:}
+    \begin{itemize}
+        \item Number of layers: \texttt{no\_layers = 2}
+        \item Vocabulary size: \(\mathtt{vocab\_size = len(vocab) + 1}\) (the extra one is for unknown word)
+        \item Embedding dimension: \texttt{embedding\_dim = 100}
+        \item Hidden dimension: \texttt{hidden\_dim = 50}
+        \item Output dimension: \texttt{output\_dim = 1}
+        \item Dropout probability: \texttt{drop\_prob = 0.5}
+        \item Batch size: \texttt{batch\_size = 50}
+    \end{itemize}
+    \item \textbf{Optimization Hyperparameters:}
+    \begin{itemize}
+        \item Learning rate: \texttt{lr = 0.001}
+        \item Loss Function: Binary Cross Entropy Loss
+        \item Optimizer: Adam with predefined learning rate.
+    \end{itemize}
+\end{itemize}
 
@@
 
+    \item \textbf{Cross Validation:}
+    In this project, a stratified k-fold cross validation was applied as part of the training loop to evaluate and compare the performance of various RNN-based model.
+And I will use the better performance of the best combination which has the highest mean accuracy.
+And this method was designed to ensure that each model was trained and validated on all data splits, providing a more comprehencive evaluation of model performance.
+    \\[1ex]
+    First the IMDB dataset was split into 5 folds (k=5) for stratified cross-validation, ensuring that each fold maintained the same proportion of positive and negative sentiment labels.
+For each fold, there is only one subset was used as the validation set, in the mean time, the rest of the subsets were combined to form as the training set.
+    \\[1ex]
+    For each fold,the model will train on the trainin set using the predefined batch size and validated on the validation set after each epoch.
+Once all the folds were completed, the performance metris were averaged to obtain the final evaluation scores for each model.
+In the other word, the average accuracy were used to compare the models and determine their overall effectiveness.
+
     \item \textbf{Validation Stage:}  
@@
 In this section, I report the classification accuracy of three model on the testing set. And I will show the performance of RNN, GRU and LSTM on IMDB dataset respectively. 
-\subsubsection{Vanilla RNN (1 layer)}
+\subsubsection{Model 1 (Vanilla RNN)}
 For the figure 2, RNN with 1 layer performance on IMDB dataset.
@@
 \end{figure}
-\subsubsection{Vanilla RNN (2 layer)}
+\subsubsection{Model 4 (Vanilla RNN with 2 layer)}
 The training accuracy improves over the eopchs, increasing steadily from 50\% to 75\%.
@@
 \end{figure}
-\subsubsection{GRU (1 layer)}
+\subsubsection{Model 3 (Gated recurrent unit)}
 For the figure 4, GRU with 1 layer, we can see that the training accuracy improves steadily and reaching around 95\%, but the validation accuracy remain the same value after epoch 2.
@@
 \end{figure}
-\subsubsection{GRU (2 layer)}
+\subsubsection{Model 5 (Gated recurrent unit with 2 layer)}
 And next is the performance of GRU with 1 layer, we can see that the training accuracy improves significantly, starting around 60\% and reaching over 95\% by the end of training.
@@
 
-\subsubsection{LSTM (1 layer)}
+\subsubsection{Model 2 (Long short-term memory)}
 And then for the training accuracy of LSTM with 1 layer, the train and validation accuracy increase steadily, for the training accuracy increase from 55\% to 90\%, for the validation loss is increase from 63\% to 83\%.
@@
 \end{figure}
-\subsubsection{LSTM (2 layer)}
+\subsubsection{Model 6 (Long short-term memory with 2 layer)}
 Next is the performance of LSTM with 1 layer, the training accuracy improves steadily over the epochs and starting at around 55\% and reaching about 90\% at the end of the training.

### 2025-03-30 paper2/main.tex
paper update

@@
 \newpage
+
+\subsubsection{Model 7 (RNN encoder-decoder)}
+\textbf{Loss over epochs (left plot)}
+\begin{enumerate}
+    \item Training Loss:
+    \begin{itemize}
+        \item The training loss consistently decreases over the epochs, which mean that the model is leaning and fitting the training data better as training progresses.
+        \item By the end of the epoch, which mean 10th epoch, the training loss drops significantly from roughtly 0.63 to 0.55.
+    \end{itemize}
+    \item Test Loss:
+    \begin{itemize}
+        \item The test loss shows more fluctuation compared to the training loss, initially, the test loss is greater that the training loss, and this is expected.
+After that, it decreases slightly in the first few epochs but then starts to fluctuate significantly, with a noticeable spike around epoch 6 and another drop by epoch 9.
+At the end of the test loss is around 0.725.
+    \end{itemize}
+\end{enumerate}
+\textbf{Accuracy Over Epochs (Right plot)}
+\begin{enumerate}
+    \item Training Accuracy:
+    \begin{itemize}
+        \item The training accuracy steadily increases from roungthly 60\% in the first epoch to above 72.5\% by the 10th epoch. 
+    \end{itemize}
+    \item Test Accuracy:
+    \begin{itemize}
+        \item Test accuracy initially improves, peaking at epoch 4 at approximately 70\%, after epoch 4, the test accuracy drop dramatically to below 60\%.
+    \end{itemize}
+\end{enumerate}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder_1_layer_performance.png}
+    \caption{LSTM with 2 layer performance on IMDB dataset}
+\end{figure}
+
+\subsubsection{Model 8 (RNN encoder-decoder with attention mechanism)}
+\textbf{Loss over epochs (Left plot)}
+\begin{enumerate}
+    \item Training Loss:
+    \begin{itemize}
+        \item The training loss cansistently decreases throughout the eopchs, starting from around 0.65 and dropping to below 0.3 by the 10th eopch.
+Which mean that the model is leaning effectively and improving its performance on the training data as training progresses.
+    \end{itemize}
+    \item Test Loss:
+    \begin{itemize}
+        \item The test loss fluctuates more than the training loss, it increases slightly in the first few epochs and peaking at around epoch 4 with a value near 0.8.
+At the end of the epoch, the loss drop back to around 0.6.
+    \end{itemize}
+\end{enumerate}
+
+\textbf{Accuracy Over Epochs (Right plot)}
+\begin{enumerate}
+    \item Training Accuracy:
+    \begin{itemize}
+        \item The training accuracy steadily increases over the epochs, starting from around 60\% and peaking at epoch 5 at 73\%. After that the accuracy drop to around 63\% dramatically. 
+    \end{itemize}
+    \item Test Accuracy:
+    \begin{itemize}
+        \item The test accuracy increases rapidly during the inital epochs, peaking at around apoch 4 with a value close to 70\%.
+After epoch 3, it drop dramatically, the shape is similar with the training accuracy.
+    \end{itemize}
+\end{enumerate}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder_with_attention_1_layer_performance.png}
+    \caption{LSTM with 2 layer performance on IMDB dataset}
+\end{figure}
+
+\newpage
 \section{Analysis}

### 2025-03-30 boxplot.py
paper update

@@ -1,0 +1,48 @@
+import yfinance as yf
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+# Define two sets of stock tickers
+set1 = ['0001.HK', '0002.HK', '0003.HK', '0005.HK','0006.HK','0011.HK','0016.HK','0027.HK','0066.HK','0101.HK','0175.HK','0267.HK','0316.HK','0386.HK','0388.HK','0669.HK','0688.HK','0700.HK','0762.HK','0823.HK','0857.HK','0883.HK','0939.HK','0941.HK','1038.HK','1044.HK','1093.HK','1177.HK','1398.HK','1928.HK','2020.HK','2313.HK','2318.HK','2319.HK','2382.HK','2388.HK','2628.HK','3988.HK']
+set2 = ['600016.ss','600196.ss','600104.ss','600019.ss','600690.ss','600519.ss','600028.ss','600010.ss','600000.ss','600036.ss','600585.ss','600050.ss','600030.ss','600900.ss','600031.ss','600089.ss','600547.ss','601988.ss','601398.ss','600048.ss','601628.ss','601166.ss','601318.ss','601328.ss','601088.ss','601857.ss','601939.ss','601601.ss','601390.ss','601169.ss','600111.ss','601899.ss','601186.ss','601668.ss','601989.ss','601607.ss','600999.ss','600887.ss','600406.ss','600893.ss','600276.ss','600570.ss','601618.ss','601788.ss','601111.ss','600009.ss','600809.ss','600660.ss','600436.ss','600309.ss','601888.ss','600438.ss','600588.ss','600760.ss','600600.ss','600176.ss','600011.ss' ,'600346.ss','600150.ss','600584.ss','600132.ss','600426.ss','601919.ss', '601600.ss','600460.ss','600085.ss','600096.ss','600188.ss','600795.ss','600875.ss','601872.ss','600026.ss','600027.ss','600160.ss','600332.ss','601766.ss','601998.ss','600895.ss','600489.ss','600418.ss','600415.ss','600118.ss','601168.ss','601006.ss','600886.ss','600482.ss','600372.ss','600362.ss','601898.ss','601699.ss','601009.ss','600845.ss','600803.ss','600674.ss','600039.ss','600018.ss','600015.ss']
+
+
+# Fetch historical data for both sets
+data1 = yf.download(set1, start='2020-01-01', end='2023-01-01')
[…]

### 2025-03-30 heatmaphsi.py

@@ -1,0 +1,27 @@
+import yfinance as yf
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+# Define the stock symbols and date range
+symbols = ['0168.HK','0175.HK','0268.HK','0270.HK','0288.HK','0316.HK','0357.HK','0358.HK','0383.HK','0390.HK','0489.HK','0525.HK','0552.HK','0568.HK','0576.HK','0586.HK','0598.HK','0670.HK','0694.HK','0728.HK','0737.HK','0753.HK','0763.HK','0777.HK','0813.HK','0902.HK','0914.HK','0936.HK','0998.HK','1288.HK','0001.HK','0002.HK','0003.HK','0004.HK','0005.HK','0006.HK','0011.HK','0012.HK','0016.HK','0017.HK','0019.HK','0023.HK','0066.HK','0083.HK','0101.HK','0135.HK','0144.HK','0151.HK','0267.HK','0291.HK','0293.HK','0330.HK','0386.HK','0388.HK','0688.HK','0700.HK','0762.HK','0823.HK','0857.HK','0883.HK','0939.HK','0941.HK','0992.HK','1038.HK','1044.HK','1088.HK','1093.HK','1109.HK','1177.HK','1299.HK','1398.HK','1928.HK','2007.HK','2318.HK','2388.HK','2628.HK','3328.HK','3988.HK','6823.HK','9618.HK','9988.HK']
+
+start_date = '2010-01-01'
+end_date = '2020-12-31'#refer to the period selected by the paper
+
+# Fetch the stock data
+data = yf.download(symbols, start=start_date, end=end_date, auto_adjust=False)['Adj Close']
+
+# Summary statistics for prices
+print("Summary Statistics for Prices:\n")
+print(data.describe())
+
+# Calculate the correlation matrix
+correlation_matrix = data.pct_change().corr()
+
+# Create a heatmap
+plt.figure(figsize=(10, 8))
+sns.cubehelix_palette(start=.5, rot=-.5, as_cmap=True)
+sns.heatmap(correlation_matrix, annot=False, xticklabels=False, yticklabels=False, cmap='crest', vmin=-1, vmax=1, center=0)
+plt.title('Correlation Heatmap of Stocks in HSI')
+plt.show()

### 2025-03-30 heatmapsse.py

@@ -1,0 +1,218 @@
+import yfinance as yf
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+# Define the stock symbols and date range
+# Choose SSE 180 constituent index as it includes a larger number of stocks and the stocks are the most representative stocks that listed on SSE
+# Only include the stocks that were being listed through 2010-2020, 97 stocks will be in the stock pool 
+symbols = ['600000.ss','600004.ss','600006.ss', '600007.ss','600008.ss','600009.ss', '600010.ss','600011.ss','600012.ss', '600015.ss','600016.ss','600017.ss', '600018.ss','600019.ss','600020.ss', '600021.ss','600022.ss','600023.ss', '600025.ss','600026.ss','600027.ss', '600028.ss','600029.ss','600030.ss', '600031.ss','600032.ss','600033.ss', '600035.ss','600036.ss','600037.ss', '600038.ss','600039.ss','600048.ss', '600050.ss','600051.ss','600052.ss', '600053.ss','600054.ss','600055.ss', '600056.ss','600057.ss','600058.ss', '600059.ss','600060.ss','600061.ss', '600062.ss','600063.ss','600064.ss', '600066.ss','600067.ss','600071.ss', '600072.ss','600073.ss','600075.ss', '600076.ss','600078.ss','600079.ss', '600080.ss','600081.ss','600082.ss', '600084.ss','600085.ss','600088.ss', '600089.ss','600094.ss','600095.ss', '600096.ss','600097.ss','600098.ss', '600099.ss','600100.ss','600101.ss', '600103.ss','600104.ss','600105.ss', '600106.ss','600107.ss','600108.ss', '600109.ss','600110.ss','600111.ss', '600113.ss','600114.ss','600115.ss', '600116.ss','600117.ss','600118.ss', '600119.ss','600120.ss','600121.ss', '600123.ss','600125.ss','600126.ss', '600127.ss','600128.ss','600129.ss', '600130.ss','600131.ss','600132.ss', '600133.ss','600135.ss','600137.ss', '600138.ss','600141.ss','600143.ss', '600148.ss','600149.ss','600150.ss', '600151.ss','600152.ss','600153.ss', '600155.ss','600156.ss','600157.ss', '600158.ss','600159.ss','600160.ss', '600161.ss','600162.ss','600163.ss
[…]

### 2025-03-30 stock_volatility.py

@@ -1,0 +1,37 @@
+import yfinance as yf
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Define the list of HSI tickers
+hsi_tickers = ['0001.HK', '0002.HK', '0003.HK', '0005.HK','0006.HK','0011.HK','0016.HK','0027.HK','0066.HK','0101.HK','0175.HK','0267.HK','0316.HK','0386.HK','0388.HK','0669.HK','0688.HK','0700.HK','0762.HK','0823.HK','0857.HK','0883.HK','0939.HK','0941.HK','1038.HK','1044.HK','1093.HK','1177.HK','1398.HK','1928.HK','2020.HK','2313.HK','2318.HK','2319.HK','2382.HK','2388.HK','2628.HK','3988.HK']
+
+# Define the list of SSE180 tickers
+sse180_tickers = ['600016.ss','600196.ss','600104.ss','600019.ss','600690.ss','600519.ss','600028.ss','600010.ss','600000.ss','600036.ss','600585.ss','600050.ss','600030.ss','600900.ss','600031.ss','600089.ss','600547.ss','601988.ss','601398.ss','600048.ss','601628.ss','601166.ss','601318.ss','601328.ss','601088.ss','601857.ss','601939.ss','601601.ss','601390.ss','601169.ss','600111.ss','601899.ss','601186.ss','601668.ss','601989.ss','601607.ss','600999.ss','600887.ss','600406.ss','600893.ss','600276.ss','600570.ss','601618.ss','601788.ss','601111.ss','600009.ss','600809.ss','600660.ss','600436.ss','600309.ss','601888.ss','600438.ss','600588.ss','600760.ss','600600.ss','600176.ss','600011.ss' ,'600346.ss','600150.ss','600584.ss','600132.ss','600426.ss','601919.ss', '601600.ss','600460.ss','600085.ss','600096.ss','600188.ss','600795.ss','600875.ss','601872.ss','600026.ss','600027.ss','600160.ss','600332.ss','601766.ss','601998.ss','600895.ss','600489.ss','600418.ss','600415.ss','600118.ss','601168.ss','601006.ss','600886.ss','600482.ss','600372.ss','600362.ss','601898.ss','601699.ss','601009.ss','600845.ss','600803.ss','600674.ss','600039.ss','600018.ss','600015.ss']
+
+start_date = '2010-01-01'
+end_date = '2020-12-31'
+
+# Function to calculate volatility
+def calculate_volatility(tickers):
+    volatility_values = []
[…]

### 2025-03-31 paper2/main.tex
finish

@@
     \begin{itemize}
-        \item The training accuracy steadily increases over the epochs, starting from around 60\% and peaking at epoch 5 at 73\%. After that the accuracy drop to around 63\% dramatically. 
+        \item The training accuracy steadily increases over the epochs, starting from around 60\% and raise to 90\% by the end of epoch.
+This demonstrates that the model becomes increaseingly proficient at predicting the correct labels for the training data.
     \end{itemize}
@@
     \begin{itemize}
-        \item The test accuracy increases rapidly during the inital epochs, peaking at around apoch 4 with a value close to 70\%.
-After epoch 3, it drop dramatically, the shape is similar with the training accuracy.
+        \item The test accuracy steadily increase from 72\% to 85\%. 
     \end{itemize}
@@
 \end{figure}
-
 \newpage
-\section{Analysis}
-\subsection{Learning Behavior}
-All models show the effective learning during the initial epochs, roughtly 55\%, as both training and validation accuracy improve, and loss decrease.
-But the common problem is overfitting occure in 3 models.
-The possible reasons is dataset.
-Since every sentence in dataset was truncated.
-I just use the first 100 token to training the model.
-So the existing tokens might not be fully represent the sentence.
-The second reason is the embedding layers, I used the built-in function in Pytorch to create the embedding layer by randomise the matrix.
-\\[1ex]
-Among the 3 models, LSTM and GRU have the better performance compare with vanilla RNN.
-The reason of GRU and LSTM perform better than RNN is the vanilla RNN struggles with long term dependencies because of the vanishing gradient problem when processing long sequences.
-In this experience is 100 time step.
-And this limits its ability to learn the meaningful patterns from data.
-On the contrast, GRU and LSTM use gate mechanisms to retain the relevant information and discard irrelevant information, allowing them to capture the long term dependencies more effectively.
-\\[1ex]
-And for the overfitting problem, we can see that overfitting problem occur in three models.
-For the vanilla RNN, since this is the basic and simple architecture to handle sequential data, it lacks mechanisms to retain or forget information selectively.
-As a result, it relying on the memorizing the training data rather than learning the meaningful patterns.
-This causes it to overfit.
-And for the GRU and LSTM, although gating mechanisms provide a more sophisticated architecture to learning from data, they aggressively optimizes the training loss, which can result in overfitting when the model continues to learn unnecessary information from the dataset.
-\\[1ex] 
-The above result and analyse shown that GRU and LSTM perform better than vanilla RNN, which show that the graident vanishing and graident exploding problem were improved by the gating mechanism.
-And there is so many way will affect the experiment result for example, the embedding layer, tokenization.
+
+\subsubsection{Word Importancy}
+\textbf{Heatmap}
+\\[1ex]
+This heatmap is from rnn encoder decoder with attention mechanism, I extract the embedding layer and the hidden state from the encoder, and calculate the dot product between the weight matrix and the sentence vector.
+In this heatmap, each row corresponds to a hidden state dimension, and each column represents a word.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/hidden_state.png}
+    \caption{Heatmap of hidden state}
+\end{figure}
+\newline
+Different dimensions in the hidden state activate for different words.
+For example, the words "highly" and "recommend" have bright activations in specific dimensions, indicating that these dimensinos are capturing the positive sentiment or recommendation context, the words like "is", "a" and "that" show weaker, less distinct activations, as they carry less information in the sentence.
+The word "must-watch" show the unique parttern, the most possible reason is that it is a compound word with significant semantic importance in the sentence.
 
 \newpage
-
+\textbf{T-SNE visualization of hidden state}
+\\[1ex]
+This graph visualizaes the hidden states of the RNN encoder-decoder with attention after processing the sentence.
+The words are plotted in a 2 dimensional space by using t-SNE, which reduces the high dimensional hidden states to 2 dimensions while preserving their information.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{../Pic/hidden_state_2.png}
+    \caption{t-SNE}
+\end{figure}
+\newline
+This plot is similar with the heatmap, is to show the relationship of how the words can be cluster together.
+The words like "highly", "recommend" and "great" are positioned close to each other, it is because they are all refer to the sentence's positive sentiment or recommendation context.
+Overall, the t-SNE plot shows how the model encodes the semantic and syntactic relationship between words.
+Words with similar meanings or roles in the sentence are encoded in similar regions of the hidden state space.
+
+\newpage
+\section{Discussion and Conclusion}
+\subsection{Results and Analysis}
+The initial epochs are all models that started effective learning shown with approximately 55\% training and validation accuracy and the loss decreased consistently.
+A major problem that always arose across all models was overfitting.
+The primary reason behind the emergence of this phenomenon can be attributed to most of the dataset’s characteristics, especially truncation of sentences.
+The input to the models is restricted to the first 100 tokens and therefore, the models adjust to this input and may not fully capture the important context of each sentence.
+Furthermore the failure to generalize might be further loudened by the fact that the embedding matrix is defined using a built in layer in PyTorch with its values initialized to be randomly.
+
+\subsection{Model Comparisons}
+The three models analysed - vanilla RNN, LSTM and GRU, showed a superior performance of GRU and LSTM models.
+The underlying cause for this improvement however, is due to the limitation of the vanilla RNN at propagating long term dependencies (vanishing gradient problem) using 100 time steps of sequence processing.
+Its ability to effectively learn meaningful patterns of data allowed for this limitation.
+LSTM and GRU, on the other hand, utilize gating mechanism to retain relevant information and discard irrelevant information so that long term dependency can be captured more effectively.
+
+\subsection{Overfitting Issues}
+All of the three models were susceptible to the overfitting problem.
+However, as simple as the vanilla RNN is, it lacks advanced mechanisms of information retention and selective forgetting, and so instead of extracting the meaningful patterns embedded in the training data, it relies on memorizing it.
+Consequently, this results in overfitting.
+GRU and LSTM introduce gating mechanisms for data learning and are able to well learn from data, however, they do better optimization of training loss which can result to over fitting when they continue to learn unnecessary information from dataset.
+
+\subsection{Deep Recurrent Neural Networks (DRNN)}
+The performance of each model shows that the deep recurrent neural networks (DRNNs) is outperforms with the single layer models.
+When we stack more layer, then the model will become DRNNs and this architecture allows it to learn more complex and abstract representations of the data, in other word, the model able to capture or extract more useful information from the data.
+Thereby reducing the impact of overfitting.
+
+\subsection{Attention Mechanism}
+For the RNN encoder-decoder, I have applied attention mechanism, and among all of the models I have evaluated, the attention mechanism stood out as the most effective and the highest accuracy.
+The attention mechanism enhances the performance by allowing the model to focus on relevant parts of the input sequence, effectively retaining and utilizing crucial contextual information.
+
+\subsection{Gradient Problem}
 \begin{figure}[h!]
@@
 
-\newpage
-
 % ######################## Results and Analysis ######################################
@@
 \subsection{Discussion}
-During the experimental phase, several insights and challenges emerged, which are discussed below:
+After I run all the model and compare their performance, serveral insights and challenges appear, which are discussed below.
 
@@
 
+    \item \textbf{Expected Results:}
+    It is expected that the attention mechanism is outperformance comparing with other model since the attention mechanism is designed for capture the information from long sentence.
+So this is the expected outcome.
+
     \item \textbf{Limitations:}  
@@
     \item \textbf{Comparative Analysis:}  
+    \\[1ex]
     Different RNN-based models, including vanilla RNN, LSTM, and GRU, were implemented and evaluated.
@@
     \item \textbf{Efficacy of Gated Architectures:}  
+    \\[1ex]
     The results indicate that the gated architectures (LSTM and GRU) significantly outperform the vanilla RNN in terms of handling vanishing and exploding gradient issues, thereby offering more stable and reliable performance.
-\end{itemize}
-
-In conclusion, while some unexpected results were observed—such as the higher loss in the two-layer LSTM model—the overall findings strongly advocate for the use of gated RNN architectures in tasks requiring modeling of long-term dependencies.
-Despite limitations due to computational resources, this report provides a valuable comparative study and lays the groundwork for future research on more complex models like Transformers in natural language processing tasks.
-
-In conclusion, while the project revealed certain limitations, it nevertheless contributes valuable insights into the training and evaluation of RNN-based models for sentiment classification, thereby setting the stage for further research in the domain of efficient and scalable NLP methodologies.
+
+    \item \textbf{Importancy of attention mechanism}
+    \\[1ex]
+    We have plot the heatmap of the hidden state of the RNN encoder-decoder with attention mechanism, also i have plotted the t-SNE to show how the word be cluter together.
+And this 2 graph can illustrate how to attention mechanism train the model and train the parameter inside the model.
+\end{itemize}
+
 % ######################## Discussion and Conclusion ######################################

### 2025-03-31 heatmapsse.py

@@ -98,107 +98,10 @@ symbols = ['600000.ss','600004.ss','600006.ss', '600007.ss','600008.ss','600009.
 '603386.ss','603387.ss','603388.ss',
 '603389.ss','603390.ss','603391.ss',
 '603392.ss','603393.ss','603394.ss',
-'603395.ss','603396.ss','603397.ss',
-'603398.ss','603399.ss','603400.ss','603401.ss','603402.ss', '603403.ss','603404.ss','603405.ss', '603406.ss','603407.ss','603408.ss', '603409.ss','603410.ss','603411.ss', '603412.ss','603413.ss','603414.ss', '603415.ss','603416.ss','603417.ss', '603418.ss','603419.ss','603420.ss', '603421.ss','603422.ss','603423.ss', '603424.ss','603425.ss','603426.ss', '603427.ss','603428.ss','603429.ss', '603430.ss','603431.ss','603432.ss', '603433.ss','603434.ss','603435.ss', '603436.ss','603437.ss','603438.ss', '603439.ss','603440.ss','603441.ss', '603442.ss','603443.ss','603444.ss', '603445.ss','603446.ss','603447.ss', '603448.ss','603449.ss','603450.ss', '603451.ss','603452.ss','603453.ss', '603454.ss','603455.ss','603456.ss', '603457.ss','603458.ss','603459.ss', '603460.ss','603461.ss','603462.ss', '603463.ss','603464.ss','603465.ss', '603466.ss','603467.ss','603468.ss', '603469.ss','603470.ss','603471.ss', '603472.ss','603473.ss','603474.ss', '603475.ss','603476.ss','603477.ss', '603478.ss','603479.ss','603480.ss', '603481.ss','603482.ss','603483.ss', '603484.ss','603485.ss','603486.ss', '603487.ss','603488.ss','603489.ss', '603490.ss','603491.ss','603492.ss', '603493.ss','603494.ss','603495.ss', '603496.ss','603497.ss','603498.ss', '603499.ss','603500.ss', '603501.SS','603505.SS','603506.SS','603507.SS','603508.SS','603511.SS','603515.SS','603516.SS','603517.SS','603518.SS',
-'603519.SS','603520.SS','603521.SS','603527.SS','603528.SS','603529.SS','603530.SS','603533.SS','603535.SS','603536.SS',
[…]

### 2025-04-01 boxplot.py
paper update

@@ -1,48 +1,0 @@
-import yfinance as yf
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-
-# Define two sets of stock tickers
-set1 = ['0001.HK', '0002.HK', '0003.HK', '0005.HK','0006.HK','0011.HK','0016.HK','0027.HK','0066.HK','0101.HK','0175.HK','0267.HK','0316.HK','0386.HK','0388.HK','0669.HK','0688.HK','0700.HK','0762.HK','0823.HK','0857.HK','0883.HK','0939.HK','0941.HK','1038.HK','1044.HK','1093.HK','1177.HK','1398.HK','1928.HK','2020.HK','2313.HK','2318.HK','2319.HK','2382.HK','2388.HK','2628.HK','3988.HK']
-set2 = ['600016.ss','600196.ss','600104.ss','600019.ss','600690.ss','600519.ss','600028.ss','600010.ss','600000.ss','600036.ss','600585.ss','600050.ss','600030.ss','600900.ss','600031.ss','600089.ss','600547.ss','601988.ss','601398.ss','600048.ss','601628.ss','601166.ss','601318.ss','601328.ss','601088.ss','601857.ss','601939.ss','601601.ss','601390.ss','601169.ss','600111.ss','601899.ss','601186.ss','601668.ss','601989.ss','601607.ss','600999.ss','600887.ss','600406.ss','600893.ss','600276.ss','600570.ss','601618.ss','601788.ss','601111.ss','600009.ss','600809.ss','600660.ss','600436.ss','600309.ss','601888.ss','600438.ss','600588.ss','600760.ss','600600.ss','600176.ss','600011.ss' ,'600346.ss','600150.ss','600584.ss','600132.ss','600426.ss','601919.ss', '601600.ss','600460.ss','600085.ss','600096.ss','600188.ss','600795.ss','600875.ss','601872.ss','600026.ss','600027.ss','600160.ss','600332.ss','601766.ss','601998.ss','600895.ss','600489.ss','600418.ss','600415.ss','600118.ss','601168.ss','601006.ss','600886.ss','600482.ss','600372.ss','600362.ss','601898.ss','601699.ss','601009.ss','600845.ss','600803.ss','600674.ss','600039.ss','600018.ss','600015.ss']
-
-
-# Fetch historical data for both sets
-data1 = yf.download(set1, start='2020-01-01', end='2023-01-01')
[…]

### 2025-04-01 heatmaphsi.py

@@ -1,27 +1,0 @@
-import yfinance as yf
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-
-# Define the stock symbols and date range
-symbols = ['0168.HK','0175.HK','0268.HK','0270.HK','0288.HK','0316.HK','0357.HK','0358.HK','0383.HK','0390.HK','0489.HK','0525.HK','0552.HK','0568.HK','0576.HK','0586.HK','0598.HK','0670.HK','0694.HK','0728.HK','0737.HK','0753.HK','0763.HK','0777.HK','0813.HK','0902.HK','0914.HK','0936.HK','0998.HK','1288.HK','0001.HK','0002.HK','0003.HK','0004.HK','0005.HK','0006.HK','0011.HK','0012.HK','0016.HK','0017.HK','0019.HK','0023.HK','0066.HK','0083.HK','0101.HK','0135.HK','0144.HK','0151.HK','0267.HK','0291.HK','0293.HK','0330.HK','0386.HK','0388.HK','0688.HK','0700.HK','0762.HK','0823.HK','0857.HK','0883.HK','0939.HK','0941.HK','0992.HK','1038.HK','1044.HK','1088.HK','1093.HK','1109.HK','1177.HK','1299.HK','1398.HK','1928.HK','2007.HK','2318.HK','2388.HK','2628.HK','3328.HK','3988.HK','6823.HK','9618.HK','9988.HK']
-
-start_date = '2010-01-01'
-end_date = '2020-12-31'#refer to the period selected by the paper
-
-# Fetch the stock data
-data = yf.download(symbols, start=start_date, end=end_date, auto_adjust=False)['Adj Close']
-
-# Summary statistics for prices
-print("Summary Statistics for Prices:\n")
-print(data.describe())
-
-# Calculate the correlation matrix
-correlation_matrix = data.pct_change().corr()
-
-# Create a heatmap
-plt.figure(figsize=(10, 8))
-sns.cubehelix_palette(start=.5, rot=-.5, as_cmap=True)
-sns.heatmap(correlation_matrix, annot=False, xticklabels=False, yticklabels=False, cmap='crest', vmin=-1, vmax=1, center=0)
-plt.title('Correlation Heatmap of Stocks in HSI')
-plt.show()

### 2025-04-01 heatmapsse.py

@@ -1,121 +1,0 @@
-import yfinance as yf
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-
-# Define the stock symbols and date range
-# Choose SSE 180 constituent index as it includes a larger number of stocks and the stocks are the most representative stocks that listed on SSE
-# Only include the stocks that were being listed through 2010-2020, 97 stocks will be in the stock pool 
-symbols = ['600000.ss','600004.ss','600006.ss', '600007.ss','600008.ss','600009.ss', '600010.ss','600011.ss','600012.ss', '600015.ss','600016.ss','600017.ss', '600018.ss','600019.ss','600020.ss', '600021.ss','600022.ss','600023.ss', '600025.ss','600026.ss','600027.ss', '600028.ss','600029.ss','600030.ss', '600031.ss','600032.ss','600033.ss', '600035.ss','600036.ss','600037.ss', '600038.ss','600039.ss','600048.ss', '600050.ss','600051.ss','600052.ss', '600053.ss','600054.ss','600055.ss', '600056.ss','600057.ss','600058.ss', '600059.ss','600060.ss','600061.ss', '600062.ss','600063.ss','600064.ss', '600066.ss','600067.ss','600071.ss', '600072.ss','600073.ss','600075.ss', '600076.ss','600078.ss','600079.ss', '600080.ss','600081.ss','600082.ss', '600084.ss','600085.ss','600088.ss', '600089.ss','600094.ss','600095.ss', '600096.ss','600097.ss','600098.ss', '600099.ss','600100.ss','600101.ss', '600103.ss','600104.ss','600105.ss', '600106.ss','600107.ss','600108.ss', '600109.ss','600110.ss','600111.ss', '600113.ss','600114.ss','600115.ss', '600116.ss','600117.ss','600118.ss', '600119.ss','600120.ss','600121.ss', '600123.ss','600125.ss','600126.ss', '600127.ss','600128.ss','600129.ss', '600130.ss','600131.ss','600132.ss', '600133.ss','600135.ss','600137.ss', '600138.ss','600141.ss','600143.ss', '600148.ss','600149.ss','600150.ss', '600151.ss','600152.ss','600153.ss', '600155.ss','600156.ss','600157.ss', '600158.ss','600159.ss','600160.ss', '600161.ss','600162.ss','600163.ss
[…]
